### 2020-01-25 r-journal/manuscript.Rmd
:sparkles:

---
title: "Partitioned Local Depth (PaLD) Clustering Analyses in R"
author:
  - name: Lucy D'Agostino McGowan
    affiliation: Wake Forest University
    address:
    - Winston-Salem, NC
    - 27106
    email:  mcgowald@wfu.edu
  - name: Katherine Moore
    affiliation: Wake Forest Unversity
    address:
    - Winston-Salem, NC
    - 27106
    email: mooreke@wfu.edu 
  - name: Kenneth Berenhaut 
    affiliation: Wake Forest University
    address:
    - Winston-Salem, NC
    - 27106
    email: berenhks@wfu.edu 
abstract: >
  An abstract of less than 150 words.
output:
  rticles::rjournal_article:
    includes:
      in_header: preamble.tex
---

```{r, echo = FALSE}
knitr::opts_chunk$set(fig.height = 4, message = FALSE, warning = FALSE)
```

## Introduction

* Describe PaLD (Ken & Kate?)
* Cite PaLD

We present a new package, \CRANpkg{pald}, for calculating partitioned local depth (PaLD) probabilities, implementing clustering analyses, and creating data visualizations to represent the clusters. This paper will describe how to use the package as well as walk through two examples.

## pald

The main functions in \CRANpkg{pald} package can be split into 3 categories:

1. Helper functions to organize data into the correct format, into distance matrices and then contribution matrices
2. Functions that convert a contribution matrix into a variety of useful formats, including partitioned local depths, clusters, and graphs
3. Plotting functions

In addition, the package provides a number of pertinent example data sets commonly used to demonstrate cluster algorithms, including a synthetic data set of two-dimensional points created by \citet{gionis1clustering} to demonstrate clustering aggregation, a sample of walking distances from a pump in the infamous cholera outbreak \citep{cholera}, clustering data generated from the scikit-learn Python package \citep{pedregosa2011scikit}, and data compiled by \cite{tissue} of tissue gene expressions.

While it is not a necessity, the \CRANpkg{pald} package is designed to function well with the pipe operator, `%>%`, from the \citet{magrittr} \CRANpkg{magrittr} package in that the first argument of each function is the data. This functionality will be demonstrated below.

### Helper functions to create contribution matrix

For demonstration purposes, below is a sample data frame with two variables, `x1` and `x2`. The methods put forth here work on data frames with higher dimensions, as described in the **Examples** section, we are simply choosing a small data frame here for demonstration purposes.

```{r}
library(pald)
d <- data.frame(
  x1 = c(6, 8, 11, 16, 4),
  x2 = c(5, 4, 13, 7, 18)
)
rownames(d) <- c("A", "B", "C", "D", "E")
```

The first step needed to calculate the partitioned local depths is to construct a _distance matrix_. If the data are already in this form, the user can skip to the next step. The `get_distance_matrix()` function converts an input data frame into a distance matrix, as demonstrated below.

```{r}
get_distance_matrix(d)
```
This will create an $n\times n$ distance matrix, where $n$ corresponds to the number of observations in the original data frame, in this example $n = 5$. By default, the distance matrix is _scaled_, is possible to create a distance matrix that is not scaled by changing the `scale` argument to `FALSE`.

```{r}
get_distance_matrix(d, scale = FALSE)
```

 As mentioned previously, the functions in this package are designed to work with the \CRANpkg{magrittr} pipe operator, `%>%`, so the same code as above could be written utilizing this format.

```{r}
d %>%
  get_distance_matrix()
```

This distance matrix can then be passed to the `get_contribution_matrix()` function in order to calculate a matrix of contributions. Again, if the user begins with a distance matrix, they can skip the first step and simple input the contribution matrix into this function.

```{r}
distance_matrix <- get_distance_matrix(d)
get_contribution_matrix(distance_matrix)
```

Again, the \CRANpkg{magrittr} pipe can be used as follows.

```{r}
d %>%
  get_distance_matrix() %>%
  get_contribution_matrix()
```

The _contribution matrix_ output by the `get_contribution_matrix()` is the main input for the majority of the remaining functions.

### Functions that convert a contribution matrix into useful formats

From the _contribution matrix_, a variety of useful quantities can be calculated. Below, we create a contribution matrix using the functions described in the previous section.

```{r}
d %>%
  get_distance_matrix() %>%
  get_contribution_matrix() -> contribution_matrix
```

To calculate the _clusters_ that each point will fall into, we can use the `get_clusters()` function. This will output a data frame with two columns, the first will correspond to the `point`, as identified by the row name of the original input data frame, `d`, the second will identify the `cluster` that each point belongs to.

```{r}
get_clusters(contribution_matrix)
```

In this example, three clusters are identified with these five points. Points `A` and `B` fall into cluster 1. Points `C` and `D` into cluster 2, and point `E` in cluster 3.

The `get_depths()` function calculates the _depths_ of each point, outputting a data frame with two columns, `point` indicating the point, as identified by the row name of the original data frame, `d`, and `depth` indicating the depth of the point. The data frame is arranged by depth, with the deepest point listed first.

```{r}
get_depths(contribution_matrix)
```

In this case, the deepest point is `C`.

The `get_bound()` function will calculate the _bound_ of the contribution matrix.

```{r}
get_bound(contribution_matrix)
```

In this case, the bound is 0.14.

The `any_isolated()` function will check whether there are any isolated points that will inadvertently be dropped by a graph.

```{r}
any_isolated(contribution_matrix)
```

In this case, there are no isolated points.

The `get_pald_cluster_matrix()` will calculate a matrix of partitioned local depth clusters. This function contains an argument `keep_all_edges` that indicates whether all edges should be kept. The default value is `FALSE`, indicating that edges that are less than the expectation will be dropped, allowing a clear clustering of points. Changing this argument to `TRUE` results in the pairwise minimum of the contribution matrix and the transpose of the contribution matrix.

```{r}
get_pald_cluster_matrix(contribution_matrix, keep_all_edges = TRUE)
```

Leaving the `keep_all_edges` as the default, `FALSE` will take this pairwise minimum seen above and drop the edges that are less than the expectation.

```{r}
get_pald_cluster_matrix(contribution_matrix)
```

Finally, the `get_pald_graph()` function takes the contribution matrix and creates an \CRANpkg{igraph} object, a graph that describes the relationship between the points.

```{r}
get_pald_graph(contribution_matrix)
```

Here we see that there are two connected components, points `A` and `B`, which form the first cluster, and points `D` and `C` which form the second.

## Plotting functions

The final category of function is functions for data visualization. We can begin by visualizing the points in data frame `d` (Figure \ref{fig:fig1}).

```{r fig1, fig.cap = "Visualize the points from data frame `d`", message = FALSE, warning = FALSE}
library(ggplot2)
ggplot(d, aes(x1, x2)) +
  geom_text(label = rownames(d))
```

We can then pass the contribution matrix to the `plot_pald_graph()` function to view the relationship between points (Figure \ref{fig:fig2}). By default, this will invoke the `layout_nicely()` function from \CRANpkg{igraph} to determine the layout of the graph.

```{r fig2, fig.cap = "PaLD graph displaying the relationship between the points in data frame `d`"}
d %>%
  get_distance_matrix() %>%
  get_contribution_matrix() %>%
  plot_pald_graph()
```

The `layout` argument allows the user to pass a matrix to dictate the layout of the graph. For example, if we wanted the graph to match the visualization displayed in Figure \ref{fig:fig1}, we can pass `as.matrix(d)`, or a matrix of the data frame `d` to the `layout` argument  (Figure \ref{fig:fig3}.  
```{r fig3, fig.cap = "PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1"}
d %>%
  get_distance_matrix() %>%
  get_contribution_matrix() %>%
  plot_pald_graph(layout = as.matrix(d))
```

This `plot_pald_graph()` function will also permit parameters that can be passed to the `plot.igraph()` function. For example, to add axes to the graph, the user can pass the `axes = TRUE` argument to the `...` in the `plot_pald_graph()` function (Figure \ref{fig:fig4}).

```{r fig4, fig.cap = "PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1, adding axes"}
d %>%
  get_distance_matrix() %>%
  get_contribution_matrix() %>%
  plot_pald_graph(layout = as.matrix(d),
                  axes = TRUE)
```

## Examples

We will demonstrate the utility of the \CRANpkg{pald} package in two clustering examples.

### Clustering tissue gene expression data

The first example data frame is a subset of tissue gene expression data from \citet{zilliox2007gene}, \citet{mccall2011gene}, and \citet{mccall2014gene}, obtained from the **tissuesGeneExpression** bioconductor package \citep{tissue}. This data set is included in the \CRANpkg{pald} package in an object called `tissue`.

The `tissue` object is a matrix with 189 rows, each with a corresponding tissue, such as `colon`, `kidney` or `cerebellum`. There are 22,215 columns corresponding to gene expression data from each of these rows.

We will first calculate the contribution matrix for this data frame.

```{r}
tissue_contribution_matrix <- tissue %>%
  get_distance_matrix(scale = FALSE) %>%
  get_contribution_matrix()
```

We can then use this contribution matrix to display the relationship between tissue samples using the `plot_pald_graph()` function (Figure \ref{fig:fig5}).

```{r fig5, fig.cap = "PaLD clustering of tissue data"}
plot_pald_graph(tissue_contribution_matrix)
```

The `get_clusters()` function can be used to identify the clusters of each tissue sample. Since the output is a data frame, we can summarize the clusters using commonly used data analysis techniques. For demonstration purposes, we will use the \CRANpkg{dplyr} package to summarize the contribution of clusters.

```{r}
library(dplyr)
get_clusters(tissue_contribution_matrix) %>%
  group_by(cluster, point) %>%
  count()
```

From this, we can glean that cluster one consists of two types of tissue, the kidney and endometrium. Cluster two is comprised of only the hippocampus.

### Clustering generated data

The \CRANpkg{pald} includes two data frames generated from the scikit-learn Python package \citep{pedregosa2011scikit}, `noisy_moons` and `noisy_circles`.

The `noisy_moons` data frame consists of 500 rows and two columns.

```{r}
moons_contribution_matrix <- noisy_moons %>%
  get_distance_matrix() %>%
  get_contribution_matrix()
```

When plotting the `noisy_moons` PaLD graph, we want the layout to match the layout of the original data, so we will pass `as.matrix(noisy_moons)` to the `layout` parameter in the `plot_pald_graph()` function. Additionally, here the row names are meaningless, they just correspond to the location of the generated data, so we can remove the labels on the plot by passing `vertex.label = NA` to the `...` of the `plot_pald_graph()` function.

```{r, eval = FALSE}
plot_pald_graph(moons_contribution_matrix,
                layout = as.matrix(noisy_moons),
                vertex.label = NA)
```

The `noisy_circles` data frame consists of 500 rows and two columns. We can create the contribution matrix for this data frame using the same methods as used for the `noisy_moons` data.

```{r}
circles_contribution_matrix <- noisy_circles %>%
  get_distance_matrix() %>%
  get_contribution_matrix()
```

Similarly, we will pass the layout and remove vertex labels for the noisy circles PaLD plot.

```{r, eval = FALSE}
plot_pald_graph(circles_contribution_matrix,
                layout = as.matrix(noisy_circles),
                vertex.label = NA)
```

Because these are \CRANpkg{igraph} objects, they can be combined as you would normally combine a plot. For example, we could add the line `par(mfrow = c(1, 2))` above the plot calls to create an output where both plots are displayed side by side (Figure \ref{fig:fig6}).

```{r fig6, fig.cap = "PaLD clustering of noisy moons (left) and noisy circles (right) data"}
par(mfrow = c(1, 2))
plot_pald_graph(moons_contribution_matrix,
                layout = as.matrix(noisy_moons),
                vertex.label = NA)
plot_pald_graph(circles_contribution_matrix,
                layout = as.matrix(noisy_circles),
                vertex.label = NA)
```

The ability of the PaLD algorithm to discern clusters is demonstrated here.

## Summary

This paper introduces the \CRANpkg{pald} package, demonstrating it's utility for providing a parameter-free clustering algorithm that can easily be applied to any data set.

\bibliography{RJreferences}

### 2020-01-25 r-journal/manuscript.tex
:sparkles:

% !TeX root = RJwrapper.tex
\title{Partitioned Local Depth (PaLD) Clustering Analyses in R}
\author{by Lucy D'Agostino McGowan, Katherine Moore, Kenneth Berenhaut}

\maketitle

\abstract{%
An abstract of less than 150 words.
}

% Any extra LaTeX you need in the preamble

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

\begin{itemize}
\tightlist
\item
  Describe PaLD (Ken \& Kate?)
\item
  Cite PaLD
\end{itemize}

We present a new package, \CRANpkg{pald}, for calculating partitioned
local depth (PaLD) probabilities, implementing clustering analyses, and
creating data visualizations to represent the clusters. This paper will
describe how to use the package as well as walk through two examples.

\hypertarget{pald}{%
\subsection{pald}\label{pald}}

The main functions in \CRANpkg{pald} package can be split into 3
categories:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Helper functions to organize data into the correct format, into
  distance matrices and then contribution matrices
\item
  Functions that convert a contribution matrix into a variety of useful
  formats, including partitioned local depths, clusters, and graphs
\item
  Plotting functions
\end{enumerate}

In addition, the package provides a number of pertinent example data
sets commonly used to demonstrate cluster algorithms, including a
synthetic data set of two-dimensional points created by
\citet{gionis1clustering} to demonstrate clustering aggregation, a
sample of walking distances from a pump in the infamous cholera outbreak
\citep{cholera}, clustering data generated from the scikit-learn Python
package \citep{pedregosa2011scikit}, and data compiled by \cite{tissue}
of tissue gene expressions.

While it is not a necessity, the \CRANpkg{pald} package is designed to
function well with the pipe operator, \texttt{\%\textgreater{}\%}, from
the \citet{magrittr} \CRANpkg{magrittr} package in that the first
argument of each function is the data. This functionality will be
demonstrated below.

\hypertarget{helper-functions-to-create-contribution-matrix}{%
\subsubsection{Helper functions to create contribution
matrix}\label{helper-functions-to-create-contribution-matrix}}

For demonstration purposes, below is a sample data frame with two
variables, \texttt{x1} and \texttt{x2}. The methods put forth here work
on data frames with higher dimensions, as described in the
\textbf{Examples} section, we are simply choosing a small data frame
here for demonstration purposes.

\begin{Schunk}
\begin{Sinput}
library(pald)
d <- data.frame(
  x1 = c(6, 8, 11, 16, 4),
  x2 = c(5, 4, 13, 7, 18)
)
rownames(d) <- c("A", "B", "C", "D", "E")
\end{Sinput}
\end{Schunk}

The first step needed to calculate the partitioned local depths is to
construct a \emph{distance matrix}. If the data are already in this
form, the user can skip to the next step. The
\texttt{get\_distance\_matrix()} function converts an input data frame
into a distance matrix, as demonstrated below.

\begin{Schunk}
\begin{Sinput}
get_distance_matrix(d)
\end{Sinput}
\begin{Soutput}
#>           A         B        C        D        E
#> A 0.0000000 0.4584177 1.717380 2.158418 2.229205
#> B 0.4584177 0.0000000 1.644296 1.778777 2.505929
#> C 1.7173802 1.6442957 0.000000 1.468398 1.713327
#> D 2.1584182 1.7787772 1.468398 0.000000 3.158040
#> E 2.2292047 2.5059291 1.713327 3.158040 0.000000
\end{Soutput}
\end{Schunk}

This will create an \(n\times n\) distance matrix, where \(n\)
corresponds to the number of observations in the original data frame, in
this example \(n = 5\). By default, the distance matrix is
\emph{scaled}, is possible to create a distance matrix that is not
scaled by changing the \texttt{scale} argument to \texttt{FALSE}.

\begin{Schunk}
\begin{Sinput}
get_distance_matrix(d, scale = FALSE)
\end{Sinput}
\begin{Soutput}
#>           A         B        C         D         E
#> A  0.000000  2.236068 9.433981 10.198039 13.152946
#> B  2.236068  0.000000 9.486833  8.544004 14.560220
#> C  9.433981  9.486833 0.000000  7.810250  8.602325
#> D 10.198039  8.544004 7.810250  0.000000 16.278821
#> E 13.152946 14.560220 8.602325 16.278821  0.000000
\end{Soutput}
\end{Schunk}

As mentioned previously, the functions in this package are designed to
work with the \CRANpkg{magrittr} pipe operator,
\texttt{\%\textgreater{}\%}, so the same code as above could be written
utilizing this format.

\begin{Schunk}
\begin{Sinput}
d %>%
  get_distance_matrix()
\end{Sinput}
\begin{Soutput}
#>           A         B        C        D        E
#> A 0.0000000 0.4584177 1.717380 2.158418 2.229205
#> B 0.4584177 0.0000000 1.644296 1.778777 2.505929
#> C 1.7173802 1.6442957 0.000000 1.468398 1.713327
#> D 2.1584182 1.7787772 1.468398 0.000000 3.158040
#> E 2.2292047 2.5059291 1.713327 3.158040 0.000000
\end{Soutput}
\end{Schunk}

This distance matrix can then be passed to the
\texttt{get\_contribution\_matrix()} function in order to calculate a
matrix of contributions. Again, if the user begins with a distance
matrix, they can skip the first step and simple input the contribution
matrix into this function.

\begin{Schunk}
\begin{Sinput}
distance_matrix <- get_distance_matrix(d)
get_contribution_matrix(distance_matrix)
\end{Sinput}
\begin{Soutput}
#>        1      2     3     4      5
#> A 0.2875 0.1625 0.000 0.050 0.0000
#> B 0.1750 0.3000 0.050 0.050 0.0000
#> C 0.0000 0.0625 0.300 0.175 0.0500
#> D 0.0500 0.0500 0.175 0.300 0.0000
#> E 0.0000 0.0000 0.050 0.000 0.2125
\end{Soutput}
\end{Schunk}

Again, the \CRANpkg{magrittr} pipe can be used as follows.

\begin{Schunk}
\begin{Sinput}
d %>%
  get_distance_matrix() %>%
  get_contribution_matrix()
\end{Sinput}
\begin{Soutput}
#>        1      2     3     4      5
#> A 0.2875 0.1625 0.000 0.050 0.0000
#> B 0.1750 0.3000 0.050 0.050 0.0000
#> C 0.0000 0.0625 0.300 0.175 0.0500
#> D 0.0500 0.0500 0.175 0.300 0.0000
#> E 0.0000 0.0000 0.050 0.000 0.2125
\end{Soutput}
\end{Schunk}

The \emph{contribution matrix} output by the
\texttt{get\_contribution\_matrix()} is the main input for the majority
of the remaining functions.

\hypertarget{functions-that-convert-a-contribution-matrix-into-useful-formats}{%
\subsubsection{Functions that convert a contribution matrix into useful
formats}\label{functions-that-convert-a-contribution-matrix-into-useful-formats}}

From the \emph{contribution matrix}, a variety of useful quantities can
be calculated. Below, we create a contribution matrix using the
functions described in the previous section.

\begin{Schunk}
\begin{Sinput}
d %>%
  get_distance_matrix() %>%
  get_contribution_matrix() -> contribution_matrix
\end{Sinput}
\end{Schunk}

To calculate the \emph{clusters} that each point will fall into, we can
use the \texttt{get\_clusters()} function. This will output a data frame
with two columns, the first will correspond to the \texttt{point}, as
identified by the row name of the original input data frame, \texttt{d},
the second will identify the \texttt{cluster} that each point belongs
to.

\begin{Schunk}
\begin{Sinput}
get_clusters(contribution_matrix)
\end{Sinput}
\begin{Soutput}
#> # A tibble: 5 x 2
#>   point cluster
#>   <chr>   <dbl>
#> 1 A           1
#> 2 B           1
#> 3 C           2
#> 4 D           2
#> 5 E           3
\end{Soutput}
\end{Schunk}

In this example, three clusters are identified with these five points.
Points \texttt{A} and \texttt{B} fall into cluster 1. Points \texttt{C}
and \texttt{D} into cluster 2, and point \texttt{E} in cluster 3.

The \texttt{get\_depths()} function calculates the \emph{depths} of each
point, outputting a data frame with two columns, \texttt{point}
indicating the point, as identified by the row name of the original data
frame, \texttt{d}, and \texttt{depth} indicating the depth of the point.
The data frame is arranged by depth, with the deepest point listed
first.

\begin{Schunk}
\begin{Sinput}
get_depths(contribution_matrix)
\end{Sinput}
\begin{Soutput}
#> # A tibble: 5 x 2
#>   point depth
#>   <chr> <dbl>
#> 1 C     0.588
#> 2 B     0.575
#> 3 D     0.575
#> 4 A     0.5  
#> 5 E     0.262
\end{Soutput}
\end{Schunk}

In this case, the deepest point is \texttt{C}.

The \texttt{get\_bound()} function will calculate the \emph{bound} of
the contribution matrix.

\begin{Schunk}
\begin{Sinput}
get_bound(contribution_matrix)
\end{Sinput}
\begin{Soutput}
#> [1] 0.14
\end{Soutput}
\end{Schunk}

In this case, the bound is 0.14.

The \texttt{any\_isolated()} function will check whether there are any
isolated points that will inadvertently be dropped by a graph.

\begin{Schunk}
\begin{Sinput}
any_isolated(contribution_matrix)
\end{Sinput}
\begin{Soutput}
#> [1] FALSE
\end{Soutput}
\end{Schunk}

In this case, there are no isolated points.

The \texttt{get\_pald\_cluster\_matrix()} will calculate a matrix of
partitioned local depth clusters. This function contains an argument
\texttt{keep\_all\_edges} that indicates whether all edges should be
kept. The default value is \texttt{FALSE}, indicating that edges that
are less than the expectation will be dropped, allowing a clear
clustering of points. Changing this argument to \texttt{TRUE} results in
the pairwise minimum of the contribution matrix and the transpose of the
contribution matrix.

\begin{Schunk}
\begin{Sinput}
get_pald_cluster_matrix(contribution_matrix, keep_all_edges = TRUE)
\end{Sinput}
\begin{Soutput}
#>        1      2     3     4      5
#> A 0.2875 0.1625 0.000 0.050 0.0000
#> B 0.1625 0.3000 0.050 0.050 0.0000
#> C 0.0000 0.0500 0.300 0.175 0.0500
#> D 0.0500 0.0500 0.175 0.300 0.0000
#> E 0.0000 0.0000 0.050 0.000 0.2125
\end{Soutput}
\end{Schunk}

Leaving the \texttt{keep\_all\_edges} as the default, \texttt{FALSE}
will take this pairwise minimum seen above and drop the edges that are
less than the expectation.

\begin{Schunk}
\begin{Sinput}
get_pald_cluster_matrix(contribution_matrix)
\end{Sinput}
\begin{Soutput}
#>        1      2     3     4 5
#> A 0.0000 0.1625 0.000 0.000 0
#> B 0.1625 0.0000 0.000 0.000 0
#> C 0.0000 0.0000 0.000 0.175 0
#> D 0.0000 0.0000 0.175 0.000 0
#> E 0.0000 0.0000 0.000 0.000 0
\end{Soutput}
\end{Schunk}

Finally, the \texttt{get\_pald\_graph()} function takes the contribution
matrix and creates an \CRANpkg{igraph} object, a graph that describes
the relationship between the points.

\begin{Schunk}
\begin{Sinput}
get_pald_graph(contribution_matrix)
\end{Sinput}
\begin{Soutput}
#> IGRAPH 2c0b54a UNW- 5 2 -- 
#> + attr: name (v/c), weight (e/n)
#> + edges from 2c0b54a (vertex names):
#> [1] A--B D--C
\end{Soutput}
\end{Schunk}

Here we see that there are two connected components, points \texttt{A}
and \texttt{B}, which form the first cluster, and points \texttt{D} and
\texttt{C} which form the second.

\hypertarget{plotting-functions}{%
\subsection{Plotting functions}\label{plotting-functions}}

The final category of function is functions for data visualization. We
can begin by visualizing the points in data frame \texttt{d} (Figure
\ref{fig:fig1}).

\begin{Schunk}
\begin{Sinput}
library(ggplot2)
ggplot(d, aes(x1, x2)) +
  geom_text(label = rownames(d))
\end{Sinput}
\begin{figure}
\includegraphics{manuscript_files/figure-latex/fig1-1} \caption[Visualize the points from data frame `d`]{Visualize the points from data frame `d`}\label{fig:fig1}
\end{figure}
\end{Schunk}

We can then pass the contribution matrix to the
\texttt{plot\_pald\_graph()} function to view the relationship between
points (Figure \ref{fig:fig2}). By default, this will invoke the
\texttt{layout\_nicely()} function from \CRANpkg{igraph} to determine
the layout of the graph.

\begin{Schunk}
\begin{Sinput}
d %>%
  get_distance_matrix() %>%
  get_contribution_matrix() %>%
  plot_pald_graph()
\end{Sinput}
\begin{figure}
\includegraphics{manuscript_files/figure-latex/fig2-1} \caption[PaLD graph displaying the relationship between the points in data frame `d`]{PaLD graph displaying the relationship between the points in data frame `d`}\label{fig:fig2}
\end{figure}
\end{Schunk}

The \texttt{layout} argument allows the user to pass a matrix to dictate
the layout of the graph. For example, if we wanted the graph to match
the visualization displayed in Figure \ref{fig:fig1}, we can pass
\texttt{as.matrix(d)}, or a matrix of the data frame \texttt{d} to the
\texttt{layout} argument (Figure \ref{fig:fig3}.

\begin{Schunk}
\begin{Sinput}
d %>%
  get_distance_matrix() %>%
  get_contribution_matrix() %>%
  plot_pald_graph(layout = as.matrix(d))
\end{Sinput}
\begin{figure}
\includegraphics{manuscript_files/figure-latex/fig3-1} \caption[PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1]{PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1}\label{fig:fig3}
\end{figure}
\end{Schunk}

This \texttt{plot\_pald\_graph()} function will also permit parameters
that can be passed to the \texttt{plot.igraph()} function. For example,
to add axes to the graph, the user can pass the \texttt{axes\ =\ TRUE}
argument to the \texttt{...} in the \texttt{plot\_pald\_graph()}
function (Figure \ref{fig:fig4}).

\begin{Schunk}
\begin{Sinput}
d %>%
  get_distance_matrix() %>%
  get_contribution_matrix() %>%
  plot_pald_graph(layout = as.matrix(d),
                  axes = TRUE)
\end{Sinput}
\begin{figure}
\includegraphics{manuscript_files/figure-latex/fig4-1} \caption[PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1, adding axes]{PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1, adding axes}\label{fig:fig4}
\end{figure}
\end{Schunk}

\hypertarget{examples}{%
\subsection{Examples}\label{examples}}

We will demonstrate the utility of the \CRANpkg{pald} package in two
clustering examples.

\hypertarget{clustering-tissue-gene-expression-data}{%
\subsubsection{Clustering tissue gene expression
data}\label{clustering-tissue-gene-expression-data}}

The first example data frame is a subset of tissue gene expression data
from \citet{zilliox2007gene}, \citet{mccall2011gene}, and
\citet{mccall2014gene}, obtained from the \textbf{tissuesGeneExpression}
bioconductor package \citep{tissue}. This data set is included in the
\CRANpkg{pald} package in an object called \texttt{tissue}.

The \texttt{tissue} object is a matrix with 189 rows, each with a
corresponding tissue, such as \texttt{colon}, \texttt{kidney} or
\texttt{cerebellum}. There are 22,215 columns corresponding to gene
expression data from each of these rows.

We will first calculate the contribution matrix for this data frame.

\begin{Schunk}
\begin{Sinput}
tissue_contribution_matrix <- tissue %>%
  get_distance_matrix(scale = FALSE) %>%
  get_contribution_matrix()
\end{Sinput}
\end{Schunk}

We can then use this contribution matrix to display the relationship
between tissue samples using the \texttt{plot\_pald\_graph()} function
(Figure \ref{fig:fig5}).

\begin{Schunk}
\begin{Sinput}
plot_pald_graph(tissue_contribution_matrix)
\end{Sinput}
\begin{figure}
\includegraphics{manuscript_files/figure-latex/fig5-1} \caption[PaLD clustering of tissue data]{PaLD clustering of tissue data}\label{fig:fig5}
\end{figure}
\end{Schunk}

The \texttt{get\_clusters()} function can be used to identify the
clusters of each tissue sample. Since the output is a data frame, we can
summarize the clusters using commonly used data analysis techniques. For
demonstration purposes, we will use the \CRANpkg{dplyr} package to
summarize the contribution of clusters.

\begin{Schunk}
\begin{Sinput}
library(dplyr)
get_clusters(tissue_contribution_matrix) %>%
  group_by(cluster, point) %>%
  count()
\end{Sinput}
\begin{Soutput}
#> # A tibble: 19 x 3
#> # Groups:   cluster, point [19]
#>    cluster point           n
#>      <dbl> <chr>       <int>
#>  1       1 endometrium    15
#>  2       1 kidney         39
#>  3       2 hippocampus    31
#>  4       3 cerebellum     26
#>  5       4 cerebellum      1
#>  6       5 colon          33
#>  7       6 colon           1
#>  8       7 liver           7
#>  9       8 cerebellum      1
#> 10       9 liver          17
#> 11      10 cerebellum      2
#> 12      11 liver           2
#> 13      12 cerebellum      1
#> 14      13 cerebellum      4
#> 15      14 cerebellum      2
#> 16      15 cerebellum      1
#> 17      16 placenta        2
#> 18      17 placenta        1
#> 19      18 placenta        3
\end{Soutput}
\end{Schunk}

From this, we can glean that cluster one consists of two types of
tissue, the kidney and endometrium. Cluster two is comprised of only the
hippocampus.

\hypertarget{clustering-generated-data}{%
\subsubsection{Clustering generated
data}\label{clustering-generated-data}}

The \CRANpkg{pald} includes two data frames generated from the
scikit-learn Python package \citep{pedregosa2011scikit},
\texttt{noisy\_moons} and \texttt{noisy\_circles}.

The \texttt{noisy\_moons} data frame consists of 500 rows and two
columns.

\begin{Schunk}
\begin{Sinput}
moons_contribution_matrix <- noisy_moons %>%
  get_distance_matrix() %>%
  get_contribution_matrix()
\end{Sinput}
\end{Schunk}

When plotting the \texttt{noisy\_moons} PaLD graph, we want the layout
to match the layout of the original data, so we will pass
\texttt{as.matrix(noisy\_moons)} to the \texttt{layout} parameter in the
\texttt{plot\_pald\_graph()} function. Additionally, here the row names
are meaningless, they just correspond to the location of the generated
data, so we can remove the labels on the plot by passing
\texttt{vertex.label\ =\ NA} to the \texttt{...} of the
\texttt{plot\_pald\_graph()} function.

\begin{Schunk}
\begin{Sinput}
plot_pald_graph(moons_contribution_matrix,
                layout = as.matrix(noisy_moons),
                vertex.label = NA)
\end{Sinput}
\end{Schunk}

The \texttt{noisy\_circles} data frame consists of 500 rows and two
columns. We can create the contribution matrix for this data frame using
the same methods as used for the \texttt{noisy\_moons} data.

\begin{Schunk}
\begin{Sinput}
circles_contribution_matrix <- noisy_circles %>%
  get_distance_matrix() %>%
  get_contribution_matrix()
\end{Sinput}
\end{Schunk}

Similarly, we will pass the layout and remove vertex labels for the
noisy circles PaLD plot.

\begin{Schunk}
\begin{Sinput}
plot_pald_graph(circles_contribution_matrix,
                layout = as.matrix(noisy_circles),
                vertex.label = NA)
\end{Sinput}
\end{Schunk}

Because these are \CRANpkg{igraph} objects, they can be combined as you
would normally combine a plot. For example, we could add the line
\texttt{par(mfrow\ =\ c(1,\ 2))} above the plot calls to create an
output where both plots are displayed side by side (Figure
\ref{fig:fig6}).

\begin{Schunk}
\begin{Sinput}
par(mfrow = c(1, 2))
plot_pald_graph(moons_contribution_matrix,
                layout = as.matrix(noisy_moons),
                vertex.label = NA)
plot_pald_graph(circles_contribution_matrix,
                layout = as.matrix(noisy_circles),
                vertex.label = NA)
\end{Sinput}
\begin{figure}
\includegraphics{manuscript_files/figure-latex/fig6-1} \caption[PaLD clustering of noisy moons (left) and noisy circles (right) data]{PaLD clustering of noisy moons (left) and noisy circles (right) data}\label{fig:fig6}
\end{figure}
\end{Schunk}

The ability of the PaLD algorithm to discern clusters is demonstrated
here.

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

This paper introduces the \CRANpkg{pald} package, demonstrating it's
utility for providing a parameter-free clustering algorithm that can
easily be applied to any data set.

\bibliography{RJreferences}

\address{%
Lucy D'Agostino McGowan\\
Wake Forest University\\
Winston-Salem, NC\\ 27106\\
}
\href{mailto:mcgowald@wfu.edu}{\nolinkurl{mcgowald@wfu.edu}}

\address{%
Katherine Moore\\
Wake Forest Unversity\\
Winston-Salem, NC\\ 27106\\
}
\href{mailto:mooreke@wfu.edu}{\nolinkurl{mooreke@wfu.edu}}

\address{%
Kenneth Berenhaut\\
Wake Forest University\\
Winston-Salem, NC\\ 27106\\
}
\href{mailto:berenhks@wfu.edu}{\nolinkurl{berenhks@wfu.edu}}

### 2022-03-13 r-journal/manuscript.Rmd
update paper

@@
 
-* Describe PaLD (Ken & Kate?)
-* Cite PaLD
+* Describe PaLD (Ken & Kate?) [@berenhaut2022social]
 
@@
 
-1. Helper functions to organize data into the correct format, into distance matrices and then contribution matrices
-2. Functions that convert a contribution matrix into a variety of useful formats, including partitioned local depths, clusters, and graphs
+1. Helper functions to organize data into the correct format, into distance matrices and then cohesion matrices
+2. Functions that convert a cohesion matrix into a variety of useful formats, including partitioned local depths, clusters, and graphs
 3. Plotting functions
@@
 
-While it is not a necessity, the \CRANpkg{pald} package is designed to function well with the pipe operator, `%>%`, from the \citet{magrittr} \CRANpkg{magrittr} package in that the first argument of each function is the data.
-This functionality will be demonstrated below.
+While it is not a necessity, the \CRANpkg{pald} package is designed to function well with the pipe operator, `|>`. This functionality will be demonstrated below.
 
@@
 For demonstration purposes, below is a sample data frame with two variables, `x1` and `x2`.
-The methods put forth here work on data frames with higher dimensions, as described in the **Examples** section, we are simply choosing a small data frame here for demonstration purposes.
+The methods put forth here work on data frames with higher dimensions, as described in the **Examples** section; we are simply choosing a small data frame here for demonstration purposes.
 
@@
 library(pald)
-d <- data.frame(
+df <- data.frame(
   x1 = c(6, 8, 11, 16, 4),
@@
 )
-rownames(d) <- c("A", "B", "C", "D", "E")
+rownames(df) <- c("A", "B", "C", "D", "E")
 ```
@@
 If the data are already in this form, the user can skip to the next step.
-The `get_distance_matrix()` function converts an input data frame into a distance matrix, as demonstrated below.
+The `dist()` function converts an input data frame into a distance matrix, as demonstrated below.
 
 ```{r}
-get_distance_matrix(d)
+d <- dist(df)
 ```
+
 This will create an $n\times n$ distance matrix, where $n$ corresponds to the number of observations in the original data frame, in this example $n = 5$. 
-By default, the distance matrix is _scaled_, is possible to create a distance matrix that is not scaled by changing the `scale` argument to `FALSE`.
+<!-- TODO: is this nxn? looks like (n-1)x(n-1) -->
+This distance matrix can then be passed to the `cohesion_matrix()` function in order to calculate the pairwise cohesion values.
+Cohesion is an interpretable probability that reflects the strength of alignment of two points.
+Again, if the user begins with a distance matrix, they can skip the first step and simply input the distance matrix into this function.
 
 ```{r}
-get_distance_matrix(d, scale = FALSE)
+d <- dist(df)
+cohesion_matrix(d)
 ```
 
- As mentioned previously, the functions in this package are designed to work with the \CRANpkg{magrittr} pipe operator, `%>%`, so the same code as above could be written utilizing this format.
+Equivalently, the user can use the native pipe `|>` as follows.
 
 ```{r}
-d %>%
-  get_distance_matrix()
+df |>
+  dist() |>
+  cohesion_matrix()
 ```
 
-This distance matrix can then be passed to the `get_contribution_matrix()` function in order to calculate a matrix of contributions.
-Again, if the user begins with a distance matrix, they can skip the first step and simple input the contribution matrix into this function.
+The _cohesion matrix_ output by the `cohesion_matrix()` is the main input for the majority of the remaining functions.
+
+### Functions that convert a cohesion matrix into useful formats
+
+From the _cohesion matrix_, a variety of useful quantities can be calculated. Below, we create a cohesion matrix using the functions described in the previous section.
 
 ```{r}
-distance_matrix <- get_distance_matrix(d)
-get_contribution_matrix(distance_matrix)
+d |>
+  dist() |>
+  cohesion_matrix() -> cohesion
 ```
 
-Again, the \CRANpkg{magrittr} pipe can be used as follows.
+To calculate the _clusters_ that each point will fall into, we can use the `community_clusters()` function.
+This will output a data frame with two columns, the first will correspond to the `point`, as identified by the row name of the original input data frame, `df`, the second will identify the `cluster` that each point belongs to.
 
 ```{r}
-d %>%
-  get_distance_matrix() %>%
-  get_contribution_matrix()
-```
-
-The _contribution matrix_ output by the `get_contribution_matrix()` is the main input for the majority of the remaining functions.
-
-### Functions that convert a contribution matrix into useful formats
-
-From the _contribution matrix_, a variety of useful quantities can be calculated. Below, we create a contribution matrix using the functions described in the previous section.
-
-```{r}
-d %>%
-  get_distance_matrix() %>%
-  get_contribution_matrix() -> contribution_matrix
-```
-
-To calculate the _clusters_ that each point will fall into, we can use the `get_clusters()` function.
-This will output a data frame with two columns, the first will correspond to the `point`, as identified by the row name of the original input data frame, `d`, the second will identify the `cluster` that each point belongs to.
-
-```{r}
-get_clusters(contribution_matrix)
+community_clusters(cohesion)
 ```
@@
 
-The `get_depths()` function calculates the _depths_ of each point, outputting a data frame with two columns, `point` indicating the point, as identified by the row name of the original data frame, `d`, and `depth` indicating the depth of the point.
-The data frame is arranged by depth, with the deepest point listed first.
+The `local_depths()` function calculates the _depths_ of each point, outputting a vector of local depths.
+Local depth is an interpretable probability which reflects aspects of relative position and centrality via distance comparisons (i.e., $d(z, x) < d(z, y)$).
 
 ```{r}
-get_depths(contribution_matrix)
+local_depths(cohesion)
 ```
@@
 
-The `get_bound()` function will calculate the _bound_ of the contribution matrix.
+The `strong_threshold()` function will calculate the cohesion threshold for strong ties.
+This is equal to half the average of the diagonal cohesion matrix.
+This is a threshold that may be used to distinguish between strong and weak ties.
 
 ```{r}
-get_bound(contribution_matrix)
+strong_threshold(cohesion)
 ```
 
-In this case, the bound is 0.14.
+In this case, the threshold is `0.14`.
 
@@
 ```{r}
-any_isolated(contribution_matrix)
+any_isolated(cohesion)
 ```
 
-In this case, there are no isolated points.
+Here, there are no isolated points.
 
-The `get_pald_cluster_matrix()` will calculate a matrix of partitioned local depth clusters.
-This function contains an argument `keep_all_edges` that indicates whether all edges should be kept.
-The default value is `FALSE`, indicating that edges that are less than the expectation will be dropped, allowing a clear clustering of points.
-Changing this argument to `TRUE` results in the pairwise minimum of the contribution matrix and the transpose of the contribution matrix.
+The function `cohesion_strong()` will update the cohesion matrix to set all weak ties to zero (via the `strong_threshold()` function).
+Optionally, the matrix will also be symmetrized, with the default parameter `symmetric = TRUE`.
 
 ```{r}
-get_pald_cluster_matrix(contribution_matrix, keep_all_edges = TRUE)
+cohesion_strong(cohesion)
 ```
 
-Leaving the `keep_all_edges` as the default, `FALSE` will take this pairwise minimum seen above and drop the edges that are less than the expectation.
+Finally, the `community_graphs()` function takes the cohesion matrix and creates \CRANpkg{igraph} objects, graphs that describes the relationship between the points.
+This function will output a list of three objects:
+
+* `G`: the weighted (community) graph whose edge weights are mutual cohesion
+* `G_strong`: the weighted (community) graph consisting of edges for which mutual cohesion is greater than the threshold for strong ties
+* `layout`: the graph layout, using the Fruchterman Reingold (FR) force-directed graph drawing for the graph `G`
 
 ```{r}
-get_pald_cluster_matrix(contribution_matrix)
-```
-
-Finally, the `get_pald_graph()` function takes the contribution matrix and creates an \CRANpkg{igraph} object, a graph that describes the relationship between the points.
-
-```{r}
-get_pald_graph(contribution_matrix)
+graphs <- community_graphs(cohesion)
+graphs[["G_strong"]]
 ```
@@
 
-The final category of function is functions for data visualization. We can begin by visualizing the points in data frame `d` (Figure \ref{fig:fig1}).
+The final category of function is functions for data visualization. We can begin by visualizing the points in data frame `df` (Figure \ref{fig:fig1}).
 
-```{r fig1, fig.cap = "Visualize the points from data frame `d`", message = FALSE, warning = FALSE}
+```{r fig1, fig.cap = "Visualize the points from data frame `df`", message = FALSE, warning = FALSE}
 library(ggplot2)
-ggplot(d, aes(x1, x2)) +
-  geom_text(label = rownames(d))
+ggplot(df, aes(x1, x2)) +
+  geom_text(label = rownames(df))
 ```
 
-We can then pass the contribution matrix to the `plot_pald_graph()` function to view the relationship between points (Figure \ref{fig:fig2}).
-By default, this will invoke the `layout_nicely()` function from \CRANpkg{igraph} to determine the layout of the graph.
+We can then pass the cohesion matrix to the `plot_community_graphs()` function to view the relationship between points (Figure \ref{fig:fig2}).
 
-```{r fig2, fig.cap = "PaLD graph displaying the relationship between the points in data frame `d`"}
-d %>%
-  get_distance_matrix() %>%
-  get_contribution_matrix() %>%
-  plot_pald_graph()
+```{r fig2, fig.cap = "PaLD graph displaying the relationship between the points in data frame `df`"}
+df |>
+  dist() |>
+  cohesion_matrix() |>
+  plot_community_graphs()
 ```
@@
 The `layout` argument allows the user to pass a matrix to dictate the layout of the graph.
-For example, if we wanted the graph to match the visualization displayed in Figure \ref{fig:fig1}, we can pass `as.matrix(d)`, or a matrix of the data frame `d` to the `layout` argument  (Figure \ref{fig:fig3}.
+For example, if we wanted the graph to match the visualization displayed in Figure \ref{fig:fig1}, we can pass `as.matrix(df)`, or a matrix of the data frame `df` to the `layout` argument  (Figure \ref{fig:fig3}.
 ```{r fig3, fig.cap = "PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1"}
-d %>%
-  get_distance_matrix() %>%
-  get_contribution_matrix() %>%
-  plot_pald_graph(layout = as.matrix(d))
+df |>
+  dist() |>
+  cohesion_matrix() |>
+  plot_community_graphs(layout = as.matrix(df))
 ```
 
-This `plot_pald_graph()` function will also permit parameters that can be passed to the `plot.igraph()` function.
-For example, to add axes to the graph, the user can pass the `axes = TRUE` argument to the `...` in the `plot_pald_graph()` function (Figure \ref{fig:fig4}).
+This `plot_community_graphs()` function will also permit parameters that can be passed to the `plot.igraph()` function.
+For example, to add axes to the graph, the user can pass the `axes = TRUE` argument to the `...` in the `plot_community_graphs()` function (Figure \ref{fig:fig4}).
 
 ```{r fig4, fig.cap = "PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1, adding axes"}
-d %>%
-  get_distance_matrix() %>%
-  get_contribution_matrix() %>%
-  plot_pald_graph(layout = as.matrix(d),
+df |>
+  dist() |>
+  cohesion_matrix() |>
+  plot_community_graphs(layout = as.matrix(df),
                         axes = TRUE)
@@
 
-The first example data frame is a subset of tissue gene expression data from \citet{zilliox2007gene}, \citet{mccall2011gene}, and \citet{mccall2014gene}, obtained from the **tissuesGeneExpression** bioconductor package \citep{tissue}.
-This data set is included in the \CRANpkg{pald} package in an object called `tissue`.
+The first example if from a subset of tissue gene expression data from \citet{zilliox2007gene}, \citet{mccall2011gene}, and \citet{mccall2014gene}, obtained from the **tissuesGeneExpression** bioconductor package \citep{tissue}.
+A cohesion matrix was created using this data set and is included the \CRANpkg{pald} package in an object called `tissue_cohesion_matrix`.
 
-The `tissue` object is a matrix with 189 rows, each with a corresponding tissue, such as `colon`, `kidney` or `cerebellum`.
-There are 22,215 columns corresponding to gene expression data from each of these rows.
+The `tissue_cohesion_matrix` object is a cohesion matrix with 189 rows and 189 columns.
 
-We will first calculate the contribution matrix for this data frame.
+We can use this contribution matrix to display the relationship between tissue samples using the `plot_community_graphs()` function (Figure \ref{fig:fig5}).
+For clarity of the display, we show how to only include a random set of the vertex labels.
+We can pass this random set through the `...` to the `plot.igraph` `vertex.label` parameter.
 
-```{r}
-tissue_contribution_matrix <- tissue %>%
-  get_distance_matrix(scale = FALSE) %>%
-  get_contribution_matrix()
+```{r fig5, fig.cap = "PaLD clustering of tissue data", fig.height = 9, fig.width = 7}
+set.seed(1)
+
+labels <- rownames(tissue_cohesion_matrix)
+labels[sample(1:189, 125)] <- ""
+plot_community_graphs(tissue_cohesion_matrix,
+                      vertex.label = labels,
+                      vertex.size = 4)
 ```
 
-We can then use this contribution matrix to display the relationship between tissue samples using the `plot_pald_graph()` function (Figure \ref{fig:fig5}).
-
-```{r fig5, fig.cap = "PaLD clustering of tissue data"}
-plot_pald_graph(tissue_contribution_matrix)
-```
-
-The `get_clusters()` function can be used to identify the clusters of each tissue sample.
+The `community_clusters()` function can be used to identify the clusters of each tissue sample.
 Since the output is a data frame, we can summarize the clusters using commonly used data analysis techniques.
@@
 library(dplyr)
-get_clusters(tissue_contribution_matrix) %>%
+community_clusters(tissue_cohesion_matrix) %>%
   group_by(cluster, point) %>%
@@
 
-The \CRANpkg{pald} includes two data frames generated from the scikit-learn Python package \citep{pedregosa2011scikit}, `noisy_moons` and `noisy_circles`.
+The \CRANpkg{pald} includes three randomly generated data frames corresponding to plots from @berenhaut2022social:
 
-The `noisy_moons` data frame consists of 500 rows and two columns.
+* `exdata1` is a data set consisting of 8 points to recreate Figure 1 in @berenhaut2022social 
+* `exdata2` is a data set consisting of 16 points to recreate Figure 2 in @berenhaut2022social 
+* `exdata3` is a data set consisting of 240 points to recreate Figure 4D in @berenhaut2022social
+
+Here, we will demonstrate how to use `exdata3`. These points were generated from bivariate normal distributions with varying means and variances.
 
 ```{r}
-moons_contribution_matrix <- noisy_moons %>%
-  get_distance_matrix() %>%
-  get_contribution_matrix()
+exdata_cohesion_matrix <- exdata3 |>
+  dist() |>
+  cohesion_matrix()
 ```
 
-When plotting the `noisy_moons` PaLD graph, we want the layout to match the layout of the original data, so we will pass `as.matrix(noisy_moons)` to the `layout` parameter in the `plot_pald_graph()` function.
-Additionally, here the row names are meaningless, they just correspond to the location of the generated data, so we can remove the labels on the plot by passing `vertex.label = NA` to the `...` of the `plot_pald_graph()` function.
+When plotting the `exdata3` PaLD graph, we want the layout to match the layout of the original data, so we will pass `as.matrix(exdata3)` to the `layout` parameter in the `plot_community_graphs()` function.
+Additionally, here the row names are meaningless, they just correspond to the location of the generated data, so we can remove the labels on the plot by setting `show_labels = FALSE`.
+The `only_strong` parameter will only display the strongly connected edges.
+Additionally, we can optionally pass parameters through the `...` to `plot::igraph` function; for example, to increase the vertex size we can set `vertex.size = 5`.
+Figure \ref{fig:fig6} shows this.
 
-```{r, eval = FALSE}
-plot_pald_graph(moons_contribution_matrix,
-                layout = as.matrix(noisy_moons),
-                vertex.label = NA)
-```
-
-The `noisy_circles` data frame consists of 500 rows and two columns. We can create the contribution matrix for this data frame using the same methods as used for the `noisy_moons` data.
-
-```{r}
-circles_contribution_matrix <- noisy_circles %>%
-  get_distance_matrix() %>%
-  get_contribution_matrix()
-```
-
-Similarly, we will pass the layout and remove vertex labels for the noisy circles PaLD plot.
-
-```{r, eval = FALSE}
-plot_pald_graph(circles_contribution_matrix,
-                layout = as.matrix(noisy_circles),
-                vertex.label = NA)
-```
-
-Because these are \CRANpkg{igraph} objects, they can be combined as you would normally combine a plot.
-For example, we could add the line `par(mfrow = c(1, 2))` above the plot calls to create an output where both plots are displayed side by side (Figure \ref{fig:fig6}).
-
-```{r fig6, fig.cap = "PaLD clustering of noisy moons (left) and noisy circles (right) data"}
-par(mfrow = c(1, 2))
-plot_pald_graph(moons_contribution_matrix,
-                layout = as.matrix(noisy_moons),
-                vertex.label = NA)
-plot_pald_graph(circles_contribution_matrix,
-                layout = as.matrix(noisy_circles),
-                vertex.label = NA)
+```{r fig6, fig.cap = "PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al. (2022))"}
+plot_community_graphs(exdata_cohesion_matrix,
+                layout = as.matrix(exdata3),
+                show_labels = FALSE, 
+                only_strong = TRUE,
+                vertex.size = 5)
 ```

### 2022-03-13 r-journal/manuscript.tex
update paper

@@
 \title{Partitioned Local Depth (PaLD) Clustering Analyses in R}
-\author{by Lucy D'Agostino McGowan, Katherine Moore, Kenneth Berenhaut}
+\author{by Lucy D'Agostino McGowan, Katherine Moore, and Kenneth Berenhaut}
 
@@
 
-% Any extra LaTeX you need in the preamble
-
 \hypertarget{introduction}{%
@@
 \item
-  Describe PaLD (Ken \& Kate?)
-\item
-  Cite PaLD
+  Describe PaLD (Ken \& Kate?) \citep{berenhaut2022social}
 \end{itemize}
@@
   Helper functions to organize data into the correct format, into
-  distance matrices and then contribution matrices
+  distance matrices and then cohesion matrices
 \item
-  Functions that convert a contribution matrix into a variety of useful
+  Functions that convert a cohesion matrix into a variety of useful
   formats, including partitioned local depths, clusters, and graphs
@@
 While it is not a necessity, the \CRANpkg{pald} package is designed to
-function well with the pipe operator, \texttt{\%\textgreater{}\%}, from
-the \citet{magrittr} \CRANpkg{magrittr} package in that the first
-argument of each function is the data. This functionality will be
-demonstrated below.
+function well with the pipe operator, \texttt{\textbar{}\textgreater{}}.
+This functionality will be demonstrated below.
 
@@
 on data frames with higher dimensions, as described in the
-\textbf{Examples} section, we are simply choosing a small data frame
+\textbf{Examples} section; we are simply choosing a small data frame
 here for demonstration purposes.
@@
 )
-rownames(d) <- c("A", "B", "C", "D", "E")
+rownames(df) <- c("A", "B", "C", "D", "E")
 \end{Sinput}
@@
 construct a \emph{distance matrix}. If the data are already in this
-form, the user can skip to the next step. The
-\texttt{get\_distance\_matrix()} function converts an input data frame
-into a distance matrix, as demonstrated below.
+form, the user can skip to the next step. The \texttt{dist()} function
+converts an input data frame into a distance matrix, as demonstrated
+below.
 
@@
 \begin{Sinput}
-get_distance_matrix(d)
+d <- dist(df)
 \end{Sinput}
-\begin{Soutput}
-#>           A         B        C        D        E
-#> A 0.0000000 0.4584177 1.717380 2.158418 2.229205
-#> B 0.4584177 0.0000000 1.644296 1.778777 2.505929
-#> C 1.7173802 1.6442957 0.000000 1.468398 1.713327
-#> D 2.1584182 1.7787772 1.468398 0.000000 3.158040
-#> E 2.2292047 2.5059291 1.713327 3.158040 0.000000
-\end{Soutput}
 \end{Schunk}
@@
 corresponds to the number of observations in the original data frame, in
-this example \(n = 5\). By default, the distance matrix is
-\emph{scaled}, is possible to create a distance matrix that is not
-scaled by changing the \texttt{scale} argument to \texttt{FALSE}.
+this example \(n = 5\). This distance matrix can then be passed to the
+\texttt{cohesion\_matrix()} function in order to calculate the pairwise
+cohesion values. Cohesion is an interpretable probability that reflects
+the strength of alignment of two points. Again, if the user begins with
+a distance matrix, they can skip the first step and simply input the
+distance matrix into this function.
 
@@
 \begin{Sinput}
-get_distance_matrix(d, scale = FALSE)
+d <- dist(df)
+cohesion_matrix(d)
 \end{Sinput}
@@
 #>        A      B         C         D         E
-#> A  0.000000  2.236068 9.433981 10.198039 13.152946
-#> B  2.236068  0.000000 9.486833  8.544004 14.560220
-#> C  9.433981  9.486833 0.000000  7.810250  8.602325
-#> D 10.198039  8.544004 7.810250  0.000000 16.278821
-#> E 13.152946 14.560220 8.602325 16.278821  0.000000
+#> A 0.2875 0.1625 0.0000000 0.0500000 0.0000000
+#> B 0.1625 0.2875 0.0000000 0.0500000 0.0000000
+#> C 0.0000 0.0000 0.3083333 0.1833333 0.1000000
+#> D 0.0500 0.0500 0.1750000 0.3000000 0.0000000
+#> E 0.0000 0.0000 0.1000000 0.0000000 0.2333333
 \end{Soutput}
@@
 
-As mentioned previously, the functions in this package are designed to
-work with the \CRANpkg{magrittr} pipe operator,
-\texttt{\%\textgreater{}\%}, so the same code as above could be written
-utilizing this format.
+Equivalently, the user can use the native pipe
+\texttt{\textbar{}\textgreater{}} as follows.
 
@@
 \begin{Sinput}
-d %>%
-  get_distance_matrix()
+df |>
+  dist() |>
+  cohesion_matrix()
 \end{Sinput}
@@
 #>        A      B         C         D         E
-#> A 0.0000000 0.4584177 1.717380 2.158418 2.229205
-#> B 0.4584177 0.0000000 1.644296 1.778777 2.505929
-#> C 1.7173802 1.6442957 0.000000 1.468398 1.713327
-#> D 2.1584182 1.7787772 1.468398 0.000000 3.158040
-#> E 2.2292047 2.5059291 1.713327 3.158040 0.000000
+#> A 0.2875 0.1625 0.0000000 0.0500000 0.0000000
+#> B 0.1625 0.2875 0.0000000 0.0500000 0.0000000
+#> C 0.0000 0.0000 0.3083333 0.1833333 0.1000000
+#> D 0.0500 0.0500 0.1750000 0.3000000 0.0000000
+#> E 0.0000 0.0000 0.1000000 0.0000000 0.2333333
 \end{Soutput}
@@
 
-This distance matrix can then be passed to the
-\texttt{get\_contribution\_matrix()} function in order to calculate a
-matrix of contributions. Again, if the user begins with a distance
-matrix, they can skip the first step and simple input the contribution
-matrix into this function.
+The \emph{cohesion matrix} output by the \texttt{cohesion\_matrix()} is
+the main input for the majority of the remaining functions.
+
+\hypertarget{functions-that-convert-a-cohesion-matrix-into-useful-formats}{%
+\subsubsection{Functions that convert a cohesion matrix into useful
+formats}\label{functions-that-convert-a-cohesion-matrix-into-useful-formats}}
+
+From the \emph{cohesion matrix}, a variety of useful quantities can be
+calculated. Below, we create a cohesion matrix using the functions
+described in the previous section.
 
@@
 \begin{Sinput}
-distance_matrix <- get_distance_matrix(d)
-get_contribution_matrix(distance_matrix)
-\end{Sinput}
-\begin{Soutput}
-#>        1      2     3     4      5
-#> A 0.2875 0.1625 0.000 0.050 0.0000
-#> B 0.1750 0.3000 0.050 0.050 0.0000
-#> C 0.0000 0.0625 0.300 0.175 0.0500
-#> D 0.0500 0.0500 0.175 0.300 0.0000
-#> E 0.0000 0.0000 0.050 0.000 0.2125
-\end{Soutput}
-\end{Schunk}
-
-Again, the \CRANpkg{magrittr} pipe can be used as follows.
-
-\begin{Schunk}
-\begin{Sinput}
-d %>%
-  get_distance_matrix() %>%
-  get_contribution_matrix()
-\end{Sinput}
-\begin{Soutput}
-#>        1      2     3     4      5
-#> A 0.2875 0.1625 0.000 0.050 0.0000
-#> B 0.1750 0.3000 0.050 0.050 0.0000
-#> C 0.0000 0.0625 0.300 0.175 0.0500
-#> D 0.0500 0.0500 0.175 0.300 0.0000
-#> E 0.0000 0.0000 0.050 0.000 0.2125
-\end{Soutput}
-\end{Schunk}
-
-The \emph{contribution matrix} output by the
-\texttt{get\_contribution\_matrix()} is the main input for the majority
-of the remaining functions.
-
-\hypertarget{functions-that-convert-a-contribution-matrix-into-useful-formats}{%
-\subsubsection{Functions that convert a contribution matrix into useful
-formats}\label{functions-that-convert-a-contribution-matrix-into-useful-formats}}
-
-From the \emph{contribution matrix}, a variety of useful quantities can
-be calculated. Below, we create a contribution matrix using the
-functions described in the previous section.
-
-\begin{Schunk}
-\begin{Sinput}
-d %>%
-  get_distance_matrix() %>%
-  get_contribution_matrix() -> contribution_matrix
+d |>
+  dist() |>
+  cohesion_matrix() -> cohesion
 \end{Sinput}
@@
 To calculate the \emph{clusters} that each point will fall into, we can
-use the \texttt{get\_clusters()} function. This will output a data frame
-with two columns, the first will correspond to the \texttt{point}, as
-identified by the row name of the original input data frame, \texttt{d},
-the second will identify the \texttt{cluster} that each point belongs
-to.
+use the \texttt{community\_clusters()} function. This will output a data
+frame with two columns, the first will correspond to the \texttt{point},
+as identified by the row name of the original input data frame,
+\texttt{df}, the second will identify the \texttt{cluster} that each
+point belongs to.
 
@@
 \begin{Sinput}
-get_clusters(contribution_matrix)
+community_clusters(cohesion)
 \end{Sinput}
 \begin{Soutput}
-#> # A tibble: 5 x 2
 #>   point cluster
-#>   <chr>   <dbl>
-#> 1 A           1
-#> 2 B           1
-#> 3 C           2
-#> 4 D           2
-#> 5 E           3
+#> A     A       1
+#> B     B       1
+#> C     C       2
+#> D     D       2
+#> E     E       3
 \end{Soutput}
@@
 
-The \texttt{get\_depths()} function calculates the \emph{depths} of each
-point, outputting a data frame with two columns, \texttt{point}
-indicating the point, as identified by the row name of the original data
-frame, \texttt{d}, and \texttt{depth} indicating the depth of the point.
-The data frame is arranged by depth, with the deepest point listed
-first.
+The \texttt{local\_depths()} function calculates the \emph{depths} of
+each point, outputting a vector of local depths. Local depth is an
+interpretable probability which reflects aspects of relative position
+and centrality via distance comparisons (i.e., \(d(z, x) < d(z, y)\)).
 
@@
 \begin{Sinput}
-get_depths(contribution_matrix)
+local_depths(cohesion)
 \end{Sinput}
 \begin{Soutput}
-#> # A tibble: 5 x 2
-#>   point depth
-#>   <chr> <dbl>
-#> 1 C     0.588
-#> 2 B     0.575
-#> 3 D     0.575
-#> 4 A     0.5  
-#> 5 E     0.262
+#>         A         B         C         D         E 
+#> 0.4062500 0.5416667 0.3833333 0.4687500 0.7000000
 \end{Soutput}
@@
 
-The \texttt{get\_bound()} function will calculate the \emph{bound} of
-the contribution matrix.
+The \texttt{strong\_threshold()} function will calculate the cohesion
+threshold for strong ties. This is equal to half the average of the
+diagonal cohesion matrix. This is a threshold that may be used to
+distinguish between strong and weak ties.
 
@@
 \begin{Sinput}
-get_bound(contribution_matrix)
+strong_threshold(cohesion)
 \end{Sinput}
@@
 
-In this case, the bound is 0.14.
+In this case, the threshold is \texttt{0.14}.
 
@@
 \begin{Sinput}
-any_isolated(contribution_matrix)
+any_isolated(cohesion)
+\end{Sinput}
+\end{Schunk}
+
+Here, there are no isolated points.
+
+The function \texttt{cohesion\_strong()} will update the cohesion matrix
+to set all weak ties to zero (via the \texttt{strong\_threshold()}
+function). Optionally, the matrix will also be symmetrized, with the
+default parameter \texttt{symmetric\ =\ TRUE}.
+
+\begin{Schunk}
+\begin{Sinput}
+cohesion_strong(cohesion)
 \end{Sinput}
 \begin{Soutput}
-#> [1] FALSE
+#>        A      B     C     D      E
+#> A 0.3000 0.1625 0.000 0.000 0.0000
+#> B 0.1625 0.2875 0.000 0.000 0.0000
+#> C 0.0000 0.0000 0.300 0.175 0.0000
+#> D 0.0000 0.0000 0.175 0.300 0.0000
+#> E 0.0000 0.0000 0.000 0.000 0.2125
 \end{Soutput}
@@
 
-In this case, there are no isolated points.
+Finally, the \texttt{community\_graphs()} function takes the cohesion
+matrix and creates \CRANpkg{igraph} objects, graphs that describes the
+relationship between the points. This function will output a list of
+three objects:
 
-The \texttt{get\_pald\_cluster\_matrix()} will calculate a matrix of
-partitioned local depth clusters. This function contains an argument
-\texttt{keep\_all\_edges} that indicates whether all edges should be
-kept. The default value is \texttt{FALSE}, indicating that edges that
-are less than the expectation will be dropped, allowing a clear
-clustering of points. Changing this argument to \texttt{TRUE} results in
-the pairwise minimum of the contribution matrix and the transpose of the
-contribution matrix.
+\begin{itemize}
+\tightlist
+\item
+  \texttt{G}: the weighted (community) graph whose edge weights are
+  mutual cohesion
+\item
+  \texttt{G\_strong}: the weighted (community) graph consisting of edges
+  for which mutual cohesion is greater than the threshold for strong
+  ties
+\item
+  \texttt{layout}: the graph layout, using the Fruchterman Reingold (FR)
+  force-directed graph drawing for the graph \texttt{G}
+\end{itemize}
 
@@
 \begin{Sinput}
-get_pald_cluster_matrix(contribution_matrix, keep_all_edges = TRUE)
+graphs <- community_graphs(cohesion)
+graphs[["G_strong"]]
 \end{Sinput}
 \begin{Soutput}
-#>        1      2     3     4      5
-#> A 0.2875 0.1625 0.000 0.050 0.0000
-#> B 0.1625 0.3000 0.050 0.050 0.0000
-#> C 0.0000 0.0500 0.300 0.175 0.0500
-#> D 0.0500 0.0500 0.175 0.300 0.0000
-#> E 0.0000 0.0000 0.050 0.000 0.2125
-\end{Soutput}
-\end{Schunk}
-
-Leaving the \texttt{keep\_all\_edges} as the default, \texttt{FALSE}
-will take this pairwise minimum seen above and drop the edges that are
-less than the expectation.
-
-\begin{Schunk}
-\begin{Sinput}
-get_pald_cluster_matrix(contribution_matrix)
-\end{Sinput}
-\begin{Soutput}
-#>        1      2     3     4 5
-#> A 0.0000 0.1625 0.000 0.000 0
-#> B 0.1625 0.0000 0.000 0.000 0
-#> C 0.0000 0.0000 0.000 0.175 0
-#> D 0.0000 0.0000 0.175 0.000 0
-#> E 0.0000 0.0000 0.000 0.000 0
-\end{Soutput}
-\end{Schunk}
-
-Finally, the \texttt{get\_pald\_graph()} function takes the contribution
-matrix and creates an \CRANpkg{igraph} object, a graph that describes
-the relationship between the points.
-
-\begin{Schunk}
-\begin{Sinput}
-get_pald_graph(contribution_matrix)
-\end{Sinput}
-\begin{Soutput}
-#> IGRAPH 2c0b54a UNW- 5 2 -- 
+#> IGRAPH 261f39e UNW- 5 2 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from 2c0b54a (vertex names):
-#> [1] A--B D--C
+#> + edges from 261f39e (vertex names):
+#> [1] A--B C--D
 \end{Soutput}
@@
 The final category of function is functions for data visualization. We
-can begin by visualizing the points in data frame \texttt{d} (Figure
+can begin by visualizing the points in data frame \texttt{df} (Figure
 \ref{fig:fig1}).
@@
 library(ggplot2)
-ggplot(d, aes(x1, x2)) +
-  geom_text(label = rownames(d))
+ggplot(df, aes(x1, x2)) +
+  geom_text(label = rownames(df))
 \end{Sinput}
 \begin{figure}
-\includegraphics{manuscript_files/figure-latex/fig1-1} \caption[Visualize the points from data frame `d`]{Visualize the points from data frame `d`}\label{fig:fig1}
+\includegraphics{manuscript_files/figure-latex/fig1-1} \caption[Visualize the points from data frame `df`]{Visualize the points from data frame `df`}\label{fig:fig1}
 \end{figure}
@@
 
-We can then pass the contribution matrix to the
-\texttt{plot\_pald\_graph()} function to view the relationship between
-points (Figure \ref{fig:fig2}). By default, this will invoke the
-\texttt{layout\_nicely()} function from \CRANpkg{igraph} to determine
-the layout of the graph.
+We can then pass the cohesion matrix to the
+\texttt{plot\_community\_graphs()} function to view the relationship
+between points (Figure \ref{fig:fig2}).
 
@@
 \begin{Sinput}
-d %>%
-  get_distance_matrix() %>%
-  get_contribution_matrix() %>%
-  plot_pald_graph()
+df |>
+  dist() |>
+  cohesion_matrix() |>
+  plot_community_graphs()
 \end{Sinput}
 \begin{figure}
-\includegraphics{manuscript_files/figure-latex/fig2-1} \caption[PaLD graph displaying the relationship between the points in data frame `d`]{PaLD graph displaying the relationship between the points in data frame `d`}\label{fig:fig2}
+\includegraphics{manuscript_files/figure-latex/fig2-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`]{PaLD graph displaying the relationship between the points in data frame `df`}\label{fig:fig2}
 \end{figure}
@@
 the visualization displayed in Figure \ref{fig:fig1}, we can pass
-\texttt{as.matrix(d)}, or a matrix of the data frame \texttt{d} to the
+\texttt{as.matrix(df)}, or a matrix of the data frame \texttt{df} to the
 \texttt{layout} argument (Figure \ref{fig:fig3}.
@@
 \begin{Sinput}
-d %>%
-  get_distance_matrix() %>%
-  get_contribution_matrix() %>%
-  plot_pald_graph(layout = as.matrix(d))
+df |>
+  dist() |>
+  cohesion_matrix() |>
+  plot_community_graphs(layout = as.matrix(df))
 \end{Sinput}
@@
 
-This \texttt{plot\_pald\_graph()} function will also permit parameters
-that can be passed to the \texttt{plot.igraph()} function. For example,
-to add axes to the graph, the user can pass the \texttt{axes\ =\ TRUE}
-argument to the \texttt{...} in the \texttt{plot\_pald\_graph()}
-function (Figure \ref{fig:fig4}).
+This \texttt{plot\_community\_graphs()} function will also permit
+parameters that can be passed to the \texttt{plot.igraph()} function.
+For example, to add axes to the graph, the user can pass the
+\texttt{axes\ =\ TRUE} argument to the \texttt{...} in the
+\texttt{plot\_community\_graphs()} function (Figure \ref{fig:fig4}).
 
@@
 \begin{Sinput}
-d %>%
-  get_distance_matrix() %>%
-  get_contribution_matrix() %>%
-  plot_pald_graph(layout = as.matrix(d),
+df |>
+  dist() |>
+  cohesion_matrix() |>
+  plot_community_graphs(layout = as.matrix(df),
                         axes = TRUE)
@@
 
-The first example data frame is a subset of tissue gene expression data
-from \citet{zilliox2007gene}, \citet{mccall2011gene}, and
+The first example if from a subset of tissue gene expression data from
+\citet{zilliox2007gene}, \citet{mccall2011gene}, and
 \citet{mccall2014gene}, obtained from the \textbf{tissuesGeneExpression}
-bioconductor package \citep{tissue}. This data set is included in the
-\CRANpkg{pald} package in an object called \texttt{tissue}.
+bioconductor package \citep{tissue}. A cohesion matrix was created using
+this data set and is included the \CRANpkg{pald} package in an object
+called \texttt{tissue\_cohesion\_matrix}.
 
-The \texttt{tissue} object is a matrix with 189 rows, each with a
-corresponding tissue, such as \texttt{colon}, \texttt{kidney} or
-\texttt{cerebellum}. There are 22,215 columns corresponding to gene
-expression data from each of these rows.
+The \texttt{tissue\_cohesion\_matrix} object is a cohesion matrix with
+189 rows and 189 columns.
 
-We will first calculate the contribution matrix for this data frame.
+We can use this contribution matrix to display the relationship between
+tissue samples using the \texttt{plot\_community\_graphs()} function
+(Figure \ref{fig:fig5}). For clarity of the display, we show how to only
+include a random set of the vertex labels. We can pass this random set
+through the \texttt{...} to the \texttt{plot.igraph}
+\texttt{vertex.label} parameter.
 
@@
 \begin{Sinput}
-tissue_contribution_matrix <- tissue %>%
-  get_distance_matrix(scale = FALSE) %>%
-  get_contribution_matrix()
-\end{Sinput}
-\end{Schunk}
+set.seed(1)
 
-We can then use this contribution matrix to display the relationship
-between tissue samples using the \texttt{plot\_pald\_graph()} function
-(Figure \ref{fig:fig5}).
-
-\begin{Schunk}
-\begin{Sinput}
-plot_pald_graph(tissue_contribution_matrix)
+labels <- rownames(tissue_cohesion_matrix)
+labels[sample(1:189, 125)] <- ""
+plot_community_graphs(tissue_cohesion_matrix,
+                      vertex.label = labels,
+                      vertex.size = 4)
 \end{Sinput}
@@
 
-The \texttt{get\_clusters()} function can be used to identify the
+The \texttt{community\_clusters()} function can be used to identify the
 clusters of each tissue sample. Since the output is a data frame, we can
@@
 
-The \CRANpkg{pald} includes two data frames generated from the
-scikit-learn Python package \citep{pedregosa2011scikit},
-\texttt{noisy\_moons} and \texttt{noisy\_circles}.
+The \CRANpkg{pald} includes three randomly generated data frames
+corresponding to plots from \citet{berenhaut2022social}:
 
-The \texttt{noisy\_moons} data frame consists of 500 rows and two
-columns.
+\begin{itemize}
+\tightlist
+\item
+  \texttt{exdata1} is a data set consisting of 8 points to recreate
+  Figure 1 in \citet{berenhaut2022social}
+\item
+  \texttt{exdata2} is a data set consisting of 16 points to recreate
+  Figure 2 in \citet{berenhaut2022social}
+\item
+  \texttt{exdata3} is a data set consisting of 240 points to recreate
+  Figure 4D in \citet{berenhaut2022social}
+\end{itemize}
+
+Here, we will demonstrate how to use \texttt{exdata3}. These points were
+generated from bivariate normal distributions with varying means and
+variances.
 
@@
 \begin{Sinput}
-moons_contribution_matrix <- noisy_moons %>%
-  get_distance_matrix() %>%
-  get_contribution_matrix()
+exdata_cohesion_matrix <- exdata3 |>
+  dist() |>
+  cohesion_matrix()
 \end{Sinput}
@@
 
-When plotting the \texttt{noisy\_moons} PaLD graph, we want the layout
-to match the layout of the original data, so we will pass
-\texttt{as.matrix(noisy\_moons)} to the \texttt{layout} parameter in the
-\texttt{plot\_pald\_graph()} function. Additionally, here the row names
-are meaningless, they just correspond to the location of the generated
-data, so we can remove the labels on the plot by passing
-\texttt{vertex.label\ =\ NA} to the \texttt{...} of the
-\texttt{plot\_pald\_graph()} function.
+When plotting the \texttt{exdata3} PaLD graph, we want the layout to
+match the layout of the original data, so we will pass
+\texttt{as.matrix(exdata3)} to the \texttt{layout} parameter in the
+\texttt{plot\_community\_graphs()} function. Additionally, here the row
+names are meaningless, they just correspond to the location of the
+generated data, so we can remove the labels on the plot by setting
+\texttt{show\_labels\ =\ FALSE}. The \texttt{only\_strong} parameter
+will only display the strongly connected edges. Additionally, we can
+optionally pass parameters through the \texttt{...} to
+\texttt{plot::igraph} function; for example, to increase the vertex size
+we can set \texttt{vertex.size\ =\ 5}. Figure \ref{fig:fig6} shows this.
 
@@
 \begin{Sinput}
-plot_pald_graph(moons_contribution_matrix,
-                layout = as.matrix(noisy_moons),
-                vertex.label = NA)
-\end{Sinput}
-\end{Schunk}
-
-The \texttt{noisy\_circles} data frame consists of 500 rows and two
-columns. We can create the contribution matrix for this data frame using
-the same methods as used for the \texttt{noisy\_moons} data.
-
-\begin{Schunk}
-\begin{Sinput}
-circles_contribution_matrix <- noisy_circles %>%
-  get_distance_matrix() %>%
-  get_contribution_matrix()
-\end{Sinput}
-\end{Schunk}
-
-Similarly, we will pass the layout and remove vertex labels for the
-noisy circles PaLD plot.
-
-\begin{Schunk}
-\begin{Sinput}
-plot_pald_graph(circles_contribution_matrix,
-                layout = as.matrix(noisy_circles),
-                vertex.label = NA)
-\end{Sinput}
-\end{Schunk}
-
-Because these are \CRANpkg{igraph} objects, they can be combined as you
-would normally combine a plot. For example, we could add the line
-\texttt{par(mfrow\ =\ c(1,\ 2))} above the plot calls to create an
-output where both plots are displayed side by side (Figure
-\ref{fig:fig6}).
-
-\begin{Schunk}
-\begin{Sinput}
-par(mfrow = c(1, 2))
-plot_pald_graph(moons_contribution_matrix,
-                layout = as.matrix(noisy_moons),
-                vertex.label = NA)
-plot_pald_graph(circles_contribution_matrix,
-                layout = as.matrix(noisy_circles),
-                vertex.label = NA)
+plot_community_graphs(exdata_cohesion_matrix,
+                layout = as.matrix(exdata3),
+                show_labels = FALSE, 
+                only_strong = TRUE,
+                vertex.size = 5)
 \end{Sinput}
 \begin{figure}
-\includegraphics{manuscript_files/figure-latex/fig6-1} \caption[PaLD clustering of noisy moons (left) and noisy circles (right) data]{PaLD clustering of noisy moons (left) and noisy circles (right) data}\label{fig:fig6}
+\includegraphics{manuscript_files/figure-latex/fig6-1} \caption[PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al]{PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al.
+(2022))}\label{fig:fig6}
 \end{figure}
@@
 Lucy D'Agostino McGowan\\
-Wake Forest University\\
+Wake Forest University\\%
 Winston-Salem, NC\\ 27106\\
+%
+%
+%
+\href{mailto:mcgowald@wfu.edu}{\nolinkurl{mcgowald@wfu.edu}}%
 }
-\href{mailto:mcgowald@wfu.edu}{\nolinkurl{mcgowald@wfu.edu}}
 
@@
 Katherine Moore\\
-Wake Forest Unversity\\
+Wake Forest Unversity\\%
 Winston-Salem, NC\\ 27106\\
+%
+%
+%
+\href{mailto:mooreke@wfu.edu}{\nolinkurl{mooreke@wfu.edu}}%
 }
-\href{mailto:mooreke@wfu.edu}{\nolinkurl{mooreke@wfu.edu}}
 
@@
 Kenneth Berenhaut\\
-Wake Forest University\\
+Wake Forest University\\%
 Winston-Salem, NC\\ 27106\\
+%
+%
+%
+\href{mailto:berenhks@wfu.edu}{\nolinkurl{berenhks@wfu.edu}}%
 }
-\href{mailto:berenhks@wfu.edu}{\nolinkurl{berenhks@wfu.edu}}

### 2022-03-13 r-journal/manuscript.R

@@ -1,0 +1,120 @@
+# Generated by `rjournal_article()` using `knitr::purl()`: do not edit by hand
+# Please edit manuscript.Rmd to modify this file
+
+## ---- echo = FALSE------------------------------------------------------------
+knitr::opts_chunk$set(fig.height = 4, message = FALSE, warning = FALSE)
+
+
+## -----------------------------------------------------------------------------
+library(pald)
+df <- data.frame(
+  x1 = c(6, 8, 11, 16, 4),
+  x2 = c(5, 4, 13, 7, 18)
+)
+rownames(df) <- c("A", "B", "C", "D", "E")
+
+
+## -----------------------------------------------------------------------------
+d <- dist(df)
+
+
+## -----------------------------------------------------------------------------
+d <- dist(df)
+cohesion_matrix(d)
+
+
+## -----------------------------------------------------------------------------
+df |>
+  dist() |>
+  cohesion_matrix()
+
+
+## -----------------------------------------------------------------------------
+d |>
+  dist() |>
+  cohesion_matrix() -> cohesion
+
+
+## -----------------------------------------------------------------------------
+community_clusters(cohesion)
+
+
+## -----------------------------------------------------------------------------
+local_depths(cohesion)
+
+
+## -----------------------------------------------------------------------------
+strong_threshold(cohesion)
+
+
+## -----------------------------------------------------------------------------
+any_isolated(cohesion)
+
+
+## -----------------------------------------------------------------------------
+cohesion_strong(cohesion)
+
+
+## -----------------------------------------------------------------------------
+graphs <- community_graphs(cohesion)
+graphs[["G_strong"]]
+
+
+## ----fig1, fig.cap = "Visualize the points from data frame `df`", message = FALSE, warning = FALSE----
+library(ggplot2)
+ggplot(df, aes(x1, x2)) +
+  geom_text(label = rownames(df))
+
+
+## ----fig2, fig.cap = "PaLD graph displaying the relationship between the points in data frame `df`"----
+df |>
+  dist() |>
+  cohesion_matrix() |>
+  plot_community_graphs()
+
+
+## ----fig3, fig.cap = "PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1"----
+df |>
+  dist() |>
+  cohesion_matrix() |>
+  plot_community_graphs(layout = as.matrix(df))
+
+
[…]

### 2022-03-17 r-journal/manuscript.Rmd
update paper

@@
 
-In addition, the package provides a number of pertinent example data sets commonly used to demonstrate cluster algorithms, including a synthetic data set of two-dimensional points created by \citet{gionis1clustering} to demonstrate clustering aggregation, a sample of walking distances from a pump in the infamous cholera outbreak \citep{cholera}, clustering data generated from the scikit-learn Python package \citep{pedregosa2011scikit}, and data compiled by \cite{tissue} of tissue gene expressions.
+In addition, the package provides a number of pertinent example data sets commonly used to demonstrate cluster algorithms, including a synthetic data set of two-dimensional points created by \citet{gionis1clustering} to demonstrate clustering aggregation, clustering data generated from the scikit-learn Python package \citep{pedregosa2011scikit}, data describing cognate relationships between words across 87 Indo-European languages \citep{dyen92}, data compiled by \cite{tissue} of tissue gene expressions, and three example data sets generated for the \citet{berenhaut2022social} paper.
 
@@
 
-### Helper functions to create contribution matrix
+### Creating the cohesion matrix
 
@@
 df <- data.frame(
-  x1 = c(6, 8, 11, 16, 4),
-  x2 = c(5, 4, 13, 7, 18)
+  x1 = c(6, 8, 11, 16, 4, 14),
+  x2 = c(5, 4, 13, 7, 6, 10)
 )
-rownames(df) <- c("A", "B", "C", "D", "E")
+rownames(df) <- c("A", "B", "C", "D", "E", "F")
 ```
@@
 
-This will create an $n\times n$ distance matrix, where $n$ corresponds to the number of observations in the original data frame, in this example $n = 5$. 
-<!-- TODO: is this nxn? looks like (n-1)x(n-1) -->
-This distance matrix can then be passed to the `cohesion_matrix()` function in order to calculate the pairwise cohesion values.
+This will create a `dist` object.
+If converted to a matrix, this will be a $n\times n$ distance matrix, where $n$ corresponds to the number of observations in the original data frame, in this example $n = 6$.
+
+This dist object, or equivalently a distance matrix, can then be passed to the `cohesion_matrix()` function in order to calculate the pairwise cohesion values.
 Cohesion is an interpretable probability that reflects the strength of alignment of two points.
@@
 ```{r}
-d |>
+df |>
   dist() |>
@@
 
-In this example, three clusters are identified with these five points. Points `A` and `B` fall into cluster 1. Points `C` and `D` into cluster 2, and point `E` in cluster 3.
+In this example, three clusters are identified with these six points. Points `A`, `B`, and `E` fall into cluster 1. Point `C` is in cluster 2 and points `D` and `F` fall into cluster 3.
 
@@
 
-In this case, the deepest point is `C`.
+In this case, the deepest point is `A`.
 
@@
 
-In this case, the threshold is `0.14`.
+In this case, the threshold is `0.12`.
 
@@
 
-Here we see that there are two connected components, points `A` and `B`, which form the first cluster, and points `D` and `C` which form the second.
+Here we see that there are three connected components, points `A` and `B`, and `A` and `E` which form the first cluster, and points `D` and `F` which form another.
 
@@
 
-The final category of function is functions for data visualization. We can begin by visualizing the points in data frame `df` (Figure \ref{fig:fig1}).
+The final category of function is functions for data visualization.
+We can begin by visualizing the points in data frame `df` (Figure \ref{fig:fig1}).
+When visualizing these points, it is important to have a square plot so as to not distort the distances.
+When using the \CRANpkg{ggplot2} package for this visualization, you can use the `aspect.ratio  = 1` argument in the `theme()` function.
+If using the `plot()` function included in the base library, you can set the plot area to be a square by using `par(pty = "s")` prior to calling the `plot()` function, as shown below.
 
@@
 ggplot(df, aes(x1, x2)) +
-  geom_text(label = rownames(df))
+  geom_text(label = rownames(df)) + 
+  theme(aspect.ratio = 1)
 ```
 
-We can then pass the cohesion matrix to the `plot_community_graphs()` function to view the relationship between points (Figure \ref{fig:fig2}).
+We can pass the cohesion matrix to the `plot_community_graphs()` function to view the relationship between points (Figure \ref{fig:fig2}).
 
 ```{r fig2, fig.cap = "PaLD graph displaying the relationship between the points in data frame `df`"}
-df |>
-  dist() |>
-  cohesion_matrix() |>
-  plot_community_graphs()
+plot_community_graphs(cohesion)
 ```
@@
 For example, if we wanted the graph to match the visualization displayed in Figure \ref{fig:fig1}, we can pass `as.matrix(df)`, or a matrix of the data frame `df` to the `layout` argument  (Figure \ref{fig:fig3}.
+Notice in the code below, we set `par(pty = "s")` in order to ensure that we have a square plotting region.
+Additionally, this `plot_community_graphs()` function will also permit parameters that can be passed to the `plot.igraph()` function.
+For example, we can pass arguments to the `plot.igraph` function via the `...`, for example to increase the vertex size and change the vertex label color, we can specify `vertex.size = 15` and `vertex.label.color = "white"`.
+
 ```{r fig3, fig.cap = "PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1"}
-df |>
-  dist() |>
-  cohesion_matrix() |>
-  plot_community_graphs(layout = as.matrix(df))
-```
+par(pty = "s")
 
-This `plot_community_graphs()` function will also permit parameters that can be passed to the `plot.igraph()` function.
-For example, to add axes to the graph, the user can pass the `axes = TRUE` argument to the `...` in the `plot_community_graphs()` function (Figure \ref{fig:fig4}).
-
-```{r fig4, fig.cap = "PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1, adding axes"}
-df |>
-  dist() |>
-  cohesion_matrix() |>
-  plot_community_graphs(layout = as.matrix(df),
-                        axes = TRUE)
+plot_community_graphs(cohesion, 
+                      layout = as.matrix(df),
+                      vertex.size = 15,
+                      vertex.label.color = "white")
 ```
@@
 
-We will demonstrate the utility of the \CRANpkg{pald} package in two clustering examples.
+We will demonstrate the utility of the \CRANpkg{pald} package in three clustering examples.
 
@@
 The first example if from a subset of tissue gene expression data from \citet{zilliox2007gene}, \citet{mccall2011gene}, and \citet{mccall2014gene}, obtained from the **tissuesGeneExpression** bioconductor package \citep{tissue}.
-A cohesion matrix was created using this data set and is included the \CRANpkg{pald} package in an object called `tissue_cohesion_matrix`.
+A `dist` object was created using this data set and is included the \CRANpkg{pald} package in an object called `tissue_dist`.
 
-The `tissue_cohesion_matrix` object is a cohesion matrix with 189 rows and 189 columns.
+The `tissue_dist` object is a `dist` object resulting in a distance matrix with 189 rows and 189 columns.
 
-We can use this contribution matrix to display the relationship between tissue samples using the `plot_community_graphs()` function (Figure \ref{fig:fig5}).
-For clarity of the display, we show how to only include a random set of the vertex labels.
-We can pass this random set through the `...` to the `plot.igraph` `vertex.label` parameter.
+We can create the cohesion matrix using the `cohesion_matrix` function.
 
-```{r fig5, fig.cap = "PaLD clustering of tissue data", fig.height = 9, fig.width = 7}
-set.seed(1)
-
-labels <- rownames(tissue_cohesion_matrix)
-labels[sample(1:189, 125)] <- ""
-plot_community_graphs(tissue_cohesion_matrix,
-                      vertex.label = labels,
-                      vertex.size = 4)
+```{r}
+tissue_cohesion <- cohesion_matrix(tissue_dist)
 ```
@@
 ```{r}
-library(dplyr)
-community_clusters(tissue_cohesion_matrix) %>%
-  group_by(cluster, point) %>%
-  count()
+community_clusters(tissue_cohesion) |>
+  dplyr::count(cluster, point)
 ```
@@
 From this, we can glean that cluster one consists of two types of tissue, the kidney and endometrium. Cluster two is comprised of only the hippocampus.
+
+We can also display the relationships between tissue samples using the `plot_community_graphs()` function (Figure \ref{fig:fig4}).
+For clarity of the display, we show how to remove the labels using `show_labels = FALSE`.
+We will instead color by the labels by passing these to the `vertex.color` parameter through the  `...` to the `plot.igraph` function.
+Similarly, we can add a legend using the `legend()` function, as you would for a \CRANpkg{igraph} visualization.
+Additionally, we use the `edge_width_factor` and `emph_strong` arguments to adjust the width of the lines between and within PaLD clusters.
+
+```{r, eval = FALSE}
+labels <- rownames(tissue_cohesion)
+plot_community_graphs(tissue_cohesion,
+                      show_labels = FALSE,
+                      vertex.size = 4,
+                      vertex.color = as.factor(labels),
+                      edge_width_factor = 35,
+                      emph_strong = 5) 
+legend("topleft", 
+       legend = unique(as.factor(labels)), 
+       pt.bg = unique(as.factor(labels)),
+       col = "black",
+       pch = 21)
+```
+
+```{r, include = FALSE, output = "hide"}
+set.seed(1)
+tissue_cohesion <- cohesion_matrix(tissue_dist)
+labels <- rownames(tissue_cohesion)
+png("fig5.png", res = 300, height = 3000, width = 3000)
+plot_community_graphs(tissue_cohesion,
+                      show_labels = FALSE,
+                      vertex.size = 4,
+                      vertex.color = as.factor(labels),
+                      edge_width_factor = 35,
+                      emph_strong = 5) 
+legend("topright", 
+       legend = unique(as.factor(labels)), 
+       pt.bg = unique(as.factor(labels)),
+       col = "black",
+       pch = 21)
+dev.off()
+```
+
+```{r fig4, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "PaLD clustering of tissue data.
+The line colors indicate the PaLD clusters, the point colors indicate the tissue classification.", out.width = "100%"}
+knitr::include_graphics("fig5.png")
+```
+
+## Cognate-based Language Families
+
+This example performs a PaLD analysis on a data set from \citet{dyen92} that examines the relationship between 87 Indo-European languages from the perspective of cognates.
+A `dist` object was created from this data set and included in the \CRANpkg{pald} package in an object called `cognate_dist`.
+
+This example will demonstrate how you can apply functions in the \CRANpkg{igraph} package to objects output from the \CRANpkg{pald} package.
+We can first use the `cohesion_matrix()` function to calculate the cohesion matrix and the `community_graphs()` function to create a list with the weighted community graph, the weighted community graph with only strong ties included (and all others set to 0), and the layout.
+From this, we can extract the graph with only the strong ties, here called `cognate_graph_strong`.
+
+```{r}
+cognate_cohesion <- cohesion_matrix(cognate_dist)
+cognate_graphs <- community_graphs(cognate_cohesion)
+
+cognate_graph_strong <- cognate_graphs[["G_strong"]]
+```
+
+We can then use the `neighbors()` function from the \CRANpkg{igraph} package to extract the neighbors in this graph.
+For example, if we wanted to extract all neighbors where the language is "French", we would run the following.
+
+```{r}
+french_neighbors <- igraph::neighbors(cognate_graph_strong, "French")
+french_neighbors
+```
+
+Similarly, we can print the associated neighborhood weights by subsetting the cohesion matrix.
+
+```{r}
+cognate_cohesion["French", french_neighbors]
+```
 
@@
 
-Here, we will demonstrate how to use `exdata3`. These points were generated from bivariate normal distributions with varying means and variances.
+Here, we will demonstrate how to use `exdata3`. These points were generated from bivariate normal distributions with varying means and variances. There are eight "true" clusters.
+
+We will demonstrate how we can compare PaLD to two clustering methods: *k*-means and hierarchical clustering.
+The code below calculates the cohesion matrix (`exdata_cohesion`) as well as the clusters via PaLD (`exdata_pald`), *k*-means (`exdata_kmeans`) and hierarchical clustering (`exdata_hclust`).
 
 ```{r}
-exdata_cohesion_matrix <- exdata3 |>
+exdata_cohesion <- exdata3 |>
   dist() |>
   cohesion_matrix()
+
+exdata_pald <- community_clusters(exdata_cohesion)$cluster
+
+exdata_kmeans <- kmeans(exdata3, 8)$cluster
+
+exdata_hclust <- exdata3 |>
+  dist() |>
+  hclust() |>
+  cutree(k = 8) 
 ```
 
-When plotting the `exdata3` PaLD graph, we want the layout to match the layout of the original data, so we will pass `as.matrix(exdata3)` to the `layout` parameter in the `plot_community_graphs()` function.
-Additionally, here the row names are meaningless, they just correspond to the location of the generated data, so we can remove the labels on the plot by setting `show_labels = FALSE`.
-The `only_strong` parameter will only display the strongly connected edges.
-Additionally, we can optionally pass parameters through the `...` to `plot::igraph` function; for example, to increase the vertex size we can set `vertex.size = 5`.
-Figure \ref{fig:fig6} shows this.
+We can compare this to the clustering generated by *k*-means and hierarchical clustering (Figure \ref{fig:fig5}).
 
-```{r fig6, fig.cap = "PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al. (2022))"}
-plot_community_graphs(exdata_cohesion_matrix,
-                layout = as.matrix(exdata3),
-                show_labels = FALSE, 
-                only_strong = TRUE,
-                vertex.size = 5)
+```{r fig5, fig.cap = "PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al. (2022)) compared to *k*-means and hierarchical clustering with k = 8."}
+par(mfrow = c(1, 3), pty = "s")
+plot(
+  exdata3,
+  pch = 16,
+  col = pald_colors[exdata_pald],
+  xlab = "",
+  ylab = "",
+  main = "PaLD Clusters"
+)
+plot(
+  exdata3,
+  pch = 16,
+  col = pald_colors[exdata_kmeans],
+  xlab = "",
+  ylab = "",
+  main = "K-Means Clusters (k = 8)"
+)
+plot(
+  exdata3,
+  pch = 16,
+  col = pald_colors[exdata_hclust],
+  xlab = "",
+  ylab = "",
+  main = "Hiearchical Clusters (k = 8)"
+)
 ```
+
+Cohesion is particularly useful when considering data with varying local density, see discussion in [@berenhaut2022social].
+Note that the PaLD algorithm is able to detect the eight natural groups within the data without the use of any additional inputs (e.g., number of clusters) nor optimization criteria.
+Despite providing the "correct" number of clusters (i.e., $k = 8$) both *k*-means and hierarchical clustering did not give the desired result.
 

### 2022-03-17 r-journal/manuscript.tex
update paper

@@
 synthetic data set of two-dimensional points created by
-\citet{gionis1clustering} to demonstrate clustering aggregation, a
-sample of walking distances from a pump in the infamous cholera outbreak
-\citep{cholera}, clustering data generated from the scikit-learn Python
-package \citep{pedregosa2011scikit}, and data compiled by \cite{tissue}
-of tissue gene expressions.
+\citet{gionis1clustering} to demonstrate clustering aggregation,
+clustering data generated from the scikit-learn Python package
+\citep{pedregosa2011scikit}, data describing cognate relationships
+between words across 87 Indo-European languages \citep{dyen92}, data
+compiled by \cite{tissue} of tissue gene expressions, and three example
+data sets generated for the \citet{berenhaut2022social} paper.
 
@@
 
-\hypertarget{helper-functions-to-create-contribution-matrix}{%
-\subsubsection{Helper functions to create contribution
-matrix}\label{helper-functions-to-create-contribution-matrix}}
+\hypertarget{creating-the-cohesion-matrix}{%
+\subsubsection{Creating the cohesion
+matrix}\label{creating-the-cohesion-matrix}}
 
@@
 df <- data.frame(
-  x1 = c(6, 8, 11, 16, 4),
-  x2 = c(5, 4, 13, 7, 18)
+  x1 = c(6, 8, 11, 16, 4, 14),
+  x2 = c(5, 4, 13, 7, 6, 10)
 )
-rownames(df) <- c("A", "B", "C", "D", "E")
+rownames(df) <- c("A", "B", "C", "D", "E", "F")
 \end{Sinput}
@@
 
-This will create an \(n\times n\) distance matrix, where \(n\)
-corresponds to the number of observations in the original data frame, in
-this example \(n = 5\). This distance matrix can then be passed to the
-\texttt{cohesion\_matrix()} function in order to calculate the pairwise
-cohesion values. Cohesion is an interpretable probability that reflects
-the strength of alignment of two points. Again, if the user begins with
-a distance matrix, they can skip the first step and simply input the
-distance matrix into this function.
+This will create a \texttt{dist} object. If converted to a matrix, this
+will be a \(n\times n\) distance matrix, where \(n\) corresponds to the
+number of observations in the original data frame, in this example
+\(n = 6\).
+
+This dist object, or equivalently a distance matrix, can then be passed
+to the \texttt{cohesion\_matrix()} function in order to calculate the
+pairwise cohesion values. Cohesion is an interpretable probability that
+reflects the strength of alignment of two points. Again, if the user
+begins with a distance matrix, they can skip the first step and simply
+input the distance matrix into this function.
 
@@
 \begin{Soutput}
-#>        A      B         C         D         E
-#> A 0.2875 0.1625 0.0000000 0.0500000 0.0000000
-#> B 0.1625 0.2875 0.0000000 0.0500000 0.0000000
-#> C 0.0000 0.0000 0.3083333 0.1833333 0.1000000
-#> D 0.0500 0.0500 0.1750000 0.3000000 0.0000000
-#> E 0.0000 0.0000 0.1000000 0.0000000 0.2333333
+#>           A         B         C         D         E         F
+#> A 0.2333333 0.1666667 0.0000000 0.0000000 0.1666667 0.0000000
+#> B 0.1333333 0.2333333 0.0000000 0.0000000 0.1000000 0.0000000
+#> C 0.0000000 0.0000000 0.2333333 0.1000000 0.0000000 0.1000000
+#> D 0.0000000 0.0000000 0.1000000 0.2666667 0.0000000 0.1666667
+#> E 0.1333333 0.1000000 0.0000000 0.0000000 0.2333333 0.0000000
+#> F 0.0000000 0.0000000 0.1000000 0.1666667 0.0000000 0.2666667
+#> attr(,"class")
+#> [1] "cohesion_matrix" "matrix"          "array"
 \end{Soutput}
@@
 \begin{Sinput}
-d |>
+df |>
   dist() |>
@@
 #> C     C       2
-#> D     D       2
-#> E     E       3
+#> D     D       3
+#> E     E       1
+#> F     F       3
 \end{Soutput}
@@
 
-In this example, three clusters are identified with these five points.
-Points \texttt{A} and \texttt{B} fall into cluster 1. Points \texttt{C}
-and \texttt{D} into cluster 2, and point \texttt{E} in cluster 3.
+In this example, three clusters are identified with these six points.
+Points \texttt{A}, \texttt{B}, and \texttt{E} fall into cluster 1. Point
+\texttt{C} is in cluster 2 and points \texttt{D} and \texttt{F} fall
+into cluster 3.
 
@@
 \begin{Soutput}
-#>         A         B         C         D         E 
-#> 0.4062500 0.5416667 0.3833333 0.4687500 0.7000000
+#>         A         B         C         D         E         F 
+#> 0.5666667 0.4666667 0.4333333 0.5333333 0.4666667 0.5333333
 \end{Soutput}
@@
 
-In this case, the deepest point is \texttt{C}.
+In this case, the deepest point is \texttt{A}.
 
@@
 \begin{Soutput}
-#> [1] 0.14
+#> [1] 0.1222222
 \end{Soutput}
@@
 
-In this case, the threshold is \texttt{0.14}.
+In this case, the threshold is \texttt{0.12}.
 
@@
 \begin{Soutput}
-#>        A      B     C     D      E
-#> A 0.3000 0.1625 0.000 0.000 0.0000
-#> B 0.1625 0.2875 0.000 0.000 0.0000
-#> C 0.0000 0.0000 0.300 0.175 0.0000
-#> D 0.0000 0.0000 0.175 0.300 0.0000
-#> E 0.0000 0.0000 0.000 0.000 0.2125
+#>           A         B         C         D         E         F
+#> A 0.2333333 0.1333333 0.0000000 0.0000000 0.1333333 0.0000000
+#> B 0.1333333 0.2333333 0.0000000 0.0000000 0.0000000 0.0000000
+#> C 0.0000000 0.0000000 0.2333333 0.0000000 0.0000000 0.0000000
+#> D 0.0000000 0.0000000 0.0000000 0.2666667 0.0000000 0.1666667
+#> E 0.1333333 0.0000000 0.0000000 0.0000000 0.2333333 0.0000000
+#> F 0.0000000 0.0000000 0.0000000 0.1666667 0.0000000 0.2666667
+#> attr(,"class")
+#> [1] "cohesion_matrix" "matrix"          "array"
 \end{Soutput}
@@
 \begin{Soutput}
-#> IGRAPH 261f39e UNW- 5 2 -- 
+#> IGRAPH dcdcffa UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from 261f39e (vertex names):
-#> [1] A--B C--D
+#> + edges from dcdcffa (vertex names):
+#> [1] A--B A--E D--F
 \end{Soutput}
@@
 
-Here we see that there are two connected components, points \texttt{A}
-and \texttt{B}, which form the first cluster, and points \texttt{D} and
-\texttt{C} which form the second.
+Here we see that there are three connected components, points \texttt{A}
+and \texttt{B}, and \texttt{A} and \texttt{E} which form the first
+cluster, and points \texttt{D} and \texttt{F} which form another.
 
@@
 can begin by visualizing the points in data frame \texttt{df} (Figure
-\ref{fig:fig1}).
+\ref{fig:fig1}). When visualizing these points, it is important to have
+a square plot so as to not distort the distances. When using the
+\CRANpkg{ggplot2} package for this visualization, you can use the
+\texttt{aspect.ratio\ \ =\ 1} argument in the \texttt{theme()} function.
+If using the \texttt{plot()} function included in the base library, you
+can set the plot area to be a square by using \texttt{par(pty\ =\ "s")}
+prior to calling the \texttt{plot()} function, as shown below.
 
@@
 ggplot(df, aes(x1, x2)) +
-  geom_text(label = rownames(df))
+  geom_text(label = rownames(df)) + 
+  theme(aspect.ratio = 1)
 \end{Sinput}
@@
 
-We can then pass the cohesion matrix to the
+We can pass the cohesion matrix to the
 \texttt{plot\_community\_graphs()} function to view the relationship
@@
 \begin{Sinput}
-df |>
-  dist() |>
-  cohesion_matrix() |>
-  plot_community_graphs()
+plot_community_graphs(cohesion)
 \end{Sinput}
@@
 \texttt{as.matrix(df)}, or a matrix of the data frame \texttt{df} to the
-\texttt{layout} argument (Figure \ref{fig:fig3}.
+\texttt{layout} argument (Figure \ref{fig:fig3}. Notice in the code
+below, we set \texttt{par(pty\ =\ "s")} in order to ensure that we have
+a square plotting region. Additionally, this
+\texttt{plot\_community\_graphs()} function will also permit parameters
+that can be passed to the \texttt{plot.igraph()} function. For example,
+we can pass arguments to the \texttt{plot.igraph} function via the
+\texttt{...}, for example to increase the vertex size and change the
+vertex label color, we can specify \texttt{vertex.size\ =\ 15} and
+\texttt{vertex.label.color\ =\ "white"}.
 
@@
 \begin{Sinput}
-df |>
-  dist() |>
-  cohesion_matrix() |>
-  plot_community_graphs(layout = as.matrix(df))
+par(pty = "s")
+
+plot_community_graphs(cohesion, 
+                      layout = as.matrix(df),
+                      vertex.size = 15,
+                      vertex.label.color = "white")
 \end{Sinput}
@@
 
-This \texttt{plot\_community\_graphs()} function will also permit
-parameters that can be passed to the \texttt{plot.igraph()} function.
-For example, to add axes to the graph, the user can pass the
-\texttt{axes\ =\ TRUE} argument to the \texttt{...} in the
-\texttt{plot\_community\_graphs()} function (Figure \ref{fig:fig4}).
-
-\begin{Schunk}
-\begin{Sinput}
-df |>
-  dist() |>
-  cohesion_matrix() |>
-  plot_community_graphs(layout = as.matrix(df),
-                        axes = TRUE)
-\end{Sinput}
-\begin{figure}
-\includegraphics{manuscript_files/figure-latex/fig4-1} \caption[PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1, adding axes]{PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1, adding axes}\label{fig:fig4}
-\end{figure}
-\end{Schunk}
-
 \hypertarget{examples}{%
@@
 
-We will demonstrate the utility of the \CRANpkg{pald} package in two
+We will demonstrate the utility of the \CRANpkg{pald} package in three
 clustering examples.
@@
 \citet{mccall2014gene}, obtained from the \textbf{tissuesGeneExpression}
-bioconductor package \citep{tissue}. A cohesion matrix was created using
-this data set and is included the \CRANpkg{pald} package in an object
-called \texttt{tissue\_cohesion\_matrix}.
+bioconductor package \citep{tissue}. A \texttt{dist} object was created
+using this data set and is included the \CRANpkg{pald} package in an
+object called \texttt{tissue\_dist}.
 
-The \texttt{tissue\_cohesion\_matrix} object is a cohesion matrix with
-189 rows and 189 columns.
+The \texttt{tissue\_dist} object is a \texttt{dist} object resulting in
+a distance matrix with 189 rows and 189 columns.
 
-We can use this contribution matrix to display the relationship between
-tissue samples using the \texttt{plot\_community\_graphs()} function
-(Figure \ref{fig:fig5}). For clarity of the display, we show how to only
-include a random set of the vertex labels. We can pass this random set
-through the \texttt{...} to the \texttt{plot.igraph}
-\texttt{vertex.label} parameter.
+We can create the cohesion matrix using the \texttt{cohesion\_matrix}
+function.
 
@@
 \begin{Sinput}
-set.seed(1)
-
-labels <- rownames(tissue_cohesion_matrix)
-labels[sample(1:189, 125)] <- ""
-plot_community_graphs(tissue_cohesion_matrix,
-                      vertex.label = labels,
-                      vertex.size = 4)
+tissue_cohesion <- cohesion_matrix(tissue_dist)
 \end{Sinput}
-\begin{figure}
-\includegraphics{manuscript_files/figure-latex/fig5-1} \caption[PaLD clustering of tissue data]{PaLD clustering of tissue data}\label{fig:fig5}
-\end{figure}
 \end{Schunk}
@@
 \begin{Sinput}
-library(dplyr)
-community_clusters(tissue_cohesion_matrix) %>%
-  group_by(cluster, point) %>%
-  count()
+community_clusters(tissue_cohesion) |>
+  dplyr::count(cluster, point)
 \end{Sinput}
 \begin{Soutput}
-#> # A tibble: 19 x 3
-#> # Groups:   cluster, point [19]
 #>    cluster       point  n
-#>      <dbl> <chr>       <int>
 #> 1        1 endometrium 15
@@
 
+We can also display the relationships between tissue samples using the
+\texttt{plot\_community\_graphs()} function (Figure \ref{fig:fig4}). For
+clarity of the display, we show how to remove the labels using
+\texttt{show\_labels\ =\ FALSE}. We will instead color by the labels by
+passing these to the \texttt{vertex.color} parameter through the
+\texttt{...} to the \texttt{plot.igraph} function. Similarly, we can add
+a legend using the \texttt{legend()} function, as you would for a
+\CRANpkg{igraph} visualization. Additionally, we use the
+\texttt{edge\_width\_factor} and \texttt{emph\_strong} arguments to
+adjust the width of the lines between and within PaLD clusters.
+
+\begin{Schunk}
+\begin{Sinput}
+labels <- rownames(tissue_cohesion)
+plot_community_graphs(tissue_cohesion,
+                      show_labels = FALSE,
+                      vertex.size = 4,
+                      vertex.color = as.factor(labels),
+                      edge_width_factor = 35,
+                      emph_strong = 5) 
+legend("topleft", 
+       legend = unique(as.factor(labels)), 
+       pt.bg = unique(as.factor(labels)),
+       col = "black",
+       pch = 21)
+\end{Sinput}
+\end{Schunk}
+
+\begin{Schunk}
+\begin{figure}
+\includegraphics[width=1\linewidth]{fig5} \caption[PaLD clustering of tissue data]{PaLD clustering of tissue data.
+The line colors indicate the PaLD clusters, the point colors indicate the tissue classification.}\label{fig:fig4}
+\end{figure}
+\end{Schunk}
+
+\hypertarget{cognate-based-language-families}{%
+\subsection{Cognate-based Language
+Families}\label{cognate-based-language-families}}
+
+This example performs a PaLD analysis on a data set from \citet{dyen92}
+that examines the relationship between 87 Indo-European languages from
+the perspective of cognates. A \texttt{dist} object was created from
+this data set and included in the \CRANpkg{pald} package in an object
+called \texttt{cognate\_dist}.
+
+This example will demonstrate how you can apply functions in the
+\CRANpkg{igraph} package to objects output from the \CRANpkg{pald}
+package. We can first use the \texttt{cohesion\_matrix()} function to
+calculate the cohesion matrix and the \texttt{community\_graphs()}
+function to create a list with the weighted community graph, the
+weighted community graph with only strong ties included (and all others
+set to 0), and the layout. From this, we can extract the graph with only
+the strong ties, here called \texttt{cognate\_graph\_strong}.
+
+\begin{Schunk}
+\begin{Sinput}
+cognate_cohesion <- cohesion_matrix(cognate_dist)
+cognate_graphs <- community_graphs(cognate_cohesion)
+
+cognate_graph_strong <- cognate_graphs[["G_strong"]]
+\end{Sinput}
+\end{Schunk}
+
+We can then use the \texttt{neighbors()} function from the
+\CRANpkg{igraph} package to extract the neighbors in this graph. For
+example, if we wanted to extract all neighbors where the language is
+``French'', we would run the following.
+
+\begin{Schunk}
+\begin{Sinput}
+french_neighbors <- igraph::neighbors(cognate_graph_strong, "French")
+french_neighbors
+\end{Sinput}
+\begin{Soutput}
+#> + 8/87 vertices, named, from 64a847b:
+#> [1] Italian         Ladin           Provencal       Walloon        
+#> [5] French_Creole_C French_Creole_D Spanish         Catalan
+\end{Soutput}
+\end{Schunk}
+
+Similarly, we can print the associated neighborhood weights by
+subsetting the cohesion matrix.
+
+\begin{Schunk}
+\begin{Sinput}
+cognate_cohesion["French", french_neighbors]
+\end{Sinput}
+\begin{Soutput}
+#>         Italian           Ladin       Provencal         Walloon French_Creole_C 
+#>      0.01997696      0.02094596      0.02871174      0.03258771      0.02406057 
+#> French_Creole_D         Spanish         Catalan 
+#>      0.02406057      0.01679733      0.01859688
+\end{Soutput}
+\end{Schunk}
+
 \hypertarget{clustering-generated-data}{%
@@
 generated from bivariate normal distributions with varying means and
-variances.
+variances. There are eight ``true'' clusters.
+
+We will demonstrate how we can compare PaLD to two clustering methods:
+\emph{k}-means and hierarchical clustering. The code below calculates
+the cohesion matrix (\texttt{exdata\_cohesion}) as well as the clusters
+via PaLD (\texttt{exdata\_pald}), \emph{k}-means
+(\texttt{exdata\_kmeans}) and hierarchical clustering
+(\texttt{exdata\_hclust}).
 
@@
 \begin{Sinput}
-exdata_cohesion_matrix <- exdata3 |>
+exdata_cohesion <- exdata3 |>
   dist() |>
   cohesion_matrix()
+
+exdata_pald <- community_clusters(exdata_cohesion)$cluster
+
+exdata_kmeans <- kmeans(exdata3, 8)$cluster
+
+exdata_hclust <- exdata3 |>
+  dist() |>
+  hclust() |>
+  cutree(k = 8) 
 \end{Sinput}
@@
 
-When plotting the \texttt{exdata3} PaLD graph, we want the layout to
-match the layout of the original data, so we will pass
-\texttt{as.matrix(exdata3)} to the \texttt{layout} parameter in the
-\texttt{plot\_community\_graphs()} function. Additionally, here the row
-names are meaningless, they just correspond to the location of the
-generated data, so we can remove the labels on the plot by setting
-\texttt{show\_labels\ =\ FALSE}. The \texttt{only\_strong} parameter
-will only display the strongly connected edges. Additionally, we can
-optionally pass parameters through the \texttt{...} to
-\texttt{plot::igraph} function; for example, to increase the vertex size
-we can set \texttt{vertex.size\ =\ 5}. Figure \ref{fig:fig6} shows this.
+We can compare this to the clustering generated by \emph{k}-means and
+hierarchical clustering (Figure \ref{fig:fig5}).
 
@@
 \begin{Sinput}
-plot_community_graphs(exdata_cohesion_matrix,
-                layout = as.matrix(exdata3),
-                show_labels = FALSE, 
-                only_strong = TRUE,
-                vertex.size = 5)
+par(mfrow = c(1, 3), pty = "s")
+plot(
+  exdata3,
+  pch = 16,
+  col = pald_colors[exdata_pald],
+  xlab = "",
+  ylab = "",
+  main = "PaLD Clusters"
+)
+plot(
+  exdata3,
+  pch = 16,
+  col = pald_colors[exdata_kmeans],
+  xlab = "",
+  ylab = "",
+  main = "K-Means Clusters (k = 8)"
+)
+plot(
+  exdata3,
+  pch = 16,
+  col = pald_colors[exdata_hclust],
+  xlab = "",
+  ylab = "",
+  main = "Hiearchical Clusters (k = 8)"
+)
 \end{Sinput}
 \begin{figure}
-\includegraphics{manuscript_files/figure-latex/fig6-1} \caption[PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al]{PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al.
-(2022))}\label{fig:fig6}
+\includegraphics{manuscript_files/figure-latex/fig5-1} \caption[PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al]{PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al.
+(2022)) compared to *k*-means and hierarchical clustering with k = 8.}\label{fig:fig5}
 \end{figure}
 \end{Schunk}
+
+Cohesion is particularly useful when considering data with varying local
+density, see discussion in \citep{berenhaut2022social}. Note that the
+PaLD algorithm is able to detect the eight natural groups within the
+data without the use of any additional inputs (e.g., number of clusters)
+nor optimization criteria. Despite providing the ``correct'' number of
+clusters (i.e., \(k = 8\)) both \emph{k}-means and hierarchical
+clustering did not give the desired result.
 

### 2022-03-17 r-journal/manuscript.R

@@ -8,10 +8,10 @@ knitr::opts_chunk$set(fig.height = 4, message = FALSE, warning = FALSE)
 ## -----------------------------------------------------------------------------
 library(pald)
 df <- data.frame(
-  x1 = c(6, 8, 11, 16, 4),
-  x2 = c(5, 4, 13, 7, 18)
+  x1 = c(6, 8, 11, 16, 4, 14),
+  x2 = c(5, 4, 13, 7, 6, 10)
 )
-rownames(df) <- c("A", "B", "C", "D", "E")
+rownames(df) <- c("A", "B", "C", "D", "E", "F")
 
 
 ## -----------------------------------------------------------------------------
@@ -30,7 +30,7 @@ df |>
 
 
 ## -----------------------------------------------------------------------------
-d |>
+df |>
   dist() |>
   cohesion_matrix() -> cohesion
 
@@ -63,58 +63,125 @@ graphs[["G_strong"]]
 ## ----fig1, fig.cap = "Visualize the points from data frame `df`", message = FALSE, warning = FALSE----
 library(ggplot2)
 ggplot(df, aes(x1, x2)) +
-  geom_text(label = rownames(df))
+  geom_text(label = rownames(df)) + 
+  theme(aspect.ratio = 1)
 
 
 ## ----fig2, fig.cap = "PaLD graph displaying the relationship between the points in data frame `df`"----
-df |>
-  dist() |>
-  cohesion_matrix() |>
-  plot_community_graphs()
+plot_community_graphs(cohesion)
 
 
 ## ----fig3, fig.cap = "PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1"----
-df |>
-  dist() |>
-  cohesion_matrix() |>
-  plot_community_graphs(layout = as.matrix(df))
+par(pty = "s")
+
+plot_community_graphs(cohesion, 
+                      layout = as.matrix(df),
+                      vertex.size = 15,
+                      vertex.label.color = "white")
 
 
-## ----fig4, fig.cap = "PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1, adding axes"----
-df |>
-  dist() |>
-  cohesion_matrix() |>
-  plot_community_graphs(layout = as.matrix(df),
-                        axes = TRUE)
+## -----------------------------------------------------------------------------
+tissue_cohesion <- cohesion_matrix(tissue_dist)
 
 
[…]

### 2022-03-22 r-journal/manuscript.Rmd
update paper

@@
 
-* Describe PaLD (Ken & Kate?) [@berenhaut2022social]
+Partitioned Local Depths (PaLD) is a framework for a holistic consideration of the community structure of distance-based data.
+Leveraging a socially inspired perspective, the method provides network-based community information which is founded on a new measure of local depth and pairwise cohesion (partitioned local depth).
+The method does not require distributional assumptions, optimization criteria, nor extraneous inputs.
+A complete description of the perspective, together with a discussion of the underlying social motivation, theoretical results, and applications to additional data sets is provided in \citet{berenhaut2022social}.
 
-We present a new package, \CRANpkg{pald}, for calculating partitioned local depth (PaLD) probabilities, implementing clustering analyses, and creating data visualizations to represent the clusters.
+Building on existing approaches to (global) depth, local depth expresses features of centrality as an interpretable probability which is free of parameters and robust to outliers.
+Then, partitioning the probability which defines local depth, we obtain a measure of cohesion between pairs of points.
+Both local depth and cohesion reflect aspects of relative position (rather than absolute distance) and provide a straightforward way to account for varying density across the space.
+Specifically, provided that two sets are separated (in the sense that the minimum between-set distance is greater than the maximum within-set distance), cohesion is invariant under the contraction and dilation of the distances within each set.
+This property may be particularly valuable when one has reason to believe that the magnitude of distances varies across the space.
+
+As cohesion captures a sense of the relationship strength between points, we can then visualize the resulting community structure with a network whose edges are weighted by (mutual) cohesion.
+The underlying social framework motivates an elegant threshold for distinguishing between strongly and weakly cohesive pairs.
+
+As seen throughout this paper, we display the network obtained from cohesion using a force-directed graph drawing algorithm and emphasize the strong ties (colored by connected component).
+We refer to the connected components of the network of strong ties as "clusters."  Note that to qualify as a cluster in this definition, one may not have any strong ties with those outside the cluster, and thus the existence of disjoint groups is a strong signal for separation.
+Here, clusters are identified without additional user inputs nor optimization criteria.
+If one wishes to further break the community graph into groups, one may use community detection methods for networks (such as spectral clustering or the Louvain algorithm).
+Though only briefly considered here, one may also use the collection of strong ties in place of (weighted) k-nearest neighbors in settings such as classification and smoothing.
+Overall, the structural information obtained from local depth, cohesion and community graphs can provide a holistic perspective on the data which does not require the use of distributional assumptions, optimization criteria nor additional user inputs.
+
+We present a new package, \CRANpkg{pald}, for calculating Partitioned Local Depths (PaLD) probabilities, implementing clustering analyses, and creating data visualizations to represent the clusters.
 This paper will describe how to use the package as well as walk through two examples.

### 2022-03-22 r-journal/manuscript.tex
update paper

@@
 
-\begin{itemize}
-\tightlist
-\item
-  Describe PaLD (Ken \& Kate?) \citep{berenhaut2022social}
-\end{itemize}
+Partitioned Local Depths (PaLD) is a framework for a holistic
+consideration of the community structure of distance-based data.
+Leveraging a socially inspired perspective, the method provides
+network-based community information which is founded on a new measure of
+local depth and pairwise cohesion (partitioned local depth). The method
+does not require distributional assumptions, optimization criteria, nor
+extraneous inputs. A complete description of the perspective, together
+with a discussion of the underlying social motivation, theoretical
+results, and applications to additional data sets is provided in
+\citet{berenhaut2022social}.
 
-We present a new package, \CRANpkg{pald}, for calculating partitioned
-local depth (PaLD) probabilities, implementing clustering analyses, and
+Building on existing approaches to (global) depth, local depth expresses
+features of centrality as an interpretable probability which is free of
+parameters and robust to outliers. Then, partitioning the probability
+which defines local depth, we obtain a measure of cohesion between pairs
+of points. Both local depth and cohesion reflect aspects of relative
+position (rather than absolute distance) and provide a straightforward
+way to account for varying density across the space. Specifically,
+provided that two sets are separated (in the sense that the minimum
+between-set distance is greater than the maximum within-set distance),
+cohesion is invariant under the contraction and dilation of the
+distances within each set. This property may be particularly valuable
+when one has reason to believe that the magnitude of distances varies
+across the space.
+
+As cohesion captures a sense of the relationship strength between
+points, we can then visualize the resulting community structure with a
+network whose edges are weighted by (mutual) cohesion. The underlying
+social framework motivates an elegant threshold for distinguishing
+between strongly and weakly cohesive pairs.
+
+As seen throughout this paper, we display the network obtained from
+cohesion using a force-directed graph drawing algorithm and emphasize
+the strong ties (colored by connected component). We refer to the
+connected components of the network of strong ties as ``clusters.'' Note
+that to qualify as a cluster in this definition, one may not have any
+strong ties with those outside the cluster, and thus the existence of
+disjoint groups is a strong signal for separation. Here, clusters are
+identified without additional user inputs nor optimization criteria. If
+one wishes to further break the community graph into groups, one may use
+community detection methods for networks (such as spectral clustering or
+the Louvain algorithm). Though only briefly considered here, one may
+also use the collection of strong ties in place of (weighted) k-nearest
+neighbors in settings such as classification and smoothing. Overall, the
+structural information obtained from local depth, cohesion and community
+graphs can provide a holistic perspective on the data which does not
+require the use of distributional assumptions, optimization criteria nor
+additional user inputs.
+
+We present a new package, \CRANpkg{pald}, for calculating Partitioned
+Local Depths (PaLD) probabilities, implementing clustering analyses, and
 creating data visualizations to represent the clusters. This paper will
@@
 \begin{Soutput}
-#> IGRAPH dcdcffa UNW- 6 3 -- 
+#> IGRAPH b3faedc UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from dcdcffa (vertex names):
+#> + edges from b3faedc (vertex names):
 #> [1] A--B A--E D--F
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from 64a847b:
+#> + 8/87 vertices, named, from 46d19b2:
 #> [1] Italian         Ladin           Provencal       Walloon        

### 2022-05-22 r-journal/manuscript.Rmd
update with Ken's comments

@@
 Both local depth and cohesion reflect aspects of relative position (rather than absolute distance) and provide a straightforward way to account for varying density across the space.
-Specifically, provided that two sets are separated (in the sense that the minimum between-set distance is greater than the maximum within-set distance), cohesion is invariant under the contraction and dilation of the distances within each set.
+Specifically, as shown in \citet{berenhaut2022social}, provided that two sets are separated (in the sense that the minimum between-set distance is greater than the maximum within-set distance), cohesion is invariant under the contraction and dilation of the distances within each set.
 This property may be particularly valuable when one has reason to believe that the magnitude of distances varies across the space.
@@
 As seen throughout this paper, we display the network obtained from cohesion using a force-directed graph drawing algorithm and emphasize the strong ties (colored by connected component).
-We refer to the connected components of the network of strong ties as "clusters."  Note that to qualify as a cluster in this definition, one may not have any strong ties with those outside the cluster, and thus the existence of disjoint groups is a strong signal for separation.
+We refer to the connected components of the network of strong ties as community "clusters."  Note that to qualify as a cluster in this definition, one may not have any strong ties with those outside the cluster, and thus the existence of disjoint groups is a strong signal for separation.
 Here, clusters are identified without additional user inputs nor optimization criteria.
@@
 
-We present a new package, \CRANpkg{pald}, for calculating Partitioned Local Depths (PaLD) probabilities, implementing clustering analyses, and creating data visualizations to represent the clusters.
+We present a new package, \CRANpkg{pald}, for calculating Partitioned Local Depths (PaLD) probabilities, implementing clustering analyses, and creating data visualizations to display community structure.
 This paper will describe how to use the package as well as walk through two examples.
@@
 
-1. Helper functions to organize data into the correct format, into distance matrices and then cohesion matrices
-2. Functions that convert a cohesion matrix into a variety of useful formats, including partitioned local depths, clusters, and graphs
-3. Plotting functions
+1. A function for computing the cohesion matrix
+2. Functions for extracting useful information from the cohesion matrix, such as local depths, neighbors, clusters, and graph objects
+3. Plotting functions for community graphs
 
@@
 
-For demonstration purposes, below is a sample data frame with two variables, `x1` and `x2`.
+The input for the Partitioned Local Depths (PaLD) is a distance matrix or `dist` object.
+Note that the collection of input distances (or dissimilarities) does not need to satisfy the triangle inequality.
+
+For demonstration purposes, we will show how one can compute a distance matrix from an input data frame with, say, two variables `x1` and `x2`.
+The input data may be of any dimension.
 The methods put forth here work on data frames with higher dimensions, as described in the **Examples** section; we are simply choosing a small data frame here for demonstration purposes.
@@
 df <- data.frame(
-  x1 = c(6, 8, 11, 16, 4, 14),
-  x2 = c(5, 4, 13, 7, 6, 10)
+  x1 = c(6, 8, 8, 16, 4, 14),
+  x2 = c(5, 4, 10, 8, 4, 10)
 )
@@
 
-The first step needed to calculate the partitioned local depths is to construct a _distance matrix_.
-If the data are already in this form, the user can skip to the next step.
 The `dist()` function converts an input data frame into a distance matrix, as demonstrated below.
+If the data are already provided as a distance matrix (or `dist` object), the user can skip to the next step.
 
@@
 
-This will create a `dist` object.
+The function above creates a `dist` object.
 If converted to a matrix, this will be a $n\times n$ distance matrix, where $n$ corresponds to the number of observations in the original data frame, in this example $n = 6$.
 
-This dist object, or equivalently a distance matrix, can then be passed to the `cohesion_matrix()` function in order to calculate the pairwise cohesion values.
-Cohesion is an interpretable probability that reflects the strength of alignment of two points.
-Again, if the user begins with a distance matrix, they can skip the first step and simply input the distance matrix into this function.
+This `dist` object, or a distance matrix, can then be passed to the `cohesion_matrix()` function in order to calculate the pairwise cohesion values.
+
+Cohesion is an interpretable probability that reflects the strength of alignment of two points within local regions.
+It captures aspects of the relative positioning of points and accounts for varying density across the space.
 
@@
 
-The _cohesion matrix_ output by the `cohesion_matrix()` is the main input for the majority of the remaining functions.
+The _cohesion matrix_ output by the `cohesion_matrix()` function is the main input for the majority of the remaining functions.
 
-### Functions that convert a cohesion matrix into useful formats
+### Functions for extracting information from the cohesion matrix
 
@@
 ```
-
-To calculate the _clusters_ that each point will fall into, we can use the `community_clusters()` function.
-This will output a data frame with two columns, the first will correspond to the `point`, as identified by the row name of the original input data frame, `df`, the second will identify the `cluster` that each point belongs to.
-
-```{r}
-community_clusters(cohesion)
-```
-
-In this example, three clusters are identified with these six points. Points `A`, `B`, and `E` fall into cluster 1. Point `C` is in cluster 2 and points `D` and `F` fall into cluster 3.
 
@@
 The `strong_threshold()` function will calculate the cohesion threshold for strong ties.
-This is equal to half the average of the diagonal cohesion matrix.
-This is a threshold that may be used to distinguish between strong and weak ties.
+This is equal to half the average of the diagonal cohesion matrix.\citep{berenhaut2022social} This is a threshold that may be used to distinguish between strong and weak ties.
 
@@
 
-In this case, the threshold is `0.12`.
-
-The `any_isolated()` function will check whether there are any isolated points that will inadvertently be dropped by a graph.
-
-```{r}
-any_isolated(cohesion)
-```
-
-Here, there are no isolated points.
+In this case, the threshold is `0.117`.
 
 The function `cohesion_strong()` will update the cohesion matrix to set all weak ties to zero (via the `strong_threshold()` function).
-Optionally, the matrix will also be symmetrized, with the default parameter `symmetric = TRUE`.
+Optionally, the matrix will also be symmetrized, using the entry-wise (parallel) minimum of the cohesion matrix and its transpose, with the default parameter `symmetric = TRUE`.
 
@@
 
-Finally, the `community_graphs()` function takes the cohesion matrix and creates \CRANpkg{igraph} objects, graphs that describes the relationship between the points.
+The `community_graphs()` function takes the cohesion matrix and creates \CRANpkg{igraph} objects, graphs that describe the relationship between the points.
 This function will output a list of three objects:
@@
 
-Here we see that there are three connected components, points `A` and `B`, and `A` and `E` which form the first cluster, and points `D` and `F` which form another.
+Here we see that there are three connected components, ties `A-B` and  and `A-E` which form the first cluster, and the tie `D-F` which forms another.
+
+The `any_isolated()` function will check whether there are any isolated points that will inadvertently be dropped by a graph.
+
+```{r}
+any_isolated(cohesion)
+```
+
+Here, there are no isolated points.
+
+The "clusters" identified by PaLD are the connected components of the graph of strong ties, `G_strong`.
+To directly calculate them, we can use the `community_clusters()` function.
+This will output a data frame with two columns, the first will correspond to the `point`, as identified by the row name of the original input data frame, `df`, the second will identify the `cluster` that each point belongs to.
+
+```{r}
+community_clusters(cohesion)
+```
+
+In this example, three clusters are identified with these six points. Points `A`, `B`, and `E` fall into cluster 1. Point `C` is in cluster 2 and points `D` and `F` fall into cluster 3.
 
@@
 We can begin by visualizing the points in data frame `df` (Figure \ref{fig:fig1}).
-When visualizing these points, it is important to have a square plot so as to not distort the distances.
-When using the \CRANpkg{ggplot2} package for this visualization, you can use the `aspect.ratio  = 1` argument in the `theme()` function.
-If using the `plot()` function included in the base library, you can set the plot area to be a square by using `par(pty = "s")` prior to calling the `plot()` function, as shown below.
+When visualizing these points, it is important to have the aspect ratio of the x and y axes equal to 1 so as to not distort the distances.
+When using the \CRANpkg{ggplot2} package for this visualization, you can use the `coord_fixed(ratio = 1)` function.
+If using the `plot()` function included in the base library, you can use the `asp = 1` argument.
 
@@
   geom_text(label = rownames(df)) + 
-  theme(aspect.ratio = 1)
+  coord_fixed(ratio = 1) + 
+  xlim(c(4, 16)) + 
+  ylim(c(4, 16))
 ```
@@
 
-The `layout` argument allows the user to pass a matrix to dictate the layout of the graph.
-For example, if we wanted the graph to match the visualization displayed in Figure \ref{fig:fig1}, we can pass `as.matrix(df)`, or a matrix of the data frame `df` to the `layout` argument  (Figure \ref{fig:fig3}.
-Notice in the code below, we set `par(pty = "s")` in order to ensure that we have a square plotting region.
+The `layout` argument allows the user to pass a matrix to dictate the 2-dimensional layout of the graph.
+For example, if we wanted the graph to match the visualization displayed in Figure \ref{fig:fig1}, we can pass `as.matrix(df)`, or a matrix of the data frame `df` to the `layout` argument  (Figure \ref{fig:fig3}).
 Additionally, this `plot_community_graphs()` function will also permit parameters that can be passed to the `plot.igraph()` function.
-For example, we can pass arguments to the `plot.igraph` function via the `...`, for example to increase the vertex size and change the vertex label color, we can specify `vertex.size = 15` and `vertex.label.color = "white"`.
+For example, we can pass arguments to the `plot.igraph` function via `...`, for example to increase the vertex size and change the vertex label color, we can specify `vertex.size = 100` and `vertex.label.color = "white"`.
+Additionally, to allow axes, we use `axes = TRUE` and to put them back on the original scale we set `rescale = FALSE`, resetting the axis limits using `xlim` and `ylim`.
+The `par(pty = "s")` function forces the subsequent plot to be square.
 
-```{r fig3, fig.cap = "PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1"}
+```{r fig3, fig.cap = "PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1"}
 par(pty = "s")
@@
                       layout = as.matrix(df),
-                      vertex.size = 15,
-                      vertex.label.color = "white")
+                      vertex.size = 100,
+                      vertex.label.color = "white",
+                      axes = TRUE,
+                      rescale = FALSE,
+                      asp = 1,
+                      xlim = c(4, 16),
+                      ylim = c(4, 16))
 ```
@@
 
-The first example if from a subset of tissue gene expression data from \citet{zilliox2007gene}, \citet{mccall2011gene}, and \citet{mccall2014gene}, obtained from the **tissuesGeneExpression** bioconductor package \citep{tissue}.
+The first example is from a subset of tissue gene expression data from \citet{zilliox2007gene}, \citet{mccall2011gene}, and \citet{mccall2014gene}, obtained from the **tissuesGeneExpression** bioconductor package \citep{tissue}.
 A `dist` object was created using this data set and is included the \CRANpkg{pald} package in an object called `tissue_dist`.
@@
 We will instead color by the labels by passing these to the `vertex.color` parameter through the  `...` to the `plot.igraph` function.
-Similarly, we can add a legend using the `legend()` function, as you would for a \CRANpkg{igraph} visualization.
+Similarly, we can add a legend using the `legend()` function, as you would for an \CRANpkg{igraph} visualization.
 Additionally, we use the `edge_width_factor` and `emph_strong` arguments to adjust the width of the lines between and within PaLD clusters.
@@
 
-```{r fig4, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "PaLD clustering of tissue data.
+```{r fig4, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Community cluster network for the tissue data.
 The line colors indicate the PaLD clusters, the point colors indicate the tissue classification.", out.width = "100%"}
@@
 
-We can then use the `neighbors()` function from the \CRANpkg{igraph} package to extract the neighbors in this graph.
+We can then use the `neighbors()` function from the \CRANpkg{igraph} package to extract the strong neighbors in this graph.
 For example, if we wanted to extract all neighbors where the language is "French", we would run the following.
@@
 
-The \CRANpkg{pald} includes three randomly generated data frames corresponding to plots from @berenhaut2022social:
+The \CRANpkg{pald} package includes three randomly generated data frames corresponding to plots from @berenhaut2022social:
 
@@
 We will demonstrate how we can compare PaLD to two clustering methods: *k*-means and hierarchical clustering.
-The code below calculates the cohesion matrix (`exdata_cohesion`) as well as the clusters via PaLD (`exdata_pald`), *k*-means (`exdata_kmeans`) and hierarchical clustering (`exdata_hclust`).
+The code below calculates the cohesion matrix (`exdata_cohesion`) as well as the clusters via PaLD (`exdata_pald`), *k*-means (`exdata_kmeans`) and hierarchical clustering using complete linkage (`exdata_hclust`).
 
@@
 
-```{r fig5, fig.cap = "PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al. (2022)) compared to *k*-means and hierarchical clustering with k = 8."}
+```{r fig5, fig.cap = "PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al. (2022)) compared to k-means and hierarchical clustering with k = 8."}
 par(mfrow = c(1, 3), pty = "s")
@@
   ylab = "",
-  main = "PaLD Clusters"
+  main = "PaLD Clusters",
+  asp = 1
 )
@@
   ylab = "",
-  main = "K-Means Clusters (k = 8)"
+  main = "K-Means Clusters (k = 8)",
+  asp = 1
 )
@@
   ylab = "",
-  main = "Hiearchical Clusters (k = 8)"
+  main = "Hiearchical Clusters (k = 8)",
+  asp = 1
 )
@@
 
-The ability of the PaLD algorithm to discern clusters is demonstrated here.
-
 ## Summary
 
-This paper introduces the \CRANpkg{pald} package, demonstrating it's utility for providing a parameter-free clustering algorithm that can easily be applied to any data set.
+This paper introduces the \CRANpkg{pald} package, demonstrating its utility for providing parameter-free clustering which can easily be implemented for a variety of data sets.
 

### 2022-05-22 r-journal/manuscript.tex
update with Ken's comments

@@
 position (rather than absolute distance) and provide a straightforward
-way to account for varying density across the space. Specifically,
-provided that two sets are separated (in the sense that the minimum
-between-set distance is greater than the maximum within-set distance),
-cohesion is invariant under the contraction and dilation of the
-distances within each set. This property may be particularly valuable
-when one has reason to believe that the magnitude of distances varies
-across the space.
+way to account for varying density across the space. Specifically, as
+shown in \citet{berenhaut2022social}, provided that two sets are
+separated (in the sense that the minimum between-set distance is greater
+than the maximum within-set distance), cohesion is invariant under the
+contraction and dilation of the distances within each set. This property
+may be particularly valuable when one has reason to believe that the
+magnitude of distances varies across the space.
 
@@
 the strong ties (colored by connected component). We refer to the
-connected components of the network of strong ties as ``clusters.'' Note
-that to qualify as a cluster in this definition, one may not have any
-strong ties with those outside the cluster, and thus the existence of
-disjoint groups is a strong signal for separation. Here, clusters are
-identified without additional user inputs nor optimization criteria. If
-one wishes to further break the community graph into groups, one may use
-community detection methods for networks (such as spectral clustering or
-the Louvain algorithm). Though only briefly considered here, one may
-also use the collection of strong ties in place of (weighted) k-nearest
-neighbors in settings such as classification and smoothing. Overall, the
-structural information obtained from local depth, cohesion and community
-graphs can provide a holistic perspective on the data which does not
-require the use of distributional assumptions, optimization criteria nor
-additional user inputs.
+connected components of the network of strong ties as community
+``clusters.'' Note that to qualify as a cluster in this definition, one
+may not have any strong ties with those outside the cluster, and thus
+the existence of disjoint groups is a strong signal for separation.
+Here, clusters are identified without additional user inputs nor
+optimization criteria. If one wishes to further break the community
+graph into groups, one may use community detection methods for networks
+(such as spectral clustering or the Louvain algorithm). Though only
+briefly considered here, one may also use the collection of strong ties
+in place of (weighted) k-nearest neighbors in settings such as
+classification and smoothing. Overall, the structural information
+obtained from local depth, cohesion and community graphs can provide a
+holistic perspective on the data which does not require the use of
+distributional assumptions, optimization criteria nor additional user
+inputs.
 
@@
 Local Depths (PaLD) probabilities, implementing clustering analyses, and
-creating data visualizations to represent the clusters. This paper will
-describe how to use the package as well as walk through two examples.
+creating data visualizations to display community structure. This paper
+will describe how to use the package as well as walk through two
+examples.
 
@@
 \item
-  Helper functions to organize data into the correct format, into
-  distance matrices and then cohesion matrices
+  A function for computing the cohesion matrix
 \item
-  Functions that convert a cohesion matrix into a variety of useful
-  formats, including partitioned local depths, clusters, and graphs
+  Functions for extracting useful information from the cohesion matrix,
+  such as local depths, neighbors, clusters, and graph objects
 \item
-  Plotting functions
+  Plotting functions for community graphs
 \end{enumerate}
@@
 
-For demonstration purposes, below is a sample data frame with two
-variables, \texttt{x1} and \texttt{x2}. The methods put forth here work
-on data frames with higher dimensions, as described in the
-\textbf{Examples} section; we are simply choosing a small data frame
+The input for the Partitioned Local Depths (PaLD) is a distance matrix
+or \texttt{dist} object. Note that the collection of input distances (or
+dissimilarities) does not need to satisfy the triangle inequality.
+
+For demonstration purposes, we will show how one can compute a distance
+matrix from an input data frame with, say, two variables \texttt{x1} and
+\texttt{x2}. The input data may be of any dimension. The methods put
+forth here work on data frames with higher dimensions, as described in
+the \textbf{Examples} section; we are simply choosing a small data frame
 here for demonstration purposes.
@@
 
-The first step needed to calculate the partitioned local depths is to
-construct a \emph{distance matrix}. If the data are already in this
-form, the user can skip to the next step. The \texttt{dist()} function
-converts an input data frame into a distance matrix, as demonstrated
-below.
+The \texttt{dist()} function converts an input data frame into a
+distance matrix, as demonstrated below. If the data are already provided
+as a distance matrix (or \texttt{dist} object), the user can skip to the
+next step.
 
@@
 
-This will create a \texttt{dist} object. If converted to a matrix, this
-will be a \(n\times n\) distance matrix, where \(n\) corresponds to the
-number of observations in the original data frame, in this example
-\(n = 6\).
+The function above creates a \texttt{dist} object. If converted to a
+matrix, this will be a \(n\times n\) distance matrix, where \(n\)
+corresponds to the number of observations in the original data frame, in
+this example \(n = 6\).
 
-This dist object, or equivalently a distance matrix, can then be passed
-to the \texttt{cohesion\_matrix()} function in order to calculate the
-pairwise cohesion values. Cohesion is an interpretable probability that
-reflects the strength of alignment of two points. Again, if the user
-begins with a distance matrix, they can skip the first step and simply
-input the distance matrix into this function.
+This \texttt{dist} object, or a distance matrix, can then be passed to
+the \texttt{cohesion\_matrix()} function in order to calculate the
+pairwise cohesion values.
+
+Cohesion is an interpretable probability that reflects the strength of
+alignment of two points within local regions. It captures aspects of the
+relative positioning of points and accounts for varying density across
+the space.
 
@@
 #>            A          B          C         D          E         F
-#> A 0.2333333 0.1666667 0.0000000 0.0000000 0.1666667 0.0000000
-#> B 0.1333333 0.2333333 0.0000000 0.0000000 0.1000000 0.0000000
-#> C 0.0000000 0.0000000 0.2333333 0.1000000 0.0000000 0.1000000
-#> D 0.0000000 0.0000000 0.1000000 0.2666667 0.0000000 0.1666667
-#> E 0.1333333 0.1000000 0.0000000 0.0000000 0.2333333 0.0000000
-#> F 0.0000000 0.0000000 0.1000000 0.1666667 0.0000000 0.2666667
+#> A 0.25000000 0.18333333 0.06666667 0.0000000 0.18333333 0.0000000
+#> B 0.14000000 0.24000000 0.05000000 0.0000000 0.10666667 0.0000000
+#> C 0.07333333 0.07333333 0.20333333 0.0000000 0.03333333 0.0800000
+#> D 0.00000000 0.00000000 0.00000000 0.2333333 0.00000000 0.1333333
+#> E 0.14000000 0.10666667 0.03333333 0.0000000 0.24000000 0.0000000
+#> F 0.00000000 0.00000000 0.05000000 0.1400000 0.00000000 0.2400000
 #> attr(,"class")
@@
 
-The \emph{cohesion matrix} output by the \texttt{cohesion\_matrix()} is
-the main input for the majority of the remaining functions.
+The \emph{cohesion matrix} output by the \texttt{cohesion\_matrix()}
+function is the main input for the majority of the remaining functions.
 
-\hypertarget{functions-that-convert-a-cohesion-matrix-into-useful-formats}{%
-\subsubsection{Functions that convert a cohesion matrix into useful
-formats}\label{functions-that-convert-a-cohesion-matrix-into-useful-formats}}
+\hypertarget{functions-for-extracting-information-from-the-cohesion-matrix}{%
+\subsubsection{Functions for extracting information from the cohesion
+matrix}\label{functions-for-extracting-information-from-the-cohesion-matrix}}
 
@@
 
-To calculate the \emph{clusters} that each point will fall into, we can
-use the \texttt{community\_clusters()} function. This will output a data
-frame with two columns, the first will correspond to the \texttt{point},
-as identified by the row name of the original input data frame,
-\texttt{df}, the second will identify the \texttt{cluster} that each
-point belongs to.
+The \texttt{local\_depths()} function calculates the \emph{depths} of
+each point, outputting a vector of local depths. Local depth is an
+interpretable probability which reflects aspects of relative position
+and centrality via distance comparisons (i.e., \(d(z, x) < d(z, y)\)).
+
+\begin{Schunk}
+\begin{Sinput}
+local_depths(cohesion)
+\end{Sinput}
+\begin{Soutput}
+#>         A         B         C         D         E         F 
+#> 0.6833333 0.5366667 0.4633333 0.3666667 0.5200000 0.4300000
+\end{Soutput}
+\end{Schunk}
+
+In this case, the deepest point is \texttt{A}.
+
+The \texttt{strong\_threshold()} function will calculate the cohesion
+threshold for strong ties. This is equal to half the average of the
+diagonal cohesion matrix.\citep{berenhaut2022social} This is a threshold
+that may be used to distinguish between strong and weak ties.
+
+\begin{Schunk}
+\begin{Sinput}
+strong_threshold(cohesion)
+\end{Sinput}
+\begin{Soutput}
+#> [1] 0.1172222
+\end{Soutput}
+\end{Schunk}
+
+In this case, the threshold is \texttt{0.117}.
+
+The function \texttt{cohesion\_strong()} will update the cohesion matrix
+to set all weak ties to zero (via the \texttt{strong\_threshold()}
+function). Optionally, the matrix will also be symmetrized, using the
+entry-wise (parallel) minimum of the cohesion matrix and its transpose,
+with the default parameter \texttt{symmetric\ =\ TRUE}.
+
+\begin{Schunk}
+\begin{Sinput}
+cohesion_strong(cohesion)
+\end{Sinput}
+\begin{Soutput}
+#>      A    B         C         D    E         F
+#> A 0.25 0.14 0.0000000 0.0000000 0.14 0.0000000
+#> B 0.14 0.24 0.0000000 0.0000000 0.00 0.0000000
+#> C 0.00 0.00 0.2033333 0.0000000 0.00 0.0000000
+#> D 0.00 0.00 0.0000000 0.2333333 0.00 0.1333333
+#> E 0.14 0.00 0.0000000 0.0000000 0.24 0.0000000
+#> F 0.00 0.00 0.0000000 0.1333333 0.00 0.2400000
+#> attr(,"class")
+#> [1] "cohesion_matrix" "matrix"          "array"
+\end{Soutput}
+\end{Schunk}
+
+The \texttt{community\_graphs()} function takes the cohesion matrix and
+creates \CRANpkg{igraph} objects, graphs that describe the relationship
+between the points. This function will output a list of three objects:
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{G}: the weighted (community) graph whose edge weights are
+  mutual cohesion
+\item
+  \texttt{G\_strong}: the weighted (community) graph consisting of edges
+  for which mutual cohesion is greater than the threshold for strong
+  ties
+\item
+  \texttt{layout}: the graph layout, using the Fruchterman Reingold (FR)
+  force-directed graph drawing for the graph \texttt{G}
+\end{itemize}
+
+\begin{Schunk}
+\begin{Sinput}
+graphs <- community_graphs(cohesion)
+graphs[["G_strong"]]
+\end{Sinput}
+\begin{Soutput}
+#> IGRAPH 8eef810 UNW- 6 3 -- 
+#> + attr: name (v/c), weight (e/n)
+#> + edges from 8eef810 (vertex names):
+#> [1] A--B A--E D--F
+\end{Soutput}
+\end{Schunk}
+
+Here we see that there are three connected components, ties \texttt{A-B}
+and and \texttt{A-E} which form the first cluster, and the tie
+\texttt{D-F} which forms another.
+
+The \texttt{any\_isolated()} function will check whether there are any
+isolated points that will inadvertently be dropped by a graph.
+
+\begin{Schunk}
+\begin{Sinput}
+any_isolated(cohesion)
+\end{Sinput}
+\end{Schunk}
+
+Here, there are no isolated points.
+
+The ``clusters'' identified by PaLD are the connected components of the
+graph of strong ties, \texttt{G\_strong}. To directly calculate them, we
+can use the \texttt{community\_clusters()} function. This will output a
+data frame with two columns, the first will correspond to the
+\texttt{point}, as identified by the row name of the original input data
+frame, \texttt{df}, the second will identify the \texttt{cluster} that
+each point belongs to.
 
@@
 
-The \texttt{local\_depths()} function calculates the \emph{depths} of
-each point, outputting a vector of local depths. Local depth is an
-interpretable probability which reflects aspects of relative position
-and centrality via distance comparisons (i.e., \(d(z, x) < d(z, y)\)).
-
-\begin{Schunk}
-\begin{Sinput}
-local_depths(cohesion)
-\end{Sinput}
-\begin{Soutput}
-#>         A         B         C         D         E         F 
-#> 0.5666667 0.4666667 0.4333333 0.5333333 0.4666667 0.5333333
-\end{Soutput}
-\end{Schunk}
-
-In this case, the deepest point is \texttt{A}.
-
-The \texttt{strong\_threshold()} function will calculate the cohesion
-threshold for strong ties. This is equal to half the average of the
-diagonal cohesion matrix. This is a threshold that may be used to
-distinguish between strong and weak ties.
-
-\begin{Schunk}
-\begin{Sinput}
-strong_threshold(cohesion)
-\end{Sinput}
-\begin{Soutput}
-#> [1] 0.1222222
-\end{Soutput}
-\end{Schunk}
-
-In this case, the threshold is \texttt{0.12}.
-
-The \texttt{any\_isolated()} function will check whether there are any
-isolated points that will inadvertently be dropped by a graph.
-
-\begin{Schunk}
-\begin{Sinput}
-any_isolated(cohesion)
-\end{Sinput}
-\end{Schunk}
-
-Here, there are no isolated points.
-
-The function \texttt{cohesion\_strong()} will update the cohesion matrix
-to set all weak ties to zero (via the \texttt{strong\_threshold()}
-function). Optionally, the matrix will also be symmetrized, with the
-default parameter \texttt{symmetric\ =\ TRUE}.
-
-\begin{Schunk}
-\begin{Sinput}
-cohesion_strong(cohesion)
-\end{Sinput}
-\begin{Soutput}
-#>           A         B         C         D         E         F
-#> A 0.2333333 0.1333333 0.0000000 0.0000000 0.1333333 0.0000000
-#> B 0.1333333 0.2333333 0.0000000 0.0000000 0.0000000 0.0000000
-#> C 0.0000000 0.0000000 0.2333333 0.0000000 0.0000000 0.0000000
-#> D 0.0000000 0.0000000 0.0000000 0.2666667 0.0000000 0.1666667
-#> E 0.1333333 0.0000000 0.0000000 0.0000000 0.2333333 0.0000000
-#> F 0.0000000 0.0000000 0.0000000 0.1666667 0.0000000 0.2666667
-#> attr(,"class")
-#> [1] "cohesion_matrix" "matrix"          "array"
-\end{Soutput}
-\end{Schunk}
-
-Finally, the \texttt{community\_graphs()} function takes the cohesion
-matrix and creates \CRANpkg{igraph} objects, graphs that describes the
-relationship between the points. This function will output a list of
-three objects:
-
-\begin{itemize}
-\tightlist
-\item
-  \texttt{G}: the weighted (community) graph whose edge weights are
-  mutual cohesion
-\item
-  \texttt{G\_strong}: the weighted (community) graph consisting of edges
-  for which mutual cohesion is greater than the threshold for strong
-  ties
-\item
-  \texttt{layout}: the graph layout, using the Fruchterman Reingold (FR)
-  force-directed graph drawing for the graph \texttt{G}
-\end{itemize}
-
-\begin{Schunk}
-\begin{Sinput}
-graphs <- community_graphs(cohesion)
-graphs[["G_strong"]]
-\end{Sinput}
-\begin{Soutput}
-#> IGRAPH b3faedc UNW- 6 3 -- 
-#> + attr: name (v/c), weight (e/n)
-#> + edges from b3faedc (vertex names):
-#> [1] A--B A--E D--F
-\end{Soutput}
-\end{Schunk}
-
-Here we see that there are three connected components, points \texttt{A}
-and \texttt{B}, and \texttt{A} and \texttt{E} which form the first
-cluster, and points \texttt{D} and \texttt{F} which form another.
-
 \hypertarget{plotting-functions}{%
@@
 \ref{fig:fig1}). When visualizing these points, it is important to have
-a square plot so as to not distort the distances. When using the
-\CRANpkg{ggplot2} package for this visualization, you can use the
-\texttt{aspect.ratio\ \ =\ 1} argument in the \texttt{theme()} function.
-If using the \texttt{plot()} function included in the base library, you
-can set the plot area to be a square by using \texttt{par(pty\ =\ "s")}
-prior to calling the \texttt{plot()} function, as shown below.
+the aspect ratio of the x and y axes equal to 1 so as to not distort the
+distances. When using the \CRANpkg{ggplot2} package for this
+visualization, you can use the \texttt{coord\_fixed(ratio\ =\ 1)}
+function. If using the \texttt{plot()} function included in the base
+library, you can use the \texttt{asp\ =\ 1} argument.
 
@@
   geom_text(label = rownames(df)) + 
-  theme(aspect.ratio = 1)
+  coord_fixed(ratio = 1) + 
+  xlim(c(4, 16)) + 
+  ylim(c(4, 16))
 \end{Sinput}
@@
 The \texttt{layout} argument allows the user to pass a matrix to dictate
-the layout of the graph. For example, if we wanted the graph to match
-the visualization displayed in Figure \ref{fig:fig1}, we can pass
-\texttt{as.matrix(df)}, or a matrix of the data frame \texttt{df} to the
-\texttt{layout} argument (Figure \ref{fig:fig3}. Notice in the code
-below, we set \texttt{par(pty\ =\ "s")} in order to ensure that we have
-a square plotting region. Additionally, this
-\texttt{plot\_community\_graphs()} function will also permit parameters
-that can be passed to the \texttt{plot.igraph()} function. For example,
-we can pass arguments to the \texttt{plot.igraph} function via the
-\texttt{...}, for example to increase the vertex size and change the
-vertex label color, we can specify \texttt{vertex.size\ =\ 15} and
-\texttt{vertex.label.color\ =\ "white"}.
+the 2-dimensional layout of the graph. For example, if we wanted the
+graph to match the visualization displayed in Figure \ref{fig:fig1}, we
+can pass \texttt{as.matrix(df)}, or a matrix of the data frame
+\texttt{df} to the \texttt{layout} argument (Figure \ref{fig:fig3}).
+Additionally, this \texttt{plot\_community\_graphs()} function will also
+permit parameters that can be passed to the \texttt{plot.igraph()}
+function. For example, we can pass arguments to the \texttt{plot.igraph}
+function via \texttt{...}, for example to increase the vertex size and
+change the vertex label color, we can specify
+\texttt{vertex.size\ =\ 100} and
+\texttt{vertex.label.color\ =\ "white"}. Additionally, to allow axes, we
+use \texttt{axes\ =\ TRUE} and to put them back on the original scale we
+set \texttt{rescale\ =\ FALSE}, resetting the axis limits using
+\texttt{xlim} and \texttt{ylim}. The \texttt{par(pty\ =\ "s")} function
+forces the subsequent plot to be square.
 
@@
                       layout = as.matrix(df),
-                      vertex.size = 15,
-                      vertex.label.color = "white")
+                      vertex.size = 100,
+                      vertex.label.color = "white",
+                      axes = TRUE,
+                      rescale = FALSE,
+                      asp = 1,
+                      xlim = c(4, 16),
+                      ylim = c(4, 16))
 \end{Sinput}
 \begin{figure}
-\includegraphics{manuscript_files/figure-latex/fig3-1} \caption[PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1]{PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1}\label{fig:fig3}
+\includegraphics{manuscript_files/figure-latex/fig3-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1]{PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1}\label{fig:fig3}
 \end{figure}
@@
 
-The first example if from a subset of tissue gene expression data from
+The first example is from a subset of tissue gene expression data from
 \citet{zilliox2007gene}, \citet{mccall2011gene}, and
@@
 \texttt{...} to the \texttt{plot.igraph} function. Similarly, we can add
-a legend using the \texttt{legend()} function, as you would for a
+a legend using the \texttt{legend()} function, as you would for an
 \CRANpkg{igraph} visualization. Additionally, we use the
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{fig5} \caption[PaLD clustering of tissue data]{PaLD clustering of tissue data.
+\includegraphics[width=1\linewidth]{fig5} \caption[Community cluster network for the tissue data]{Community cluster network for the tissue data.
 The line colors indicate the PaLD clusters, the point colors indicate the tissue classification.}\label{fig:fig4}
@@
 We can then use the \texttt{neighbors()} function from the
-\CRANpkg{igraph} package to extract the neighbors in this graph. For
-example, if we wanted to extract all neighbors where the language is
+\CRANpkg{igraph} package to extract the strong neighbors in this graph.
+For example, if we wanted to extract all neighbors where the language is
 ``French'', we would run the following.
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from 46d19b2:
+#> + 8/87 vertices, named, from 9c17296:
 #> [1] Italian         Ladin           Provencal       Walloon        
@@
 
-The \CRANpkg{pald} includes three randomly generated data frames
+The \CRANpkg{pald} package includes three randomly generated data frames
 corresponding to plots from \citet{berenhaut2022social}:
@@
 via PaLD (\texttt{exdata\_pald}), \emph{k}-means
-(\texttt{exdata\_kmeans}) and hierarchical clustering
-(\texttt{exdata\_hclust}).
+(\texttt{exdata\_kmeans}) and hierarchical clustering using complete
+linkage (\texttt{exdata\_hclust}).
 
@@
 \includegraphics{manuscript_files/figure-latex/fig5-1} \caption[PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al]{PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al.
-(2022)) compared to *k*-means and hierarchical clustering with k = 8.}\label{fig:fig5}
+(2022)) compared to k-means and hierarchical clustering with k = 8.}\label{fig:fig5}
 \end{figure}
@@
 
-The ability of the PaLD algorithm to discern clusters is demonstrated
-here.
-
 \hypertarget{summary}{%
@@
 
-This paper introduces the \CRANpkg{pald} package, demonstrating it's
-utility for providing a parameter-free clustering algorithm that can
-easily be applied to any data set.
+This paper introduces the \CRANpkg{pald} package, demonstrating its
+utility for providing parameter-free clustering which can easily be
+implemented for a variety of data sets.
 

### 2022-05-22 r-journal/manuscript.R

@@ -8,8 +8,8 @@ knitr::opts_chunk$set(fig.height = 4, message = FALSE, warning = FALSE)
 ## -----------------------------------------------------------------------------
 library(pald)
 df <- data.frame(
-  x1 = c(6, 8, 11, 16, 4, 14),
-  x2 = c(5, 4, 13, 7, 6, 10)
+  x1 = c(6, 8, 8, 16, 4, 14),
+  x2 = c(5, 4, 10, 8, 4, 10)
 )
 rownames(df) <- c("A", "B", "C", "D", "E", "F")
 
@@ -36,19 +36,11 @@ df |>
 
 
 ## -----------------------------------------------------------------------------
-community_clusters(cohesion)
-
-
-## -----------------------------------------------------------------------------
 local_depths(cohesion)
 
 
 ## -----------------------------------------------------------------------------
 strong_threshold(cohesion)
-
-
-## -----------------------------------------------------------------------------
-any_isolated(cohesion)
 
 
 ## -----------------------------------------------------------------------------
@@ -60,24 +52,39 @@ graphs <- community_graphs(cohesion)
 graphs[["G_strong"]]
 
 
+## -----------------------------------------------------------------------------
+any_isolated(cohesion)
+
+
+## -----------------------------------------------------------------------------
+community_clusters(cohesion)
+
+
 ## ----fig1, fig.cap = "Visualize the points from data frame `df`", message = FALSE, warning = FALSE----
 library(ggplot2)
 ggplot(df, aes(x1, x2)) +
   geom_text(label = rownames(df)) + 
-  theme(aspect.ratio = 1)
+  coord_fixed(ratio = 1) + 
+  xlim(c(4, 16)) + 
+  ylim(c(4, 16))
 
 
 ## ----fig2, fig.cap = "PaLD graph displaying the relationship between the points in data frame `df`"----
 plot_community_graphs(cohesion)
 
 
-## ----fig3, fig.cap = "PaLD graph displaying the relationship between the points in data frame `d`, matching the original layout in Figure 1"----
+## ----fig3, fig.cap = "PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1"----
 par(pty = "s")
 
 plot_community_graphs(cohesion, 
                       layout = as.matrix(df),
-                      vertex.size = 15,
[…]

### 2022-05-23 r-journal/manuscript.tex
typo

@@
 
-The main functions in \CRANpkg{pald} package can be split into 3
+The main functions in the \CRANpkg{pald} package can be split into 3
 categories:
@@
 \begin{Soutput}
-#> IGRAPH 8eef810 UNW- 6 3 -- 
+#> IGRAPH 62fc4b4 UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from 8eef810 (vertex names):
+#> + edges from 62fc4b4 (vertex names):
 #> [1] A--B A--E D--F
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from 9c17296:
+#> + 8/87 vertices, named, from 8ad41e8:
 #> [1] Italian         Ladin           Provencal       Walloon        

### 2022-05-23 r-journal/manuscript.Rmd
typo
update paper

@@
 abstract: >
-  An abstract of less than 150 words.
+  Partitioned Local Depths (PaLD) is a framework for a holistic consideration of the community structure of distance-based data.
+This paper describes an R package, \CRANpkg{pald}, for calculating Partitioned Local Depths (PaLD) probabilities, implementing clustering analyses, and creating data visualizations to display community structure.
+We describe how to use the package as well as walk through several examples.
 output:
@@
 We present a new package, \CRANpkg{pald}, for calculating Partitioned Local Depths (PaLD) probabilities, implementing clustering analyses, and creating data visualizations to display community structure.
-This paper will describe how to use the package as well as walk through two examples.
+This paper will describe how to use the package as well as walk through several examples.
 
@@
 
-The main functions in \CRANpkg{pald} package can be split into 3 categories:
+The main functions in the \CRANpkg{pald} package can be split into 3 categories:
 

### 2022-05-23 r-journal/manuscript.tex
update paper

@@
 \abstract{%
-An abstract of less than 150 words.
+Partitioned Local Depths (PaLD) is a framework for a holistic
+consideration of the community structure of distance-based data. This
+paper describes an R package, \CRANpkg{pald}, for calculating
+Partitioned Local Depths (PaLD) probabilities, implementing clustering
+analyses, and creating data visualizations to display community
+structure. We describe how to use the package as well as walk through
+several examples.
 }
@@
 creating data visualizations to display community structure. This paper
-will describe how to use the package as well as walk through two
+will describe how to use the package as well as walk through several
 examples.
@@
 \begin{Soutput}
-#> IGRAPH 62fc4b4 UNW- 6 3 -- 
+#> IGRAPH 8cc278d UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from 62fc4b4 (vertex names):
+#> + edges from 8cc278d (vertex names):
 #> [1] A--B A--E D--F
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from 8ad41e8:
+#> + 8/87 vertices, named, from 7044f2e:
 #> [1] Italian         Ladin           Provencal       Walloon        

### 2022-06-02 r-journal/dagostino-mcgowan.Rmd
update paper

--- 2022-05-23 r-journal/manuscript.Rmd
@@
 ---
-title: "Partitioned Local Depth (PaLD) Clustering Analyses in R"
+title: "Partitioned Local Depth (PaLD) Community Cluster Analyses in R"
 author:
@@
 abstract: >
-  Partitioned Local Depths (PaLD) is a framework for a holistic consideration of the community structure of distance-based data.
-This paper describes an R package, \CRANpkg{pald}, for calculating Partitioned Local Depths (PaLD) probabilities, implementing clustering analyses, and creating data visualizations to display community structure.
+  Partitioned Local Depth (PaLD) is a framework for a holistic consideration of the community structure of distance-based data.
+This paper describes an R package, \CRANpkg{pald}, for calculating Partitioned Local Depth (PaLD) probabilities, implementing community clustering analyses, and creating data visualizations to display community structure.
 We describe how to use the package as well as walk through several examples.
-output:
-  rticles::rjournal_article:
-    includes:
-      in_header: preamble.tex
+output: rticles::rjournal_article
 ---
@@
 
-Partitioned Local Depths (PaLD) is a framework for a holistic consideration of the community structure of distance-based data.
+Partitioned Local Depth (PaLD) is a framework for a holistic consideration of community structure for distance-based data.
 Leveraging a socially inspired perspective, the method provides network-based community information which is founded on a new measure of local depth and pairwise cohesion (partitioned local depth).
@@
 
-Building on existing approaches to (global) depth, local depth expresses features of centrality as an interpretable probability which is free of parameters and robust to outliers.
+Building on existing approaches to (global) depth, local depth expresses features of centrality via an interpretable probability which is free of parameters and robust to outliers.
 Then, partitioning the probability which defines local depth, we obtain a measure of cohesion between pairs of points.
@@
 Specifically, as shown in \citet{berenhaut2022social}, provided that two sets are separated (in the sense that the minimum between-set distance is greater than the maximum within-set distance), cohesion is invariant under the contraction and dilation of the distances within each set.
-This property may be particularly valuable when one has reason to believe that the magnitude of distances varies across the space.
+This property may be particularly valuable when one has reason to believe that there is heterogeneity in density across the space.
 
 As cohesion captures a sense of the relationship strength between points, we can then visualize the resulting community structure with a network whose edges are weighted by (mutual) cohesion.
-The underlying social framework motivates an elegant threshold for distinguishing between strongly and weakly cohesive pairs.
+The underlying social framework motivates a straightforward yet elegant threshold for distinguishing between strongly and weakly cohesive pairs.
 
-As seen throughout this paper, we display the network obtained from cohesion using a force-directed graph drawing algorithm and emphasize the strong ties (colored by connected component).
+Throughout this paper, we will display the network obtained from cohesion using a force-directed graph drawing algorithm and emphasize the strong ties (colored by connected component).
 We refer to the connected components of the network of strong ties as community "clusters."  Note that to qualify as a cluster in this definition, one may not have any strong ties with those outside the cluster, and thus the existence of disjoint groups is a strong signal for separation.
 Here, clusters are identified without additional user inputs nor optimization criteria.
-If one wishes to further break the community graph into groups, one may use community detection methods for networks (such as spectral clustering or the Louvain algorithm).
+If one wishes to further break the community graph into groups, one may consider using community detection methods (such as  spectral clustering or the Louvain algorithm), as available, say, in the \CRANpkg{igraph} package.
 Though only briefly considered here, one may also use the collection of strong ties in place of (weighted) k-nearest neighbors in settings such as classification and smoothing.
@@
 
-We present a new package, \CRANpkg{pald}, for calculating Partitioned Local Depths (PaLD) probabilities, implementing clustering analyses, and creating data visualizations to display community structure.
+We present a new package, \CRANpkg{pald}, for calculating Partitioned Local Depth (PaLD) probabilities, implementing community clustering analyses, and creating data visualizations to display community structure.
 This paper will describe how to use the package as well as walk through several examples.
@@
 1. A function for computing the cohesion matrix
-2. Functions for extracting useful information from the cohesion matrix, such as local depths, neighbors, clusters, and graph objects
+2. Functions for extracting useful information from the cohesion matrix, such as local depths, neighbors, community clusters, and graph objects
 3. Plotting functions for community graphs
@@
 
-The input for the Partitioned Local Depths (PaLD) is a distance matrix or `dist` object.
-Note that the collection of input distances (or dissimilarities) does not need to satisfy the triangle inequality.
+The input for the Partitioned Local Depths (PaLD) computations is a distance matrix or `dist` object.
+Note that the collection of input distances (or dissimilarities) does not need to satisfy the triangle inequality nor be symmetric.
 
 For demonstration purposes, we will show how one can compute a distance matrix from an input data frame with, say, two variables `x1` and `x2`.
-The input data may be of any dimension.
-The methods put forth here work on data frames with higher dimensions, as described in the **Examples** section; we are simply choosing a small data frame here for demonstration purposes.
+The input data may be of any dimension; in fact the PaLD framework provides advantages when considering high-dimensional data (see the **Examples** section as well as \citet{berenhaut2022social}).
 
@@
 If the data are already provided as a distance matrix (or `dist` object), the user can skip to the next step.
+Note that the distance matrix needed for the subsequent functions doesn't need to be a `dist` object and *need not* be symmetric.
 
@@
 
-From the _cohesion matrix_, a variety of useful quantities can be calculated. Below, we create a cohesion matrix using the functions described in the previous section.
+From the _cohesion matrix_, a variety of useful quantities can be computed. Below, we create a cohesion matrix using the functions described in the previous section.
 
@@
 
-The `local_depths()` function calculates the _depths_ of each point, outputting a vector of local depths.
-Local depth is an interpretable probability which reflects aspects of relative position and centrality via distance comparisons (i.e., $d(z, x) < d(z, y)$).
+The `local_depths()` function calculates the _depth_ of each point, outputting a vector of local depths.
+Local depth is an interpretable probability which reflects aspects of relative position and centrality via distance comparisons (i.e., $d(z, x) < d(z, y)$, see \citet{berenhaut2022social}).
 
@@
 The `strong_threshold()` function will calculate the cohesion threshold for strong ties.
-This is equal to half the average of the diagonal cohesion matrix.\citep{berenhaut2022social} This is a threshold that may be used to distinguish between strong and weak ties.
+This is equal to half the average of the diagonal of the cohesion matrix \citep{berenhaut2022social}, and is a threshold that may be used to distinguish between strong and weak ties.
 
@@
 
-In this case, the threshold is `0.117`.
+In this case, the threshold is a little above `0.117`.
 
@@
 * `G`: the weighted (community) graph whose edge weights are mutual cohesion
-* `G_strong`: the weighted (community) graph consisting of edges for which mutual cohesion is greater than the threshold for strong ties
+* `G_strong`: the weighted (community) graph consisting of edges for which mutual (symmetrized) cohesion  (i.e.
+the minimum of the two directed cohesion values for any given pair) is greater than the threshold for strong ties
 * `layout`: the graph layout, using the Fruchterman Reingold (FR) force-directed graph drawing for the graph `G`
@@
 
-Here we see that there are three connected components, ties `A-B` and  and `A-E` which form the first cluster, and the tie `D-F` which forms another.
+Here we see that there are three connected components, ties `A-B` and `A-E` which form the first cluster, and the tie `D-F` which forms another.
 
-The `any_isolated()` function will check whether there are any isolated points that will inadvertently be dropped by a graph.
+The `any_isolated()` function will check whether there are any isolated points.
 
@@
 
-Here, there are no isolated points.
+\noindent Here, there are no isolated points, i.e. points having zero cohesion with all other points in the data (an extreme form of outlier).
 
-The "clusters" identified by PaLD are the connected components of the graph of strong ties, `G_strong`.
+The "community clusters" identified by PaLD are the connected components of the graph of strong ties, `G_strong`.
 To directly calculate them, we can use the `community_clusters()` function.
@@
 
-In this example, three clusters are identified with these six points. Points `A`, `B`, and `E` fall into cluster 1. Point `C` is in cluster 2 and points `D` and `F` fall into cluster 3.
+In this example, three clusters are identified with these six points.
+Points `A`, `B`, and `E` fall into cluster 1.
+Point `C` is in cluster 2 (a cluster of size 1) and points `D` and `F` fall into cluster 3.
 
@@
 
-The final category of function is functions for data visualization.
-We can begin by visualizing the points in data frame `df` (Figure \ref{fig:fig1}).
+The final category of function is that for data visualization.
+We can begin by visualizing the points in the data frame `df` (Figure \ref{fig:fig1}).
 When visualizing these points, it is important to have the aspect ratio of the x and y axes equal to 1 so as to not distort the distances.
-When using the \CRANpkg{ggplot2} package for this visualization, you can use the `coord_fixed(ratio = 1)` function.
-If using the `plot()` function included in the base library, you can use the `asp = 1` argument.
+When using the \CRANpkg{ggplot2} package for this visualization, one can use the `coord_fixed(ratio = 1)` function.
+If using the `plot()` function included in the base library, one can use the `asp = 1` argument.
 
-```{r fig1, fig.cap = "Visualize the points from data frame `df`", message = FALSE, warning = FALSE}
+```{r fig1, fig.cap = "Visualization of the points from the data frame `df`", message = FALSE, warning = FALSE}
 library(ggplot2)
@@
 We can pass the cohesion matrix to the `plot_community_graphs()` function to view the relationship between points (Figure \ref{fig:fig2}).
+Notice in this plot the layout does not match that of the original data frame as seen in Figure \ref{fig:fig1}.
+Since our original data is two dimensional, it may be reasonable to use this as the layout.
+Figure \ref{fig:fig3} will make this update as well as update some of the aesthetics, such as making the labels more readable.
 
@@
 The `layout` argument allows the user to pass a matrix to dictate the 2-dimensional layout of the graph.
-For example, if we wanted the graph to match the visualization displayed in Figure \ref{fig:fig1}, we can pass `as.matrix(df)`, or a matrix of the data frame `df` to the `layout` argument  (Figure \ref{fig:fig3}).
-Additionally, this `plot_community_graphs()` function will also permit parameters that can be passed to the `plot.igraph()` function.
-For example, we can pass arguments to the `plot.igraph` function via `...`, for example to increase the vertex size and change the vertex label color, we can specify `vertex.size = 100` and `vertex.label.color = "white"`.
-Additionally, to allow axes, we use `axes = TRUE` and to put them back on the original scale we set `rescale = FALSE`, resetting the axis limits using `xlim` and `ylim`.
+For example, if we wanted the graph to match the visualization displayed in Figure \ref{fig:fig1}, we could pass `as.matrix(df)`, or a matrix of the data frame `df` to the `layout` argument  (Figure \ref{fig:fig3}).
+Additionally, the `plot_community_graphs()` function will also permit parameters that can be passed to the `plot.igraph()` function.
+We can pass arguments to the `plot.igraph` function via the `...` argument; for example to increase the vertex size and change the vertex label color, we can specify `vertex.size = 100` and `vertex.label.color = "white"`.
+Additionally, to allow axes, we use `axes = TRUE`, and to put them back on the original scale we set `rescale = FALSE`, resetting the axis limits using `xlim` and `ylim`.
 The `par(pty = "s")` function forces the subsequent plot to be square.
@@
 
-The first example is from a subset of tissue gene expression data from \citet{zilliox2007gene}, \citet{mccall2011gene}, and \citet{mccall2014gene}, obtained from the **tissuesGeneExpression** bioconductor package \citep{tissue}.
-A `dist` object was created using this data set and is included the \CRANpkg{pald} package in an object called `tissue_dist`.
+The first example is from a subset of data from \citet{zilliox2007gene}, \citet{mccall2011gene}, and \citet{mccall2014gene}, obtained from the **tissuesGeneExpression** bioconductor package \citep{tissue} consisting of 22,215-dimensional gene expression data from 189 tissue samples.
+A `dist` object was created using this data set and is included in the \CRANpkg{pald} package in an object called `tissue_dist`.
 
@@
 
-The `community_clusters()` function can be used to identify the clusters of each tissue sample.
+The `community_clusters()` function can be used to identify the community cluster corresponding to each tissue sample.
 Since the output is a data frame, we can summarize the clusters using commonly used data analysis techniques.
@@
 
-From this, we can glean that cluster one consists of two types of tissue, the kidney and endometrium. Cluster two is comprised of only the hippocampus.
+From this, we can glean that cluster 1 consists of two types of tissue, the kidney and endometrium. Cluster 2 is comprised of only the hippocampus.
 
@@
 For clarity of the display, we show how to remove the labels using `show_labels = FALSE`.
-We will instead color by the labels by passing these to the `vertex.color` parameter through the  `...` to the `plot.igraph` function.
+We will instead color by the labels by passing these to the `vertex.color` parameter to the `plot.igraph` function (via the `...` argument).
 Similarly, we can add a legend using the `legend()` function, as you would for an \CRANpkg{igraph} visualization.
@@
 
-This example performs a PaLD analysis on a data set from \citet{dyen92} that examines the relationship between 87 Indo-European languages from the perspective of cognates.
-A `dist` object was created from this data set and included in the \CRANpkg{pald} package in an object called `cognate_dist`.
+This example performs a PaLD analysis on a data set from \citet{dyen92} that examines the relationship between 87 Indo-European languages from the perspective of cognates, coded using 2,655-dimensional binary vectors.
+A `dist` object was created from this data set and is included in the \CRANpkg{pald} package in an object called `cognate_dist`.
 
 This example will demonstrate how you can apply functions in the \CRANpkg{igraph} package to objects output from the \CRANpkg{pald} package.
-We can first use the `cohesion_matrix()` function to calculate the cohesion matrix and the `community_graphs()` function to create a list with the weighted community graph, the weighted community graph with only strong ties included (and all others set to 0), and the layout.
+We can first use the `cohesion_matrix()` function to calculate the cohesion matrix and the `community_graphs()` function to create a list with the weighted community graph, the weighted community graph with only strong ties included, and the layout.
 From this, we can extract the graph with only the strong ties, here called `cognate_graph_strong`.
@@
 We can then use the `neighbors()` function from the \CRANpkg{igraph} package to extract the strong neighbors in this graph.
-For example, if we wanted to extract all neighbors where the language is "French", we would run the following.
+For example, if we wanted to extract all neighbors for the language "French", we could run the following.
 
@@
 
+We can again use the `plot_community_graphs()` function to visualize the community clusters (Figure \ref{fig:figlang}).
+One may note the commonly identifiable language clusters and that, under a slight rotation, some of the underlying geography is mirrored in the plot.
+
+```{r, eval = FALSE}
+plot_community_graphs(
+  cognate_cohesion,
+  edge_width_factor = 30,
+  emph_strong = 3,
+  vertex.size = 3,
+  vertex.label.cex = 0.7,
+  vertex.label.dist = 1
+)
+```
+
+```{r, include = FALSE, output = "hide"}
+set.seed(7)
+png("fig6.png", res = 300, height = 3000, width = 3000)
+plot_community_graphs(
+  cognate_cohesion,
+  edge_width_factor = 30,
+  emph_strong = 3,
+  vertex.label.cex = 0.7,
+  vertex.size = 3,
+  vertex.label.dist = 1
+)
+dev.off()
+```
+
+```{r figlang, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Community structure for 87 Indo-European languages, which employs cognate information that was coded via 2,665-dimensional binary vectors.
+Commonly identifiable language clusters arise along with informative inter- and intra-cluster structure.
+Several ancient languages are centrally located.", out.width = "100%"}
+knitr::include_graphics("fig6.png")
+```
+
 ### Clustering generated data
 
-The \CRANpkg{pald} package includes three randomly generated data frames corresponding to plots from @berenhaut2022social:
+The \CRANpkg{pald} package includes three randomly generated data frames corresponding to plots from \citet{berenhaut2022social}:
 
-* `exdata1` is a data set consisting of 8 points to recreate Figure 1 in @berenhaut2022social 
-* `exdata2` is a data set consisting of 16 points to recreate Figure 2 in @berenhaut2022social 
-* `exdata3` is a data set consisting of 240 points to recreate Figure 4D in @berenhaut2022social
+* `exdata1` is a data set consisting of 8 points to recreate Figure 1 in \citet{berenhaut2022social} 
+* `exdata2` is a data set consisting of 16 points to recreate Figure 2 in \citet{berenhaut2022social} 
+* `exdata3` is a data set consisting of 240 points to recreate Figure 4D in \citet{berenhaut2022social}
 
@@
 
-Cohesion is particularly useful when considering data with varying local density, see discussion in [@berenhaut2022social].
+Cohesion is particularly useful when considering data with varying local density, see discussion in \citet{berenhaut2022social}.
 Note that the PaLD algorithm is able to detect the eight natural groups within the data without the use of any additional inputs (e.g., number of clusters) nor optimization criteria.
-Despite providing the "correct" number of clusters (i.e., $k = 8$) both *k*-means and hierarchical clustering did not give the desired result.
+Despite the user input of the "correct" number of clusters (i.e., $k = 8$) both *k*-means and hierarchical clustering did not give the desired result.
 

### 2022-06-02 r-journal/dagostino-mcgowan.tex
update paper

--- 2022-05-23 r-journal/manuscript.tex
@@
 % !TeX root = RJwrapper.tex
-\title{Partitioned Local Depth (PaLD) Clustering Analyses in R}
+\title{Partitioned Local Depth (PaLD) Community Cluster Analyses in R}
 \author{by Lucy D'Agostino McGowan, Katherine Moore, and Kenneth Berenhaut}
@@
 \abstract{%
-An abstract of less than 150 words.
+Partitioned Local Depth (PaLD) is a framework for a holistic
+consideration of the community structure of distance-based data. This
+paper describes an R package, \CRANpkg{pald}, for calculating
+Partitioned Local Depth (PaLD) probabilities, implementing community
+clustering analyses, and creating data visualizations to display
+community structure. We describe how to use the package as well as walk
+through several examples.
 }
@@
 
-Partitioned Local Depths (PaLD) is a framework for a holistic
-consideration of the community structure of distance-based data.
-Leveraging a socially inspired perspective, the method provides
-network-based community information which is founded on a new measure of
-local depth and pairwise cohesion (partitioned local depth). The method
-does not require distributional assumptions, optimization criteria, nor
+Partitioned Local Depth (PaLD) is a framework for a holistic
+consideration of community structure for distance-based data. Leveraging
+a socially inspired perspective, the method provides network-based
+community information which is founded on a new measure of local depth
+and pairwise cohesion (partitioned local depth). The method does not
+require distributional assumptions, optimization criteria, nor
 extraneous inputs. A complete description of the perspective, together
@@
 Building on existing approaches to (global) depth, local depth expresses
-features of centrality as an interpretable probability which is free of
+features of centrality via an interpretable probability which is free of
 parameters and robust to outliers. Then, partitioning the probability
@@
 contraction and dilation of the distances within each set. This property
-may be particularly valuable when one has reason to believe that the
-magnitude of distances varies across the space.
+may be particularly valuable when one has reason to believe that there
+is heterogeneity in density across the space.
 
@@
 network whose edges are weighted by (mutual) cohesion. The underlying
-social framework motivates an elegant threshold for distinguishing
-between strongly and weakly cohesive pairs.
+social framework motivates a straightforward yet elegant threshold for
+distinguishing between strongly and weakly cohesive pairs.
 
-As seen throughout this paper, we display the network obtained from
+Throughout this paper, we will display the network obtained from
 cohesion using a force-directed graph drawing algorithm and emphasize
@@
 optimization criteria. If one wishes to further break the community
-graph into groups, one may use community detection methods for networks
-(such as spectral clustering or the Louvain algorithm). Though only
-briefly considered here, one may also use the collection of strong ties
-in place of (weighted) k-nearest neighbors in settings such as
-classification and smoothing. Overall, the structural information
-obtained from local depth, cohesion and community graphs can provide a
-holistic perspective on the data which does not require the use of
-distributional assumptions, optimization criteria nor additional user
-inputs.
+graph into groups, one may consider using community detection methods
+(such as spectral clustering or the Louvain algorithm), as available,
+say, in the \CRANpkg{igraph} package. Though only briefly considered
+here, one may also use the collection of strong ties in place of
+(weighted) k-nearest neighbors in settings such as classification and
+smoothing. Overall, the structural information obtained from local
+depth, cohesion and community graphs can provide a holistic perspective
+on the data which does not require the use of distributional
+assumptions, optimization criteria nor additional user inputs.
 
 We present a new package, \CRANpkg{pald}, for calculating Partitioned
-Local Depths (PaLD) probabilities, implementing clustering analyses, and
-creating data visualizations to display community structure. This paper
-will describe how to use the package as well as walk through two
-examples.
+Local Depth (PaLD) probabilities, implementing community clustering
+analyses, and creating data visualizations to display community
+structure. This paper will describe how to use the package as well as
+walk through several examples.
 
@@
   Functions for extracting useful information from the cohesion matrix,
-  such as local depths, neighbors, clusters, and graph objects
+  such as local depths, neighbors, community clusters, and graph objects
 \item
@@
 
-The input for the Partitioned Local Depths (PaLD) is a distance matrix
-or \texttt{dist} object. Note that the collection of input distances (or
-dissimilarities) does not need to satisfy the triangle inequality.
+The input for the Partitioned Local Depths (PaLD) computations is a
+distance matrix or \texttt{dist} object. Note that the collection of
+input distances (or dissimilarities) does not need to satisfy the
+triangle inequality nor be symmetric.
 
@@
 matrix from an input data frame with, say, two variables \texttt{x1} and
-\texttt{x2}. The input data may be of any dimension. The methods put
-forth here work on data frames with higher dimensions, as described in
-the \textbf{Examples} section; we are simply choosing a small data frame
-here for demonstration purposes.
+\texttt{x2}. The input data may be of any dimension; in fact the PaLD
+framework provides advantages when considering high-dimensional data
+(see the \textbf{Examples} section as well as
+\citet{berenhaut2022social}).
 
@@
 as a distance matrix (or \texttt{dist} object), the user can skip to the
-next step.
+next step. Note that the distance matrix needed for the subsequent
+functions doesn't need to be a \texttt{dist} object and \emph{need not}
+be symmetric.
 
@@
 From the \emph{cohesion matrix}, a variety of useful quantities can be
-calculated. Below, we create a cohesion matrix using the functions
+computed. Below, we create a cohesion matrix using the functions
 described in the previous section.
@@
 
-The \texttt{local\_depths()} function calculates the \emph{depths} of
+The \texttt{local\_depths()} function calculates the \emph{depth} of
 each point, outputting a vector of local depths. Local depth is an
 interpretable probability which reflects aspects of relative position
-and centrality via distance comparisons (i.e., \(d(z, x) < d(z, y)\)).
+and centrality via distance comparisons (i.e., \(d(z, x) < d(z, y)\),
+see \citet{berenhaut2022social}).
 
@@
 threshold for strong ties. This is equal to half the average of the
-diagonal cohesion matrix.\citep{berenhaut2022social} This is a threshold
-that may be used to distinguish between strong and weak ties.
+diagonal of the cohesion matrix \citep{berenhaut2022social}, and is a
+threshold that may be used to distinguish between strong and weak ties.
 
@@
 
-In this case, the threshold is \texttt{0.117}.
+In this case, the threshold is a little above \texttt{0.117}.
 
@@
   \texttt{G\_strong}: the weighted (community) graph consisting of edges
-  for which mutual cohesion is greater than the threshold for strong
-  ties
+  for which mutual (symmetrized) cohesion (i.e.~the minimum of the two
+  directed cohesion values for any given pair) is greater than the
+  threshold for strong ties
 \item
@@
 \begin{Soutput}
-#> IGRAPH 62fc4b4 UNW- 6 3 -- 
+#> IGRAPH b2835f9 UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from 62fc4b4 (vertex names):
+#> + edges from b2835f9 (vertex names):
 #> [1] A--B A--E D--F
@@
 Here we see that there are three connected components, ties \texttt{A-B}
-and and \texttt{A-E} which form the first cluster, and the tie
-\texttt{D-F} which forms another.
+and \texttt{A-E} which form the first cluster, and the tie \texttt{D-F}
+which forms another.
 
 The \texttt{any\_isolated()} function will check whether there are any
-isolated points that will inadvertently be dropped by a graph.
+isolated points.
 
@@
 
-Here, there are no isolated points.
+\noindent Here, there are no isolated points, i.e.~points having zero
+cohesion with all other points in the data (an extreme form of outlier).
 
-The ``clusters'' identified by PaLD are the connected components of the
-graph of strong ties, \texttt{G\_strong}. To directly calculate them, we
-can use the \texttt{community\_clusters()} function. This will output a
-data frame with two columns, the first will correspond to the
-\texttt{point}, as identified by the row name of the original input data
-frame, \texttt{df}, the second will identify the \texttt{cluster} that
-each point belongs to.
+The ``community clusters'' identified by PaLD are the connected
+components of the graph of strong ties, \texttt{G\_strong}. To directly
+calculate them, we can use the \texttt{community\_clusters()} function.
+This will output a data frame with two columns, the first will
+correspond to the \texttt{point}, as identified by the row name of the
+original input data frame, \texttt{df}, the second will identify the
+\texttt{cluster} that each point belongs to.
 
@@
 Points \texttt{A}, \texttt{B}, and \texttt{E} fall into cluster 1. Point
-\texttt{C} is in cluster 2 and points \texttt{D} and \texttt{F} fall
-into cluster 3.
+\texttt{C} is in cluster 2 (a cluster of size 1) and points \texttt{D}
+and \texttt{F} fall into cluster 3.
 
@@
 
-The final category of function is functions for data visualization. We
-can begin by visualizing the points in data frame \texttt{df} (Figure
+The final category of function is that for data visualization. We can
+begin by visualizing the points in the data frame \texttt{df} (Figure
 \ref{fig:fig1}). When visualizing these points, it is important to have
@@
 distances. When using the \CRANpkg{ggplot2} package for this
-visualization, you can use the \texttt{coord\_fixed(ratio\ =\ 1)}
+visualization, one can use the \texttt{coord\_fixed(ratio\ =\ 1)}
 function. If using the \texttt{plot()} function included in the base
-library, you can use the \texttt{asp\ =\ 1} argument.
+library, one can use the \texttt{asp\ =\ 1} argument.
 
@@
 \begin{figure}
-\includegraphics{manuscript_files/figure-latex/fig1-1} \caption[Visualize the points from data frame `df`]{Visualize the points from data frame `df`}\label{fig:fig1}
+\includegraphics{dagostino-mcgowan_files/figure-latex/fig1-1} \caption[Visualization of the points from the data frame `df`]{Visualization of the points from the data frame `df`}\label{fig:fig1}
 \end{figure}
@@
 \texttt{plot\_community\_graphs()} function to view the relationship
-between points (Figure \ref{fig:fig2}).
+between points (Figure \ref{fig:fig2}). Notice in this plot the layout
+does not match that of the original data frame as seen in Figure
+\ref{fig:fig1}. Since our original data is two dimensional, it may be
+reasonable to use this as the layout. Figure \ref{fig:fig3} will make
+this update as well as update some of the aesthetics, such as making the
+labels more readable.
 
@@
 \begin{figure}
-\includegraphics{manuscript_files/figure-latex/fig2-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`]{PaLD graph displaying the relationship between the points in data frame `df`}\label{fig:fig2}
+\includegraphics{dagostino-mcgowan_files/figure-latex/fig2-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`]{PaLD graph displaying the relationship between the points in data frame `df`}\label{fig:fig2}
 \end{figure}
@@
 graph to match the visualization displayed in Figure \ref{fig:fig1}, we
-can pass \texttt{as.matrix(df)}, or a matrix of the data frame
+could pass \texttt{as.matrix(df)}, or a matrix of the data frame
 \texttt{df} to the \texttt{layout} argument (Figure \ref{fig:fig3}).
-Additionally, this \texttt{plot\_community\_graphs()} function will also
+Additionally, the \texttt{plot\_community\_graphs()} function will also
 permit parameters that can be passed to the \texttt{plot.igraph()}
-function. For example, we can pass arguments to the \texttt{plot.igraph}
-function via \texttt{...}, for example to increase the vertex size and
+function. We can pass arguments to the \texttt{plot.igraph} function via
+the \texttt{...} argument; for example to increase the vertex size and
 change the vertex label color, we can specify
@@
 \texttt{vertex.label.color\ =\ "white"}. Additionally, to allow axes, we
-use \texttt{axes\ =\ TRUE} and to put them back on the original scale we
-set \texttt{rescale\ =\ FALSE}, resetting the axis limits using
+use \texttt{axes\ =\ TRUE}, and to put them back on the original scale
+we set \texttt{rescale\ =\ FALSE}, resetting the axis limits using
 \texttt{xlim} and \texttt{ylim}. The \texttt{par(pty\ =\ "s")} function
@@
 \begin{figure}
-\includegraphics{manuscript_files/figure-latex/fig3-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1]{PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1}\label{fig:fig3}
+\includegraphics{dagostino-mcgowan_files/figure-latex/fig3-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1]{PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1}\label{fig:fig3}
 \end{figure}
@@
 
-The first example is from a subset of tissue gene expression data from
-\citet{zilliox2007gene}, \citet{mccall2011gene}, and
-\citet{mccall2014gene}, obtained from the \textbf{tissuesGeneExpression}
-bioconductor package \citep{tissue}. A \texttt{dist} object was created
-using this data set and is included the \CRANpkg{pald} package in an
-object called \texttt{tissue\_dist}.
+The first example is from a subset of data from \citet{zilliox2007gene},
+\citet{mccall2011gene}, and \citet{mccall2014gene}, obtained from the
+\textbf{tissuesGeneExpression} bioconductor package \citep{tissue}
+consisting of 22,215-dimensional gene expression data from 189 tissue
+samples. A \texttt{dist} object was created using this data set and is
+included in the \CRANpkg{pald} package in an object called
+\texttt{tissue\_dist}.
 
@@
 The \texttt{community\_clusters()} function can be used to identify the
-clusters of each tissue sample. Since the output is a data frame, we can
-summarize the clusters using commonly used data analysis techniques. For
-demonstration purposes, we will use the \CRANpkg{dplyr} package to
-summarize the contribution of clusters.
+community cluster corresponding to each tissue sample. Since the output
+is a data frame, we can summarize the clusters using commonly used data
+analysis techniques. For demonstration purposes, we will use the
+\CRANpkg{dplyr} package to summarize the contribution of clusters.
 
@@
 
-From this, we can glean that cluster one consists of two types of
-tissue, the kidney and endometrium. Cluster two is comprised of only the
+From this, we can glean that cluster 1 consists of two types of tissue,
+the kidney and endometrium. Cluster 2 is comprised of only the
 hippocampus.
@@
 \texttt{show\_labels\ =\ FALSE}. We will instead color by the labels by
-passing these to the \texttt{vertex.color} parameter through the
-\texttt{...} to the \texttt{plot.igraph} function. Similarly, we can add
-a legend using the \texttt{legend()} function, as you would for an
-\CRANpkg{igraph} visualization. Additionally, we use the
-\texttt{edge\_width\_factor} and \texttt{emph\_strong} arguments to
+passing these to the \texttt{vertex.color} parameter to the
+\texttt{plot.igraph} function (via the \texttt{...} argument).
+Similarly, we can add a legend using the \texttt{legend()} function, as
+you would for an \CRANpkg{igraph} visualization. Additionally, we use
+the \texttt{edge\_width\_factor} and \texttt{emph\_strong} arguments to
 adjust the width of the lines between and within PaLD clusters.
@@
 that examines the relationship between 87 Indo-European languages from
-the perspective of cognates. A \texttt{dist} object was created from
-this data set and included in the \CRANpkg{pald} package in an object
-called \texttt{cognate\_dist}.
+the perspective of cognates, coded using 2,655-dimensional binary
+vectors. A \texttt{dist} object was created from this data set and is
+included in the \CRANpkg{pald} package in an object called
+\texttt{cognate\_dist}.
 
@@
 function to create a list with the weighted community graph, the
-weighted community graph with only strong ties included (and all others
-set to 0), and the layout. From this, we can extract the graph with only
-the strong ties, here called \texttt{cognate\_graph\_strong}.
+weighted community graph with only strong ties included, and the layout.
+From this, we can extract the graph with only the strong ties, here
+called \texttt{cognate\_graph\_strong}.
 
@@
 \CRANpkg{igraph} package to extract the strong neighbors in this graph.
-For example, if we wanted to extract all neighbors where the language is
-``French'', we would run the following.
+For example, if we wanted to extract all neighbors for the language
+``French'', we could run the following.
 
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from 8ad41e8:
+#> + 8/87 vertices, named, from e40f165:
 #> [1] Italian         Ladin           Provencal       Walloon        
@@
 \end{Soutput}
+\end{Schunk}
+
+We can again use the \texttt{plot\_community\_graphs()} function to
+visualize the community clusters (Figure \ref{fig:figlang}). One may
+note the commonly identifiable language clusters and that, under a
+slight rotation, some of the underlying geography is mirrored in the
+plot.
+
+\begin{Schunk}
+\begin{Sinput}
+plot_community_graphs(
+  cognate_cohesion,
+  edge_width_factor = 30,
+  emph_strong = 3,
+  vertex.size = 3,
+  vertex.label.cex = 0.7,
+  vertex.label.dist = 1
+)
+\end{Sinput}
+\end{Schunk}
+
+\begin{Schunk}
+\begin{figure}
+\includegraphics[width=1\linewidth]{fig6} \caption[Community structure for 87 Indo-European languages, which employs cognate information that was coded via 2,665-dimensional binary vectors]{Community structure for 87 Indo-European languages, which employs cognate information that was coded via 2,665-dimensional binary vectors.
+Commonly identifiable language clusters arise along with informative inter- and intra-cluster structure.
+Several ancient languages are centrally located.}\label{fig:figlang}
+\end{figure}
 \end{Schunk}
@@
 \begin{figure}
-\includegraphics{manuscript_files/figure-latex/fig5-1} \caption[PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al]{PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al.
+\includegraphics{dagostino-mcgowan_files/figure-latex/fig5-1} \caption[PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al]{PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al.
 (2022)) compared to k-means and hierarchical clustering with k = 8.}\label{fig:fig5}
@@
 Cohesion is particularly useful when considering data with varying local
-density, see discussion in \citep{berenhaut2022social}. Note that the
+density, see discussion in \citet{berenhaut2022social}. Note that the
 PaLD algorithm is able to detect the eight natural groups within the
 data without the use of any additional inputs (e.g., number of clusters)
-nor optimization criteria. Despite providing the ``correct'' number of
-clusters (i.e., \(k = 8\)) both \emph{k}-means and hierarchical
-clustering did not give the desired result.
+nor optimization criteria. Despite the user input of the ``correct''
+number of clusters (i.e., \(k = 8\)) both \emph{k}-means and
+hierarchical clustering did not give the desired result.
 

### 2022-06-02 r-journal/dagostino-mcgowan.R

@@ -1,0 +1,226 @@
+# Generated by `rjournal_article()` using `knitr::purl()`: do not edit by hand
+# Please edit dagostino-mcgowan.Rmd to modify this file
+
+## ---- echo = FALSE------------------------------------------------------------
+knitr::opts_chunk$set(fig.height = 4, message = FALSE, warning = FALSE)
+
+
+## -----------------------------------------------------------------------------
+library(pald)
+df <- data.frame(
+  x1 = c(6, 8, 8, 16, 4, 14),
+  x2 = c(5, 4, 10, 8, 4, 10)
+)
+rownames(df) <- c("A", "B", "C", "D", "E", "F")
+
+
+## -----------------------------------------------------------------------------
+d <- dist(df)
+
+
+## -----------------------------------------------------------------------------
+d <- dist(df)
+cohesion_matrix(d)
+
+
+## -----------------------------------------------------------------------------
+df |>
+  dist() |>
+  cohesion_matrix()
+
+
+## -----------------------------------------------------------------------------
+df |>
+  dist() |>
+  cohesion_matrix() -> cohesion
+
+
+## -----------------------------------------------------------------------------
+local_depths(cohesion)
+
+
+## -----------------------------------------------------------------------------
+strong_threshold(cohesion)
+
+
+## -----------------------------------------------------------------------------
+cohesion_strong(cohesion)
+
+
+## -----------------------------------------------------------------------------
+graphs <- community_graphs(cohesion)
+graphs[["G_strong"]]
+
+
+## -----------------------------------------------------------------------------
+any_isolated(cohesion)
+
+
+## -----------------------------------------------------------------------------
+community_clusters(cohesion)
+
+
+## ----fig1, fig.cap = "Visualization of the points from the data frame `df`", message = FALSE, warning = FALSE----
+library(ggplot2)
+ggplot(df, aes(x1, x2)) +
+  geom_text(label = rownames(df)) + 
+  coord_fixed(ratio = 1) + 
+  xlim(c(4, 16)) + 
+  ylim(c(4, 16))
+
+
+## ----fig2, fig.cap = "PaLD graph displaying the relationship between the points in data frame `df`"----
+plot_community_graphs(cohesion)
+
+
[…]

### 2022-06-02 r-journal/manuscript.R

@@ -1,197 +1,0 @@
-# Generated by `rjournal_article()` using `knitr::purl()`: do not edit by hand
-# Please edit manuscript.Rmd to modify this file
-
-## ---- echo = FALSE------------------------------------------------------------
-knitr::opts_chunk$set(fig.height = 4, message = FALSE, warning = FALSE)
-
-
-## -----------------------------------------------------------------------------
-library(pald)
-df <- data.frame(
-  x1 = c(6, 8, 8, 16, 4, 14),
-  x2 = c(5, 4, 10, 8, 4, 10)
-)
-rownames(df) <- c("A", "B", "C", "D", "E", "F")
-
-
-## -----------------------------------------------------------------------------
-d <- dist(df)
-
-
-## -----------------------------------------------------------------------------
-d <- dist(df)
-cohesion_matrix(d)
-
-
-## -----------------------------------------------------------------------------
-df |>
-  dist() |>
-  cohesion_matrix()
-
-
-## -----------------------------------------------------------------------------
-df |>
-  dist() |>
-  cohesion_matrix() -> cohesion
-
-
-## -----------------------------------------------------------------------------
-local_depths(cohesion)
-
-
-## -----------------------------------------------------------------------------
-strong_threshold(cohesion)
-
-
-## -----------------------------------------------------------------------------
-cohesion_strong(cohesion)
-
-
-## -----------------------------------------------------------------------------
-graphs <- community_graphs(cohesion)
-graphs[["G_strong"]]
-
-
-## -----------------------------------------------------------------------------
-any_isolated(cohesion)
-
-
-## -----------------------------------------------------------------------------
-community_clusters(cohesion)
-
-
-## ----fig1, fig.cap = "Visualize the points from data frame `df`", message = FALSE, warning = FALSE----
-library(ggplot2)
-ggplot(df, aes(x1, x2)) +
-  geom_text(label = rownames(df)) + 
-  coord_fixed(ratio = 1) + 
-  xlim(c(4, 16)) + 
-  ylim(c(4, 16))
-
-
-## ----fig2, fig.cap = "PaLD graph displaying the relationship between the points in data frame `df`"----
-plot_community_graphs(cohesion)
-
-
-## ----fig3, fig.cap = "PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1"----
-par(pty = "s")
-
[…]

### 2022-06-02 r-journal/dagostino-mcgowan.Rmd
add cultures data

@@
 
-In addition, the package provides a number of pertinent example data sets commonly used to demonstrate cluster algorithms, including a synthetic data set of two-dimensional points created by \citet{gionis1clustering} to demonstrate clustering aggregation, clustering data generated from the scikit-learn Python package \citep{pedregosa2011scikit}, data describing cognate relationships between words across 87 Indo-European languages \citep{dyen92}, data compiled by \cite{tissue} of tissue gene expressions, and three example data sets generated for the \citet{berenhaut2022social} paper.
+In addition, the package provides a number of pertinent example data sets commonly used to demonstrate cluster algorithms, including a synthetic data set of two-dimensional points created by \citet{gionis1clustering} to demonstrate clustering aggregation, clustering data generated from the scikit-learn Python package \citep{pedregosa2011scikit}, data describing cognate relationships between words across 87 Indo-European languages \citep{dyen92}, data compiled by \cite{tissue} of tissue gene expressions, data from the World Values Survey \citep{inglehart2014world} on cultural values regarding family, religion, education, and institutions for several regions \citep{muthukrishna2020beyond}, and three example data sets generated for the \citet{berenhaut2022social} paper.
 

### 2022-06-02 r-journal/dagostino-mcgowan.tex
add cultures data

@@
 between words across 87 Indo-European languages \citep{dyen92}, data
-compiled by \cite{tissue} of tissue gene expressions, and three example
-data sets generated for the \citet{berenhaut2022social} paper.
+compiled by \cite{tissue} of tissue gene expressions, data from the
+World Values Survey \citep{inglehart2014world} on cultural values
+regarding family, religion, education, and institutions for several
+regions \citep{muthukrishna2020beyond}, and three example data sets
+generated for the \citet{berenhaut2022social} paper.
 
@@
 \begin{Soutput}
-#> IGRAPH b2835f9 UNW- 6 3 -- 
+#> IGRAPH 4190215 UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from b2835f9 (vertex names):
+#> + edges from 4190215 (vertex names):
 #> [1] A--B A--E D--F
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from e40f165:
+#> + 8/87 vertices, named, from 4db266d:
 #> [1] Italian         Ladin           Provencal       Walloon        

### 2023-05-02 r-journal/dagostino-mcgowan.Rmd
update with new

@@
 ---
-title: "Partitioned Local Depth (PaLD) Community Cluster Analyses in R"
+title: "Partitioned Local Depth (PaLD) Community Analyses in R"
 author:
@@
   Partitioned Local Depth (PaLD) is a framework for a holistic consideration of the community structure of distance-based data.
-This paper describes an R package, \CRANpkg{pald}, for calculating Partitioned Local Depth (PaLD) probabilities, implementing community clustering analyses, and creating data visualizations to display community structure.
+This paper describes an R package, \CRANpkg{pald}, for calculating Partitioned Local Depth (PaLD) probabilities, implementing community analyses, and creating data visualizations to display community structure.
 We describe how to use the package as well as walk through several examples.
@@
 
-We present a new package, \CRANpkg{pald}, for calculating Partitioned Local Depth (PaLD) probabilities, implementing community clustering analyses, and creating data visualizations to display community structure.
-This paper will describe how to use the package as well as walk through several examples.
+We present a new package, \CRANpkg{pald}, for calculating Partitioned Local Depth (PaLD) probabilities, implementing community analyses, and creating data visualizations to display community structure.
+This paper will describe how to use the package, walk through several examples, and compare the method to commonly used techniques in R.
+Together, these demonstrate both the novelty of the method and utility of the implementation in package described.
 
@@
 The function above creates a `dist` object.
-If converted to a matrix, this will be a $n\times n$ distance matrix, where $n$ corresponds to the number of observations in the original data frame, in this example $n = 6$.
+If converted to a matrix, this will be an $n\times n$ distance matrix, where $n$ corresponds to the number of observations in the original data frame, in this example $n = 6$.
 
@@
 We can begin by visualizing the points in the data frame `df` (Figure \ref{fig:fig1}).
-When visualizing these points, it is important to have the aspect ratio of the x and y axes equal to 1 so as to not distort the distances.
+When visualizing these points, it is important to have the aspect ratio of the x and y axes equal to 1 so as to not distort distances.
 When using the \CRANpkg{ggplot2} package for this visualization, one can use the `coord_fixed(ratio = 1)` function.
@@
 Since our original data is two dimensional, it may be reasonable to use this as the layout.
-Figure \ref{fig:fig3} will make this update as well as update some of the aesthetics, such as making the labels more readable.
+Figure \ref{fig:fig3} will make this update as well as update some of the aesthetics, such as employing more readable labels.
 
@@
 
-We will demonstrate the utility of the \CRANpkg{pald} package in three clustering examples.
+We will demonstrate the utility of the \CRANpkg{pald} package in three illustrative examples.
 
@@
 The `community_clusters()` function can be used to identify the community cluster corresponding to each tissue sample.
-Since the output is a data frame, we can summarize the clusters using commonly used data analysis techniques.
+Since the output is a data frame, we can summarize the result using commonly employed data analysis techniques.
 For demonstration purposes, we will use the \CRANpkg{dplyr} package to summarize the contribution of clusters.
@@
 
-### Clustering generated data
+### Community analysis for generated data
 
@@
 
-Here, we will demonstrate how to use `exdata3`. These points were generated from bivariate normal distributions with varying means and variances. There are eight "true" clusters.
+Here, we will demonstrate how to use `exdata3`. These points were generated from bivariate normal distributions with varying means and variances. There are eight "true" communities.
 
@@
 Despite the user input of the "correct" number of clusters (i.e., $k = 8$) both *k*-means and hierarchical clustering did not give the desired result.
+For further examples, discussion, and theoretical results, see \citet{berenhaut2022social}.
 
@@
 This paper introduces the \CRANpkg{pald} package, demonstrating its utility for providing parameter-free clustering which can easily be implemented for a variety of data sets.
+We provide example code as well as compare the method to commonly used clustering techniques, *k*-means and hierarchical clustering.
 

### 2023-05-02 r-journal/dagostino-mcgowan.tex
update with new

@@
 % !TeX root = RJwrapper.tex
-\title{Partitioned Local Depth (PaLD) Community Cluster Analyses in R}
+\title{Partitioned Local Depth (PaLD) Community Analyses in R}
 \author{by Lucy D'Agostino McGowan, Katherine Moore, and Kenneth Berenhaut}
@@
 Partitioned Local Depth (PaLD) probabilities, implementing community
-clustering analyses, and creating data visualizations to display
-community structure. We describe how to use the package as well as walk
-through several examples.
+analyses, and creating data visualizations to display community
+structure. We describe how to use the package as well as walk through
+several examples.
 }
@@
 We present a new package, \CRANpkg{pald}, for calculating Partitioned
-Local Depth (PaLD) probabilities, implementing community clustering
-analyses, and creating data visualizations to display community
-structure. This paper will describe how to use the package as well as
-walk through several examples.
+Local Depth (PaLD) probabilities, implementing community analyses, and
+creating data visualizations to display community structure. This paper
+will describe how to use the package, walk through several examples, and
+compare the method to commonly used clustering algorithms.
 
@@
 The function above creates a \texttt{dist} object. If converted to a
-matrix, this will be a \(n\times n\) distance matrix, where \(n\)
+matrix, this will be an \(n\times n\) distance matrix, where \(n\)
 corresponds to the number of observations in the original data frame, in
@@
 \begin{Soutput}
-#> IGRAPH 4190215 UNW- 6 3 -- 
+#> IGRAPH aa6899c UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from 4190215 (vertex names):
+#> + edges from aa6899c (vertex names):
 #> [1] A--B A--E D--F
@@
 \ref{fig:fig1}). When visualizing these points, it is important to have
-the aspect ratio of the x and y axes equal to 1 so as to not distort the
+the aspect ratio of the x and y axes equal to 1 so as to not distort
 distances. When using the \CRANpkg{ggplot2} package for this
@@
 reasonable to use this as the layout. Figure \ref{fig:fig3} will make
-this update as well as update some of the aesthetics, such as making the
-labels more readable.
+this update as well as update some of the aesthetics, such as employing
+more readable labels.
 
@@
 We will demonstrate the utility of the \CRANpkg{pald} package in three
-clustering examples.
+illustrative examples.
 
@@
 community cluster corresponding to each tissue sample. Since the output
-is a data frame, we can summarize the clusters using commonly used data
-analysis techniques. For demonstration purposes, we will use the
+is a data frame, we can summarize the result using commonly employed
+data analysis techniques. For demonstration purposes, we will use the
 \CRANpkg{dplyr} package to summarize the contribution of clusters.
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from 4db266d:
+#> + 8/87 vertices, named, from 7ab7e40:
 #> [1] Italian         Ladin           Provencal       Walloon        
@@
 
-\hypertarget{clustering-generated-data}{%
-\subsubsection{Clustering generated
-data}\label{clustering-generated-data}}
+\hypertarget{community-analysis-for-generated-data}{%
+\subsubsection{Community analysis for generated
+data}\label{community-analysis-for-generated-data}}
 
@@
 generated from bivariate normal distributions with varying means and
-variances. There are eight ``true'' clusters.
+variances. There are eight ``true'' communities.
 
@@
 number of clusters (i.e., \(k = 8\)) both \emph{k}-means and
-hierarchical clustering did not give the desired result.
+hierarchical clustering did not give the desired result. For further
+examples, discussion, and theoretical results, see
+\citet{berenhaut2022social}.
 
@@
 utility for providing parameter-free clustering which can easily be
-implemented for a variety of data sets.
+implemented for a variety of data sets. We provide example code as well
+as compare the method to commonly used clustering techniques,
+\emph{k}-means and hierarchical clustering.
 

### 2023-05-02 r-journal/dagostino-mcgowan.tex
add sentence

@@
 will describe how to use the package, walk through several examples, and
-compare the method to commonly used clustering algorithms.
+compare the method to commonly used techniques in R. Together, these
+demonstrate both the novelty of the method and utility of the
+implementation in package described.
 
@@
 \begin{Soutput}
-#> IGRAPH aa6899c UNW- 6 3 -- 
+#> IGRAPH f9306e2 UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from aa6899c (vertex names):
+#> + edges from f9306e2 (vertex names):
 #> [1] A--B A--E D--F
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from 7ab7e40:
+#> + 8/87 vertices, named, from 2660ad9:
 #> [1] Italian         Ladin           Provencal       Walloon        

### 2023-05-02 r-journal/dagostino-mcgowan.tex
update abstract

@@
 analyses, and creating data visualizations to display community
-structure. We describe how to use the package as well as walk through
-several examples.
+structure. We describe how to use the package, walk through several
+examples, and contrast the results with commonly used techniques.
 }
@@
 \begin{Soutput}
-#> IGRAPH f9306e2 UNW- 6 3 -- 
+#> IGRAPH 9daec2b UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from f9306e2 (vertex names):
+#> + edges from 9daec2b (vertex names):
 #> [1] A--B A--E D--F
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from 2660ad9:
+#> + 8/87 vertices, named, from 8f39182:
 #> [1] Italian         Ladin           Provencal       Walloon        

### 2023-05-02 r-journal/dagostino-mcgowan.Rmd
update abstract
capitalize cluster

@@
 This paper describes an R package, \CRANpkg{pald}, for calculating Partitioned Local Depth (PaLD) probabilities, implementing community analyses, and creating data visualizations to display community structure.
-We describe how to use the package as well as walk through several examples.
+We describe how to use the package, walk through several examples, and contrast the results with commonly used techniques.
 output: rticles::rjournal_article
@@
 In this example, three clusters are identified with these six points.
-Points `A`, `B`, and `E` fall into cluster 1.
-Point `C` is in cluster 2 (a cluster of size 1) and points `D` and `F` fall into cluster 3.
+Points `A`, `B`, and `E` fall into Cluster 1.
+Point `C` is in Cluster 2 (a cluster of size 1) and points `D` and `F` fall into Cluster 3.
 
@@
 
-From this, we can glean that cluster 1 consists of two types of tissue, the kidney and endometrium. Cluster 2 is comprised of only the hippocampus.
+From this, we can glean that Cluster 1 consists of two types of tissue, the kidney and endometrium. Cluster 2 is comprised of only the hippocampus.
 

### 2023-05-02 r-journal/dagostino-mcgowan.tex
capitalize cluster

@@
 \begin{Soutput}
-#> IGRAPH 9daec2b UNW- 6 3 -- 
+#> IGRAPH e4c9a8b UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from 9daec2b (vertex names):
+#> + edges from e4c9a8b (vertex names):
 #> [1] A--B A--E D--F
@@
 In this example, three clusters are identified with these six points.
-Points \texttt{A}, \texttt{B}, and \texttt{E} fall into cluster 1. Point
-\texttt{C} is in cluster 2 (a cluster of size 1) and points \texttt{D}
-and \texttt{F} fall into cluster 3.
+Points \texttt{A}, \texttt{B}, and \texttt{E} fall into Cluster 1. Point
+\texttt{C} is in Cluster 2 (a cluster of size 1) and points \texttt{D}
+and \texttt{F} fall into Cluster 3.
 
@@
 
-From this, we can glean that cluster 1 consists of two types of tissue,
+From this, we can glean that Cluster 1 consists of two types of tissue,
 the kidney and endometrium. Cluster 2 is comprised of only the
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from 8f39182:
+#> + 8/87 vertices, named, from 81f58d9:
 #> [1] Italian         Ladin           Provencal       Walloon        

### 2023-05-02 r-journal/dagostino-mcgowan.Rmd
update

@@
 
-Cohesion is particularly useful when considering data with varying local density, see discussion in \citet{berenhaut2022social}.
+Cohesion is particularly useful when considering data with varying local density; see \citet{berenhaut2022social} for further examples, discussion, and theoretical results.
 Note that the PaLD algorithm is able to detect the eight natural groups within the data without the use of any additional inputs (e.g., number of clusters) nor optimization criteria.
 Despite the user input of the "correct" number of clusters (i.e., $k = 8$) both *k*-means and hierarchical clustering did not give the desired result.
-For further examples, discussion, and theoretical results, see \citet{berenhaut2022social}.
+
+### Cultural and Pyschological distance analysis
+
+**TODO intro to data here** \citep{muthukrishna2020beyond}
+
+First, we will create the cohesion matrix using the `cultures` data, provided in \CRANpkg{pald}, and proceed to plot the community structure.
+
+```{r}
+cultures_cohesion <- cohesion_matrix(cultures)
+```
+
+```{r, eval = FALSE}
+plot_community_graphs(
+  cultures_cohesion,
+  edge_width_factor = 30,
+  emph_strong = 3,
+  vertex.label.cex = 0.7,
+  vertex.size = 3,
+  vertex.label.dist = 1
+)
+```
+
+```{r, include = FALSE, results = "hide"}
+set.seed(7)
+png("fig7.png", res = 300, height = 3000, width = 3000)
+plot_community_graphs(
+  cultures_cohesion,
+  edge_width_factor = 30,
+  emph_strong = 3,
+  vertex.label.cex = 0.7,
+  vertex.size = 3,
+  vertex.label.dist = 1
+)
+dev.off()
+```
+
+```{r figculture, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Community structure for cultural distance data.", out.width = "100%"}
+knitr::include_graphics("fig7.png")
+```
+
+This example demonstrates key features of the PaLD approach that distinguish it from commonly used clustering techniques.
+In addition to viewing the community structure as seen in Figure \ref{fig:figculture}, we can also visualize a two dimensional plot of cohesion versus distance for each of the included distance pairs using the `dist_cohesion_plot()` function, as seen below (Figure \ref{fig:figco}).
+
+```{r figco, fig.cap = "Distance versus cohesion for each TODO The identified communties are shown via the 5 colors -- these map to the same clusters idenfitied in Figure 7."}
+dist_cohesion_plot(cultures)
+```
+
+Notice here the magnitude of the distances within each of the identified communities varies substantially between regions, yet the cohesion remains strong.
+In other words, focusing particularly on identified community consisting of regions in the United States (light purple) compared to the community consisting of regions in India (turquoise) the strong cohesions are similar (Figure \ref{fig:figculture}), yet the distances themselves are markedly different (Figure \ref{fig:figco}).
+In fact, the largest distance between regions in the Unites States is smaller than that between any two regions in India (Figure \ref{fig:figco}).
+Classic distance-based clustering techniques tend to be unable to maintain this distinction often breaking India into several smaller communities.
+Additionally, currently available techniques require specification of parameters, as seen in the previous section.
+For further discussion of the cultural distance data in relation to community analysis see \citet{berenhaut2022social}.
 

### 2023-05-02 r-journal/dagostino-mcgowan.tex
update

@@
 \begin{Soutput}
-#> IGRAPH e4c9a8b UNW- 6 3 -- 
+#> IGRAPH 2f2c953 UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from e4c9a8b (vertex names):
+#> + edges from 2f2c953 (vertex names):
 #> [1] A--B A--E D--F
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from 81f58d9:
+#> + 8/87 vertices, named, from 828fab3:
 #> [1] Italian         Ladin           Provencal       Walloon        
@@
 Cohesion is particularly useful when considering data with varying local
-density, see discussion in \citet{berenhaut2022social}. Note that the
-PaLD algorithm is able to detect the eight natural groups within the
-data without the use of any additional inputs (e.g., number of clusters)
-nor optimization criteria. Despite the user input of the ``correct''
-number of clusters (i.e., \(k = 8\)) both \emph{k}-means and
-hierarchical clustering did not give the desired result. For further
-examples, discussion, and theoretical results, see
-\citet{berenhaut2022social}.
+density; see \citet{berenhaut2022social} for further examples,
+discussion, and theoretical results. Note that the PaLD algorithm is
+able to detect the eight natural groups within the data without the use
+of any additional inputs (e.g., number of clusters) nor optimization
+criteria. Despite the user input of the ``correct'' number of clusters
+(i.e., \(k = 8\)) both \emph{k}-means and hierarchical clustering did
+not give the desired result.
+
+\hypertarget{cultural-and-pyschological-distance-analysis}{%
+\subsubsection{Cultural and Pyschological distance
+analysis}\label{cultural-and-pyschological-distance-analysis}}
+
+\textbf{TODO intro to data here} \citep{muthukrishna2020beyond}
+
+First, we will create the cohesion matrix using the \texttt{cultures}
+data, provided in \CRANpkg{pald}, and proceed to plot the community
+structure.
+
+\begin{Schunk}
+\begin{Sinput}
+cultures_cohesion <- cohesion_matrix(cultures)
+\end{Sinput}
+\end{Schunk}
+
+\begin{Schunk}
+\begin{Sinput}
+plot_community_graphs(
+  cultures_cohesion,
+  edge_width_factor = 30,
+  emph_strong = 3,
+  vertex.label.cex = 0.7,
+  vertex.size = 3,
+  vertex.label.dist = 1
+)
+\end{Sinput}
+\end{Schunk}
+
+\begin{Schunk}
+\begin{figure}
+\includegraphics[width=1\linewidth]{fig7} \caption[Community structure for cultural distance data]{Community structure for cultural distance data.}\label{fig:figculture}
+\end{figure}
+\end{Schunk}
+
+This example demonstrates key features of the PaLD approach that
+distinguish it from commonly used clustering techniques. In addition to
+viewing the community structure as seen in Figure \ref{fig:figculture},
+we can also visualize a two dimensional plot of cohesion versus distance
+for each of the included distance pairs using the
+\texttt{dist\_cohesion\_plot()} function, as seen below (Figure
+\ref{fig:figco}).
+
+\begin{Schunk}
+\begin{Sinput}
+dist_cohesion_plot(cultures)
+\end{Sinput}
+\begin{figure}
+\includegraphics{dagostino-mcgowan_files/figure-latex/figco-1} \caption[Distance versus cohesion for each TODO The identified communties are shown via the 5 colors -- these map to the same clusters idenfitied in Figure 7]{Distance versus cohesion for each TODO The identified communties are shown via the 5 colors -- these map to the same clusters idenfitied in Figure 7.}\label{fig:figco}
+\end{figure}
+\end{Schunk}
+
+Notice here the magnitude of the distances within each of the identified
+communities varies substantially between regions, yet the cohesion
+remains strong. In other words, focusing particularly on identified
+community consisting of regions in the United States (light purple)
+compared to the community consisting of regions in India (turquoise) the
+strong cohesions are similar (Figure \ref{fig:figculture}), yet the
+distances themselves are markedly different (Figure \ref{fig:figco}). In
+fact, the largest distance between regions in the Unites States is
+smaller than that between any two regions in India (Figure
+\ref{fig:figco}). Classic distance-based clustering techniques tend to
+be unable to maintain this distinction often breaking India into several
+smaller communities. Additionally, currently available techniques
+require specification of parameters, as seen in the previous section.
+For further discussion of the cultural distance data in relation to
+community analysis see \citet{berenhaut2022social}.
 

### 2023-05-02 r-journal/dagostino-mcgowan.R

@@ -224,3 +224,40 @@ plot(
   asp = 1
 )
 
+
+## -----------------------------------------------------------------------------
+cultures_cohesion <- cohesion_matrix(cultures)
+
+
+## ---- eval = FALSE------------------------------------------------------------
+#> plot_community_graphs(
+#>   cultures_cohesion,
+#>   edge_width_factor = 30,
+#>   emph_strong = 3,
+#>   vertex.label.cex = 0.7,
+#>   vertex.size = 3,
+#>   vertex.label.dist = 1
+#> )
+
+
+## ---- include = FALSE, results = "hide"---------------------------------------
+set.seed(7)
+png("fig7.png", res = 300, height = 3000, width = 3000)
+plot_community_graphs(
+  cultures_cohesion,
+  edge_width_factor = 30,
+  emph_strong = 3,
+  vertex.label.cex = 0.7,
+  vertex.size = 3,
+  vertex.label.dist = 1
+)
+dev.off()
+
+
+## ----figculture, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Community structure for cultural distance data.", out.width = "100%"----
+knitr::include_graphics("fig7.png")
+
+
+## ----figco, fig.cap = "Distance versus cohesion for each TODO The identified communties are shown via the 5 colors -- these map to the same clusters idenfitied in Figure 7."----
+dist_cohesion_plot(cultures)
+

### 2023-05-02 r-journal/dagostino-mcgowan.tex
update!

@@
 \begin{Soutput}
-#> IGRAPH 2f2c953 UNW- 6 3 -- 
+#> IGRAPH 6d52e5e UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from 2f2c953 (vertex names):
+#> + edges from 6d52e5e (vertex names):
 #> [1] A--B A--E D--F
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from 828fab3:
+#> + 8/87 vertices, named, from 3aecdd1:
 #> [1] Italian         Ladin           Provencal       Walloon        

### 2023-05-03 r-journal/dagostino-mcgowan.Rmd
update with Kate's new affiliation

@@
   - name: Katherine Moore
-    affiliation: Wake Forest Unversity
+    affiliation: Amherst College
     address:
-    - Winston-Salem, NC
-    - 27106
-    email: mooreke@wfu.edu 
+    - Amherst, MA
+    - 01002
+    email: kmoore@amherst.edu 
   - name: Kenneth Berenhaut 

### 2023-05-03 r-journal/dagostino-mcgowan.tex
update with Kate's new affiliation

@@
 \begin{Soutput}
-#> IGRAPH 6d52e5e UNW- 6 3 -- 
+#> IGRAPH 4d37a4f UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from 6d52e5e (vertex names):
+#> + edges from 4d37a4f (vertex names):
 #> [1] A--B A--E D--F
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from 3aecdd1:
+#> + 8/87 vertices, named, from a461d26:
 #> [1] Italian         Ladin           Provencal       Walloon        
@@
 Katherine Moore\\
-Wake Forest Unversity\\%
-Winston-Salem, NC\\ 27106\\
+Amherst College\\%
+Amherst, MA\\ 1002\\
 %
@@
 %
-\href{mailto:mooreke@wfu.edu}{\nolinkurl{mooreke@wfu.edu}}%
+\href{mailto:kmoore@amherst.edu}{\nolinkurl{kmoore@amherst.edu}}%
 }

### 2023-09-28 2023-r-journal/dagostino-mcgowan.Rmd
upload files

--- 2023-05-02 r-journal/dagostino-mcgowan.Rmd
@@
 abstract: >
-  Partitioned Local Depth (PaLD) is a framework for a holistic consideration of the community structure of distance-based data.
+  Partitioned Local Depth (PaLD) is a framework for holistic consideration of community structure for distance-based data.
 This paper describes an R package, \CRANpkg{pald}, for calculating Partitioned Local Depth (PaLD) probabilities, implementing community analyses, and creating data visualizations to display community structure.
-We describe how to use the package, walk through several examples, and contrast the results with commonly used techniques.
+We describe how to use the package, walk through several examples, and discuss the approach in relation to commonly used techniques.
 output: rticles::rjournal_article
@@
 
-Partitioned Local Depth (PaLD) is a framework for a holistic consideration of community structure for distance-based data.
+Partitioned Local Depth (PaLD) is a framework for holistic consideration of community structure for distance-based data.
 Leveraging a socially inspired perspective, the method provides network-based community information which is founded on a new measure of local depth and pairwise cohesion (partitioned local depth).
@@
 Building on existing approaches to (global) depth, local depth expresses features of centrality via an interpretable probability which is free of parameters and robust to outliers.
-Then, partitioning the probability which defines local depth, we obtain a measure of cohesion between pairs of points.
+Then, partitioning the probability which defines local depth, we obtain a measure of cohesion between pairs of individuals.
 Both local depth and cohesion reflect aspects of relative position (rather than absolute distance) and provide a straightforward way to account for varying density across the space.
-Specifically, as shown in \citet{berenhaut2022social}, provided that two sets are separated (in the sense that the minimum between-set distance is greater than the maximum within-set distance), cohesion is invariant under the contraction and dilation of the distances within each set.
+As shown in \citet{berenhaut2022social}, provided that two sets are separated (in the sense that the minimum between-set distance is greater than the maximum within-set distance), cohesion is invariant under the contraction and dilation of the distances within each set.
 This property may be particularly valuable when one has reason to believe that there is heterogeneity in density across the space.
@@
 
-Throughout this paper, we will display the network obtained from cohesion using a force-directed graph drawing algorithm and emphasize the strong ties (colored by connected component).
+Networks obtained from cohesion can be displayed using a force-directed graph drawing algorithm; here, we will graphically emphasize strong ties (colored by connected component).
 We refer to the connected components of the network of strong ties as community "clusters."  Note that to qualify as a cluster in this definition, one may not have any strong ties with those outside the cluster, and thus the existence of disjoint groups is a strong signal for separation.
@@
 If one wishes to further break the community graph into groups, one may consider using community detection methods (such as  spectral clustering or the Louvain algorithm), as available, say, in the \CRANpkg{igraph} package.
-Though only briefly considered here, one may also use the collection of strong ties in place of (weighted) k-nearest neighbors in settings such as classification and smoothing.
-Overall, the structural information obtained from local depth, cohesion and community graphs can provide a holistic perspective on the data which does not require the use of distributional assumptions, optimization criteria nor additional user inputs.
+One may also use the collection of strong ties in place of (weighted) $k$-nearest neighbors in settings such as classification and smoothing.
+Overall, the structural information obtained from local depth, cohesion and community graphs can provide a holistic perspective to the data which does not require the use of distributional assumptions, optimization criteria nor additional user inputs.
 
 We present a new package, \CRANpkg{pald}, for calculating Partitioned Local Depth (PaLD) probabilities, implementing community analyses, and creating data visualizations to display community structure.
-This paper will describe how to use the package, walk through several examples, and compare the method to commonly used techniques in R.
-Together, these demonstrate both the novelty of the method and utility of the implementation in package described.
+This paper describes how to use the package, walks through several examples, and discusses the approach in the context of commonly used techniques, demonstrating both the novelty of the method and utility of the implementation in the package described.
 
@@
 
-In addition, the package provides a number of pertinent example data sets commonly used to demonstrate cluster algorithms, including a synthetic data set of two-dimensional points created by \citet{gionis1clustering} to demonstrate clustering aggregation, clustering data generated from the scikit-learn Python package \citep{pedregosa2011scikit}, data describing cognate relationships between words across 87 Indo-European languages \citep{dyen92}, data compiled by \cite{tissue} of tissue gene expressions, data from the World Values Survey \citep{inglehart2014world} on cultural values regarding family, religion, education, and institutions for several regions \citep{muthukrishna2020beyond}, and three example data sets generated for the \citet{berenhaut2022social} paper.
+In addition, the package provides a number of pertinent example data sets which may be used to demonstrate community analysis, including a synthetic data set of two-dimensional points created by \citet{gionis1clustering} to demonstrate aggregation, clustering data generated from the scikit-learn Python package \citep{pedregosa2011scikit}, data describing cognate relationships between words across 87 Indo-European languages \citep{dyen92}, data compiled by \cite{tissue} of tissue gene expressions, data emplying information from the World Values Survey \citep{inglehart2014world} on cultural values regarding family, religion, education, and institutions for several regions \citep{muthukrishna2020beyond}, and three example data sets generated for the \citet{berenhaut2022social} paper.
 
@@
 
-The input for the Partitioned Local Depths (PaLD) computations is a distance matrix or `dist` object.
+For the purposes of the \CRANpkg{pald} package, the  sole input for the Partitioned Local Depth (PaLD) computations is a distance matrix or `dist` object.
 Note that the collection of input distances (or dissimilarities) does not need to satisfy the triangle inequality nor be symmetric.
+More generally, the method only requires triplet distance comparisons, as opposed to exact numeric distances (see \citet{berenhaut2022social}).
 
-For demonstration purposes, we will show how one can compute a distance matrix from an input data frame with, say, two variables `x1` and `x2`.
+For demonstration purposes, we first show how one can compute a distance matrix from an input data frame with, say, two variables `x1` and `x2`.
 The input data may be of any dimension; in fact the PaLD framework provides advantages when considering high-dimensional data (see the **Examples** section as well as \citet{berenhaut2022social}).
@@
 
-The `dist()` function converts an input data frame into a distance matrix, as demonstrated below.
+The `dist` function returns a (default Euclidean) pairwise distance matrix for an input data frame, as demonstrated below.
 If the data are already provided as a distance matrix (or `dist` object), the user can skip to the next step.
-Note that the distance matrix needed for the subsequent functions doesn't need to be a `dist` object and *need not* be symmetric.
+Note that the distance matrix needed for the subsequent functions does not need to be a `dist` object and *need not* be symmetric.
 
@@
 The function above creates a `dist` object.
-If converted to a matrix, this will be an $n\times n$ distance matrix, where $n$ corresponds to the number of observations in the original data frame, in this example $n = 6$.
+If converted to a matrix, this will be an $n\times n$ distance matrix, where $n$ corresponds to the number of observations in the original data frame (in this example $n = 6$).
 
-This `dist` object, or a distance matrix, can then be passed to the `cohesion_matrix()` function in order to calculate the pairwise cohesion values.
+This `dist` object, or a distance matrix, can then be passed to the `cohesion_matrix` function in order to calculate pairwise cohesion values.
 
 Cohesion is an interpretable probability that reflects the strength of alignment of two points within local regions.
-It captures aspects of the relative positioning of points and accounts for varying density across the space.
+It captures aspects of the relative positioning of points and accounts for varying density across the space (see \citet{berenhaut2022social}).
 
@@
 
-The _cohesion matrix_ output by the `cohesion_matrix()` function is the main input for the majority of the remaining functions.
+The _cohesion matrix_ output by the `cohesion_matrix` function is the main input for the majority of the remaining functions.
 
@@
 
-The `local_depths()` function calculates the _depth_ of each point, outputting a vector of local depths.
-Local depth is an interpretable probability which reflects aspects of relative position and centrality via distance comparisons (i.e., $d(z, x) < d(z, y)$, see \citet{berenhaut2022social}).
+The `local_depths` function calculates the _depth_ of each point, outputting a vector of local depths.
+Local depth is an interpretable probability which reflects aspects of relative position and centrality via distance comparisons (e.g., $d(z, x) < d(z, y)$, see \citet{berenhaut2022social}).
 
@@
 
-The `strong_threshold()` function will calculate the cohesion threshold for strong ties.
-This is equal to half the average of the diagonal of the cohesion matrix \citep{berenhaut2022social}, and is a threshold that may be used to distinguish between strong and weak ties.
+The `strong_threshold` function will calculate the cohesion threshold for strong ties, which reflects typical cohesion for local points (see \citet{berenhaut2022social}).
+Computationally, this is equal to half the average of the diagonal of the cohesion matrix, and is a threshold that may be used to distinguish between strong and weak ties.
 
@@
 
-The function `cohesion_strong()` will update the cohesion matrix to set all weak ties to zero (via the `strong_threshold()` function).
+The function `cohesion_strong` will update the cohesion matrix to set all weak ties to zero (via the `strong_threshold` function).
 Optionally, the matrix will also be symmetrized, using the entry-wise (parallel) minimum of the cohesion matrix and its transpose, with the default parameter `symmetric = TRUE`.
@@
 
-The `community_graphs()` function takes the cohesion matrix and creates \CRANpkg{igraph} objects, graphs that describe the relationship between the points.
+The `community_graphs` function takes the cohesion matrix and creates \CRANpkg{igraph} objects, graphs that describe the (symmetrized) relationship structure between points.
 This function will output a list of three objects:
@@
 the minimum of the two directed cohesion values for any given pair) is greater than the threshold for strong ties
-* `layout`: the graph layout, using the Fruchterman Reingold (FR) force-directed graph drawing for the graph `G`
+* `layout`: the graph layout. By default this is provided by the Fruchterman Reingold (FR) force-directed graph drawing algorithm for the graph `G`, as implemented in the \CRANpkg{igraph} package.
 
@@
 
-Here we see that there are three connected components, ties `A-B` and `A-E` which form the first cluster, and the tie `D-F` which forms another.
+Here we see that there are three connected components, ties `A-B` and `A-E` form the first community cluster, and the tie `D-F` which forms another.
 
-The `any_isolated()` function will check whether there are any isolated points.
+The `any_isolated()` function will check whether there are any isolated points (according to cohesion).
 
@@
 The "community clusters" identified by PaLD are the connected components of the graph of strong ties, `G_strong`.
-To directly calculate them, we can use the `community_clusters()` function.
-This will output a data frame with two columns, the first will correspond to the `point`, as identified by the row name of the original input data frame, `df`, the second will identify the `cluster` that each point belongs to.
+To directly calculate these, we can use the `community_clusters` function.
+This will output a data frame with two columns; the first corresponds to the `point`, as identified by the row name of the original input data frame, `df`, the second identifies the `community` that the point belongs to.
 
@@
 
-In this example, three clusters are identified with these six points.
-Points `A`, `B`, and `E` fall into Cluster 1.
-Point `C` is in Cluster 2 (a cluster of size 1) and points `D` and `F` fall into Cluster 3.
+In this example, three communities are identified with these six points.
+Points `A`, `B`, and `E` fall into Community 1.
+Point `C` is in Community 2 (a community of size 1) and points `D` and `F` fall into Community 3.
 
@@
 When visualizing these points, it is important to have the aspect ratio of the x and y axes equal to 1 so as to not distort distances.
-When using the \CRANpkg{ggplot2} package for this visualization, one can use the `coord_fixed(ratio = 1)` function.
-If using the `plot()` function included in the base library, one can use the `asp = 1` argument.
+When using the \CRANpkg{ggplot2} package for this visualization, one can use the command `coord_fixed(ratio = 1)`.
+If using the `plot` function included in the base library, one can use the `asp = 1` argument.
 
@@
 
-We can pass the cohesion matrix to the `plot_community_graphs()` function to view the relationship between points (Figure \ref{fig:fig2}).
-Notice in this plot the layout does not match that of the original data frame as seen in Figure \ref{fig:fig1}.
-Since our original data is two dimensional, it may be reasonable to use this as the layout.
-Figure \ref{fig:fig3} will make this update as well as update some of the aesthetics, such as employing more readable labels.
+We can pass the cohesion matrix to the `plot_community_graphs` function to view the relationship between points (Figure \ref{fig:fig2}).
+The function will also permit parameters that can be passed to `plot.igraph` via the `...` argument.
 
 ```{r fig2, fig.cap = "PaLD graph displaying the relationship between the points in data frame `df`"}
-plot_community_graphs(cohesion)
+plot_community_graphs(cohesion, 
+                      vertex.label.cex = 2,
+                      vertex.label.dist = 0.9)
 ```
 
+Notice in this plot the force-directed layout does not match that of the original data frame as seen in Figure \ref{fig:fig1}.
+Since our original data is two-dimensional, it may be reasonable to use the latter as the layout for plotting.
+Figure \ref{fig:fig3} includes this update as well as others addressing some of the aesthetics, such as employing more readable labels.
 The `layout` argument allows the user to pass a matrix to dictate the 2-dimensional layout of the graph.
-For example, if we wanted the graph to match the visualization displayed in Figure \ref{fig:fig1}, we could pass `as.matrix(df)`, or a matrix of the data frame `df` to the `layout` argument  (Figure \ref{fig:fig3}).
-Additionally, the `plot_community_graphs()` function will also permit parameters that can be passed to the `plot.igraph()` function.
-We can pass arguments to the `plot.igraph` function via the `...` argument; for example to increase the vertex size and change the vertex label color, we can specify `vertex.size = 100` and `vertex.label.color = "white"`.
-Additionally, to allow axes, we use `axes = TRUE`, and to put them back on the original scale we set `rescale = FALSE`, resetting the axis limits using `xlim` and `ylim`.
+For example, if we wanted the graph to match the visualization displayed in Figure \ref{fig:fig1}, we could pass `as.matrix(df)` (a matrix of the data frame `df`) to the `layout` argument  (Figure \ref{fig:fig3}).
+Here, we increase the vertex size and change the vertex label color, through the  argument specifications `vertex.size = 100` and `vertex.label.color = "white"`.
+Additionally, to allow axes, we use `axes = TRUE`, and to put these back on the original scale we set `rescale = FALSE`, resetting the axis limits using `xlim` and `ylim`.
 The `par(pty = "s")` function forces the subsequent plot to be square.
@@
 
-We will demonstrate the utility of the \CRANpkg{pald} package in three illustrative examples.
+We will demonstrate the utility of the \CRANpkg{pald} package through several illustrative examples.
 
-### Clustering tissue gene expression data
+### Community analysis for tissue gene expression data
 
 The first example is from a subset of data from \citet{zilliox2007gene}, \citet{mccall2011gene}, and \citet{mccall2014gene}, obtained from the **tissuesGeneExpression** bioconductor package \citep{tissue} consisting of 22,215-dimensional gene expression data from 189 tissue samples.
-A `dist` object was created using this data set and is included in the \CRANpkg{pald} package in an object called `tissue_dist`.
+A (Euclidean) `dist` object was created using this data set and is included in the \CRANpkg{pald} package in an object called `tissue_dist`.
 
@@
 
-The `community_clusters()` function can be used to identify the community cluster corresponding to each tissue sample.
-Since the output is a data frame, we can summarize the result using commonly employed data analysis techniques.
-For demonstration purposes, we will use the \CRANpkg{dplyr} package to summarize the contribution of clusters.
+The `community_clusters` function can be used to identify the community corresponding to each tissue sample.
+Since the output is a data frame, we can summarize the result using commonly employed R functions.
+For demonstration purposes, we will use the \CRANpkg{dplyr} package to summarize the community contributions.
 
@@
 community_clusters(tissue_cohesion) |>
-  dplyr::count(cluster, point)
+  dplyr::count(community, point)
 ```
 
-From this, we can glean that Cluster 1 consists of two types of tissue, the kidney and endometrium. Cluster 2 is comprised of only the hippocampus.
-
-We can also display the relationships between tissue samples using the `plot_community_graphs()` function (Figure \ref{fig:fig4}).
+From this, we can glean, for instance, that Community 1 consists of two types of tissue, the kidney and endometrium, Community 2 is comprised of only the hippocampus, etc.
+More informatively, we can display relationships between tissue samples, both locally and globally through the `plot_community_graphs` function  (Figure \ref{fig:fig4}).
 For clarity of the display, we show how to remove the labels using `show_labels = FALSE`.
-We will instead color by the labels by passing these to the `vertex.color` parameter to the `plot.igraph` function (via the `...` argument).
+We will instead color according to the labels by passing these to the `vertex.color` argument for the `plot.igraph` function (via the `...` argument).
 Similarly, we can add a legend using the `legend()` function, as you would for an \CRANpkg{igraph} visualization.
-Additionally, we use the `edge_width_factor` and `emph_strong` arguments to adjust the width of the lines between and within PaLD clusters.
+Additionally, we use the `edge_width_factor` and `emph_strong` arguments to adjust the width of the lines between and within PaLD communities.
 
@@
 ```{r fig4, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Community cluster network for the tissue data.
-The line colors indicate the PaLD clusters, the point colors indicate the tissue classification.", out.width = "100%"}
+The line colors indicate the PaLD communities; the point colors indicate the tissue classification.", out.width = "100%"}
 knitr::include_graphics("fig5.png")
+```
+
+Note that while Community 1 does contain kidney and endometrium tissues, the respective sub-communities are well-separated.
+Of the 357 strong inter-community ties within Community 1, only three are between the two sub-communities.
+In fact, these are the only strong ties between different tissue types in the community cluster network (of 1064 total); a summary of community strong ties can be produced via the following code.
+Note that `tissue_graph_strong` is an \CRANpkg{igraph} object corresponding to the graph displayed in Figure \ref{fig:fig4}, restricted only to strong ties.
+
+```{r}
+tissue_graphs <- community_graphs(tissue_cohesion)
+tissue_graph_strong <- tissue_graphs[["G_strong"]]
+E <- igraph::get.edgelist(tissue_graph_strong)
+table(apply(E, 1, paste, collapse = ","))
 ```
@@
 
-This example performs a PaLD analysis on a data set from \citet{dyen92} that examines the relationship between 87 Indo-European languages from the perspective of cognates, coded using 2,655-dimensional binary vectors.
+This example explors a data set from \citet{dyen92} that summarizes relationships between 87 Indo-European languages from the perspective of cognates, coded using 2,655-dimensional binary vectors.
 A `dist` object was created from this data set and is included in the \CRANpkg{pald} package in an object called `cognate_dist`.
 
-This example will demonstrate how you can apply functions in the \CRANpkg{igraph} package to objects output from the \CRANpkg{pald} package.
-We can first use the `cohesion_matrix()` function to calculate the cohesion matrix and the `community_graphs()` function to create a list with the weighted community graph, the weighted community graph with only strong ties included, and the layout.
+Here we will demonstrate how one can further apply functions in the \CRANpkg{igraph} package to objects output from the \CRANpkg{pald} package.
+We can first use the `cohesion_matrix` function to calculate the cohesion matrix and the `community_graphs` function to create a list with the weighted community graph, the weighted community graph with only strong ties included, and the layout.
 From this, we can extract the graph with only the strong ties, here called `cognate_graph_strong`.
@@
 
-We can then use the `neighbors()` function from the \CRANpkg{igraph} package to extract the strong neighbors in this graph.
-For example, if we wanted to extract all neighbors for the language "French", we could run the following.
+We can then use the `neighbors` function from the \CRANpkg{igraph} package to extract the strong neighbors in this graph.
+For example, we can extract all neighbors for the language "French", via the followign code.
 
@@
 
-We can again use the `plot_community_graphs()` function to visualize the community clusters (Figure \ref{fig:figlang}).
+We can again use the `plot_community_graphs` function to visualize the community clusters (Figure \ref{fig:figlang}).
 One may note the commonly identifiable language clusters and that, under a slight rotation, some of the underlying geography is mirrored in the plot.
@@
 
-* `exdata1` is a data set consisting of 8 points to recreate Figure 1 in \citet{berenhaut2022social} 
-* `exdata2` is a data set consisting of 16 points to recreate Figure 2 in \citet{berenhaut2022social} 
-* `exdata3` is a data set consisting of 240 points to recreate Figure 4D in \citet{berenhaut2022social}
+* `exdata1` is a data set consisting of 8 points used to create Figure 1 in \citet{berenhaut2022social} 
+* `exdata2` is a data set consisting of 16 points used to create Figure 2 in \citet{berenhaut2022social} 
+* `exdata3` is a data set consisting of 240 points used to create Figure 4D in \citet{berenhaut2022social}
 
@@
 
-We will demonstrate how we can compare PaLD to two clustering methods: *k*-means and hierarchical clustering.
-The code below calculates the cohesion matrix (`exdata_cohesion`) as well as the clusters via PaLD (`exdata_pald`), *k*-means (`exdata_kmeans`) and hierarchical clustering using complete linkage (`exdata_hclust`).
+We will contrast community analysis via PaLD to standard culuster analysis by considering two common clustering methods.
+The code below calculates the cohesion matrix (`exdata_cohesion`) as well as the community clusters obtained via PaLD (`exdata_pald`), along with clusters obtained via *k*-means (`exdata_kmeans`) and hierarchical clustering using complete linkage (`exdata_hclust`).
 
@@
 
-We can compare this to the clustering generated by *k*-means and hierarchical clustering (Figure \ref{fig:fig5}).
+The information is displayed in Figure \ref{fig:fig5}.
 
-```{r fig5, fig.cap = "PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al. (2022)) compared to k-means and hierarchical clustering with k = 8."}
+```{r fig5, fig.cap = "PaLD, k-means, and hierarchical 8-clustering of randomly generated example data (from \\citet{berenhaut2022social}; Figure 4D)."}
 par(mfrow = c(1, 3), pty = "s")
@@
   ylab = "",
-  main = "PaLD Clusters",
+  main = "PaLD Communities",
   asp = 1
@@
 
-Cohesion is particularly useful when considering data with varying local density, see discussion in \citet{berenhaut2022social}.
-Note that the PaLD algorithm is able to detect the eight natural groups within the data without the use of any additional inputs (e.g., number of clusters) nor optimization criteria.
-Despite the user input of the "correct" number of clusters (i.e., $k = 8$) both *k*-means and hierarchical clustering did not give the desired result.
-For further examples, discussion, and theoretical results, see \citet{berenhaut2022social}.
+Cohesion is particularly useful when considering data with varying local density; see \citet{berenhaut2022social} for further examples, discussion, and theoretical results.
+Note that the PaLD algorithm is able to detect the eight natural groups within the data  (along with inter- and intra-community structure not displayed here) without the use of any additional inputs (e.g., number of clusters) nor optimization criteria.
+Despite the user input of the "correct" number of clusters (i.e., $k = 8$) both *k*-means and hierarchical clustering do not provide the desired result.
+
+### Cultural and Pyschological distance analysis
+
+In this example we perform a PaLD analysis for cultural distances obtained in \citet{muthukrishna2020beyond} from two recent waves of the World Values Survey (2005 to 2009 and 2010 to 2014; see \citet{inglehart2014world}).
+Distances are
+computed using the cultural fixation index (CFST), which is a measure built on the framework of fixation indices from population
+biology (\citet{bell2009culture,cavalli1994history}).
+Recall that the foundation of PaLD in within-triplet comparisons allows for the employment of application-dependent and non-Euclidean measures of dissimilarity.
+The `dist` object is
+included in the \CRANpkg{pald} package (`cultures`). We will first create the cohesion matrix using the `dist` object `cultures`, and proceed to plot the community graph.
+
+```{r}
+cultures_cohesion <- cohesion_matrix(cultures)
+```
+
+```{r, eval = FALSE}
+plot_community_graphs(
+  cultures_cohesion,
+  edge_width_factor = 30,
+  emph_strong = 3,
+  vertex.label.cex = 0.7,
+  vertex.size = 3,
+  vertex.label.dist = 1
+)
+```
+
+```{r, include = FALSE, results = "hide"}
+set.seed(7)
+png("fig7.png", res = 300, height = 3000, width = 3000)
+plot_community_graphs(
+  cultures_cohesion,
+  edge_width_factor = 30,
+  emph_strong = 3,
+  vertex.label.cex = 0.7,
+  vertex.size = 3,
+  vertex.label.dist = 1
+)
+dev.off()
+```
+
+```{r figculture, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Community structure for cultural distance data.", out.width = "100%"}
+knitr::include_graphics("fig7.png")
+```
+
+In addition to viewing the local and global community structure as seen in Figure \ref{fig:figculture}, the \CRANpkg{pald} package allows for a two-dimensional display of cohesion against distance for the data, via the `dist_cohesion_plot` function, as seen below (Figure \ref{fig:figco}).
+
+```{r figco, fig.cap = "A plot of cohesion versus distance for the data. The identified communities are colored as in Figure \\ref{fig:figculture}."}
+dist_cohesion_plot(cultures, mutual = TRUE)
+```
+
+Notice here that the magnitude of the distances within each of the identified communities varies substantially between regions; in fact, the most disparate two regions in the United States (at distance $\approx$ 0.027) are far closer than the two most similar in India (at distance $\approx$ 0.043).
+Despite this, India remains a cohesive whole, and locally disparate regions in the United States such as East South Central and California are not strongly cohesive.
+For discussion of subtleties in local density (see \citet{berenhaut2022social}.
+Additionally, currently available techniques require specification of parameters, as seen in the previous section.
+For further discussion of the cultural distance data in relation to community analysis see \citet{berenhaut2022social}.
 
@@
 
-This paper introduces the \CRANpkg{pald} package, demonstrating its utility for providing parameter-free clustering which can easily be implemented for a variety of data sets.
-We provide example code as well as compare the method to commonly used clustering techniques, *k*-means and hierarchical clustering.
+This paper introduces the \CRANpkg{pald} package, demonstrating its utility for providing parameter-free community analysis which can easily be implemented for a variety of data sets, supplementing results from other methods for clustering, embedding, nearest neighbors, etc.
+Example code is provided advantages over other techniques are highlighted.
 

### 2023-09-28 2023-r-journal/dagostino-mcgowan.tex
upload files

--- 2023-05-03 r-journal/dagostino-mcgowan.tex
@@
 \abstract{%
-Partitioned Local Depth (PaLD) is a framework for a holistic
-consideration of the community structure of distance-based data. This
-paper describes an R package, \CRANpkg{pald}, for calculating
-Partitioned Local Depth (PaLD) probabilities, implementing community
-analyses, and creating data visualizations to display community
-structure. We describe how to use the package, walk through several
-examples, and contrast the results with commonly used techniques.
+Partitioned Local Depth (PaLD) is a framework for holistic consideration
+of community structure for distance-based data. This paper describes an
+R package, \CRANpkg{pald}, for calculating Partitioned Local Depth
+(PaLD) probabilities, implementing community analyses, and creating data
+visualizations to display community structure. We describe how to use
+the package, walk through several examples, and discuss the approach in
+relation to commonly used techniques.
 }
@@
 
-Partitioned Local Depth (PaLD) is a framework for a holistic
-consideration of community structure for distance-based data. Leveraging
-a socially inspired perspective, the method provides network-based
-community information which is founded on a new measure of local depth
-and pairwise cohesion (partitioned local depth). The method does not
-require distributional assumptions, optimization criteria, nor
-extraneous inputs. A complete description of the perspective, together
-with a discussion of the underlying social motivation, theoretical
-results, and applications to additional data sets is provided in
+Partitioned Local Depth (PaLD) is a framework for holistic consideration
+of community structure for distance-based data. Leveraging a socially
+inspired perspective, the method provides network-based community
+information which is founded on a new measure of local depth and
+pairwise cohesion (partitioned local depth). The method does not require
+distributional assumptions, optimization criteria, nor extraneous
+inputs. A complete description of the perspective, together with a
+discussion of the underlying social motivation, theoretical results, and
+applications to additional data sets is provided in
 \citet{berenhaut2022social}.
@@
 which defines local depth, we obtain a measure of cohesion between pairs
-of points. Both local depth and cohesion reflect aspects of relative
-position (rather than absolute distance) and provide a straightforward
-way to account for varying density across the space. Specifically, as
+of individuals. Both local depth and cohesion reflect aspects of
+relative position (rather than absolute distance) and provide a
+straightforward way to account for varying density across the space. As
 shown in \citet{berenhaut2022social}, provided that two sets are
@@
 
-Throughout this paper, we will display the network obtained from
-cohesion using a force-directed graph drawing algorithm and emphasize
-the strong ties (colored by connected component). We refer to the
-connected components of the network of strong ties as community
-``clusters.'' Note that to qualify as a cluster in this definition, one
-may not have any strong ties with those outside the cluster, and thus
-the existence of disjoint groups is a strong signal for separation.
-Here, clusters are identified without additional user inputs nor
-optimization criteria. If one wishes to further break the community
-graph into groups, one may consider using community detection methods
-(such as spectral clustering or the Louvain algorithm), as available,
-say, in the \CRANpkg{igraph} package. Though only briefly considered
-here, one may also use the collection of strong ties in place of
-(weighted) k-nearest neighbors in settings such as classification and
+Networks obtained from cohesion can be displayed using a force-directed
+graph drawing algorithm; here, we will graphically emphasize strong ties
+(colored by connected component). We refer to the connected components
+of the network of strong ties as community ``clusters.'' Note that to
+qualify as a cluster in this definition, one may not have any strong
+ties with those outside the cluster, and thus the existence of disjoint
+groups is a strong signal for separation. Here, clusters are identified
+without additional user inputs nor optimization criteria. If one wishes
+to further break the community graph into groups, one may consider using
+community detection methods (such as spectral clustering or the Louvain
+algorithm), as available, say, in the \CRANpkg{igraph} package. One may
+also use the collection of strong ties in place of (weighted)
+\(k\)-nearest neighbors in settings such as classification and
 smoothing. Overall, the structural information obtained from local
 depth, cohesion and community graphs can provide a holistic perspective
-on the data which does not require the use of distributional
+to the data which does not require the use of distributional
 assumptions, optimization criteria nor additional user inputs.
@@
 creating data visualizations to display community structure. This paper
-will describe how to use the package, walk through several examples, and
-compare the method to commonly used techniques in R. Together, these
-demonstrate both the novelty of the method and utility of the
-implementation in package described.
+describes how to use the package, walks through several examples, and
+discusses the approach in the context of commonly used techniques,
+demonstrating both the novelty of the method and utility of the
+implementation in the package described.
 
@@
 In addition, the package provides a number of pertinent example data
-sets commonly used to demonstrate cluster algorithms, including a
+sets which may be used to demonstrate community analysis, including a
 synthetic data set of two-dimensional points created by
-\citet{gionis1clustering} to demonstrate clustering aggregation,
-clustering data generated from the scikit-learn Python package
+\citet{gionis1clustering} to demonstrate aggregation, clustering data
+generated from the scikit-learn Python package
 \citep{pedregosa2011scikit}, data describing cognate relationships
 between words across 87 Indo-European languages \citep{dyen92}, data
-compiled by \cite{tissue} of tissue gene expressions, data from the
-World Values Survey \citep{inglehart2014world} on cultural values
-regarding family, religion, education, and institutions for several
-regions \citep{muthukrishna2020beyond}, and three example data sets
-generated for the \citet{berenhaut2022social} paper.
+compiled by \cite{tissue} of tissue gene expressions, data emplying
+information from the World Values Survey \citep{inglehart2014world} on
+cultural values regarding family, religion, education, and institutions
+for several regions \citep{muthukrishna2020beyond}, and three example
+data sets generated for the \citet{berenhaut2022social} paper.
 
@@
 
-The input for the Partitioned Local Depths (PaLD) computations is a
-distance matrix or \texttt{dist} object. Note that the collection of
-input distances (or dissimilarities) does not need to satisfy the
-triangle inequality nor be symmetric.
+For the purposes of the \CRANpkg{pald} package, the sole input for the
+Partitioned Local Depth (PaLD) computations is a distance matrix or
+\texttt{dist} object. Note that the collection of input distances (or
+dissimilarities) does not need to satisfy the triangle inequality nor be
+symmetric. More generally, the method only requires triplet distance
+comparisons, as opposed to exact numeric distances (see
+\citet{berenhaut2022social}).
 
-For demonstration purposes, we will show how one can compute a distance
+For demonstration purposes, we first show how one can compute a distance
 matrix from an input data frame with, say, two variables \texttt{x1} and
@@
 
-The \texttt{dist()} function converts an input data frame into a
-distance matrix, as demonstrated below. If the data are already provided
-as a distance matrix (or \texttt{dist} object), the user can skip to the
-next step. Note that the distance matrix needed for the subsequent
-functions doesn't need to be a \texttt{dist} object and \emph{need not}
-be symmetric.
+The \texttt{dist} function returns a (default Euclidean) pairwise
+distance matrix for an input data frame, as demonstrated below. If the
+data are already provided as a distance matrix (or \texttt{dist}
+object), the user can skip to the next step. Note that the distance
+matrix needed for the subsequent functions does not need to be a
+\texttt{dist} object and \emph{need not} be symmetric.
 
@@
 matrix, this will be an \(n\times n\) distance matrix, where \(n\)
-corresponds to the number of observations in the original data frame, in
-this example \(n = 6\).
+corresponds to the number of observations in the original data frame (in
+this example \(n = 6\)).
 
 This \texttt{dist} object, or a distance matrix, can then be passed to
-the \texttt{cohesion\_matrix()} function in order to calculate the
-pairwise cohesion values.
+the \texttt{cohesion\_matrix} function in order to calculate pairwise
+cohesion values.
 
@@
 relative positioning of points and accounts for varying density across
-the space.
+the space (see \citet{berenhaut2022social}).
 
@@
 
-The \emph{cohesion matrix} output by the \texttt{cohesion\_matrix()}
+The \emph{cohesion matrix} output by the \texttt{cohesion\_matrix}
 function is the main input for the majority of the remaining functions.
@@
 
-The \texttt{local\_depths()} function calculates the \emph{depth} of
-each point, outputting a vector of local depths. Local depth is an
+The \texttt{local\_depths} function calculates the \emph{depth} of each
+point, outputting a vector of local depths. Local depth is an
 interpretable probability which reflects aspects of relative position
-and centrality via distance comparisons (i.e., \(d(z, x) < d(z, y)\),
+and centrality via distance comparisons (e.g., \(d(z, x) < d(z, y)\),
 see \citet{berenhaut2022social}).
@@
 
-The \texttt{strong\_threshold()} function will calculate the cohesion
-threshold for strong ties. This is equal to half the average of the
-diagonal of the cohesion matrix \citep{berenhaut2022social}, and is a
+The \texttt{strong\_threshold} function will calculate the cohesion
+threshold for strong ties, which reflects typical cohesion for local
+points (see \citet{berenhaut2022social}). Computationally, this is equal
+to half the average of the diagonal of the cohesion matrix, and is a
 threshold that may be used to distinguish between strong and weak ties.
@@
 
-The function \texttt{cohesion\_strong()} will update the cohesion matrix
-to set all weak ties to zero (via the \texttt{strong\_threshold()}
+The function \texttt{cohesion\_strong} will update the cohesion matrix
+to set all weak ties to zero (via the \texttt{strong\_threshold}
 function). Optionally, the matrix will also be symmetrized, using the
@@
 
-The \texttt{community\_graphs()} function takes the cohesion matrix and
-creates \CRANpkg{igraph} objects, graphs that describe the relationship
-between the points. This function will output a list of three objects:
+The \texttt{community\_graphs} function takes the cohesion matrix and
+creates \CRANpkg{igraph} objects, graphs that describe the (symmetrized)
+relationship structure between points. This function will output a list
+of three objects:
 
@@
 \item
-  \texttt{layout}: the graph layout, using the Fruchterman Reingold (FR)
-  force-directed graph drawing for the graph \texttt{G}
+  \texttt{layout}: the graph layout. By default this is provided by the
+  Fruchterman Reingold (FR) force-directed graph drawing algorithm for
+  the graph \texttt{G}, as implemented in the \CRANpkg{igraph} package.
 \end{itemize}
@@
 \begin{Soutput}
-#> IGRAPH 4d37a4f UNW- 6 3 -- 
+#> IGRAPH 198a54e UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from 4d37a4f (vertex names):
+#> + edges from 198a54e (vertex names):
 #> [1] A--B A--E D--F
@@
 Here we see that there are three connected components, ties \texttt{A-B}
-and \texttt{A-E} which form the first cluster, and the tie \texttt{D-F}
-which forms another.
+and \texttt{A-E} form the first community cluster, and the tie
+\texttt{D-F} which forms another.
 
 The \texttt{any\_isolated()} function will check whether there are any
-isolated points.
+isolated points (according to cohesion).
 
@@
 components of the graph of strong ties, \texttt{G\_strong}. To directly
-calculate them, we can use the \texttt{community\_clusters()} function.
-This will output a data frame with two columns, the first will
-correspond to the \texttt{point}, as identified by the row name of the
-original input data frame, \texttt{df}, the second will identify the
-\texttt{cluster} that each point belongs to.
+calculate these, we can use the \texttt{community\_clusters} function.
+This will output a data frame with two columns; the first corresponds to
+the \texttt{point}, as identified by the row name of the original input
+data frame, \texttt{df}, the second identifies the \texttt{community}
+that the point belongs to.
 
@@
 \begin{Soutput}
-#>   point cluster
+#>   point community
 #> A     A         1
@@
 
-In this example, three clusters are identified with these six points.
-Points \texttt{A}, \texttt{B}, and \texttt{E} fall into Cluster 1. Point
-\texttt{C} is in Cluster 2 (a cluster of size 1) and points \texttt{D}
-and \texttt{F} fall into Cluster 3.
+In this example, three communities are identified with these six points.
+Points \texttt{A}, \texttt{B}, and \texttt{E} fall into Community 1.
+Point \texttt{C} is in Community 2 (a community of size 1) and points
+\texttt{D} and \texttt{F} fall into Community 3.
 
@@
 distances. When using the \CRANpkg{ggplot2} package for this
-visualization, one can use the \texttt{coord\_fixed(ratio\ =\ 1)}
-function. If using the \texttt{plot()} function included in the base
-library, one can use the \texttt{asp\ =\ 1} argument.
+visualization, one can use the command
+\texttt{coord\_fixed(ratio\ =\ 1)}. If using the \texttt{plot} function
+included in the base library, one can use the \texttt{asp\ =\ 1}
+argument.
 
@@
 
-We can pass the cohesion matrix to the
-\texttt{plot\_community\_graphs()} function to view the relationship
-between points (Figure \ref{fig:fig2}). Notice in this plot the layout
-does not match that of the original data frame as seen in Figure
-\ref{fig:fig1}. Since our original data is two dimensional, it may be
-reasonable to use this as the layout. Figure \ref{fig:fig3} will make
-this update as well as update some of the aesthetics, such as employing
-more readable labels.
+We can pass the cohesion matrix to the \texttt{plot\_community\_graphs}
+function to view the relationship between points (Figure
+\ref{fig:fig2}). The function will also permit parameters that can be
+passed to \texttt{plot.igraph} via the \texttt{...} argument.
 
@@
 \begin{Sinput}
-plot_community_graphs(cohesion)
+plot_community_graphs(cohesion, 
+                      vertex.label.cex = 2,
+                      vertex.label.dist = 0.9)
 \end{Sinput}
@@
 
-The \texttt{layout} argument allows the user to pass a matrix to dictate
-the 2-dimensional layout of the graph. For example, if we wanted the
-graph to match the visualization displayed in Figure \ref{fig:fig1}, we
-could pass \texttt{as.matrix(df)}, or a matrix of the data frame
-\texttt{df} to the \texttt{layout} argument (Figure \ref{fig:fig3}).
-Additionally, the \texttt{plot\_community\_graphs()} function will also
-permit parameters that can be passed to the \texttt{plot.igraph()}
-function. We can pass arguments to the \texttt{plot.igraph} function via
-the \texttt{...} argument; for example to increase the vertex size and
-change the vertex label color, we can specify
+Notice in this plot the force-directed layout does not match that of the
+original data frame as seen in Figure \ref{fig:fig1}. Since our original
+data is two-dimensional, it may be reasonable to use the latter as the
+layout for plotting. Figure \ref{fig:fig3} includes this update as well
+as others addressing some of the aesthetics, such as employing more
+readable labels. The \texttt{layout} argument allows the user to pass a
+matrix to dictate the 2-dimensional layout of the graph. For example, if
+we wanted the graph to match the visualization displayed in Figure
+\ref{fig:fig1}, we could pass \texttt{as.matrix(df)} (a matrix of the
+data frame \texttt{df}) to the \texttt{layout} argument (Figure
+\ref{fig:fig3}). Here, we increase the vertex size and change the vertex
+label color, through the argument specifications
 \texttt{vertex.size\ =\ 100} and
 \texttt{vertex.label.color\ =\ "white"}. Additionally, to allow axes, we
-use \texttt{axes\ =\ TRUE}, and to put them back on the original scale
+use \texttt{axes\ =\ TRUE}, and to put these back on the original scale
 we set \texttt{rescale\ =\ FALSE}, resetting the axis limits using
@@
 
-We will demonstrate the utility of the \CRANpkg{pald} package in three
-illustrative examples.
+We will demonstrate the utility of the \CRANpkg{pald} package through
+several illustrative examples.
 
-\hypertarget{clustering-tissue-gene-expression-data}{%
-\subsubsection{Clustering tissue gene expression
-data}\label{clustering-tissue-gene-expression-data}}
+\hypertarget{community-analysis-for-tissue-gene-expression-data}{%
+\subsubsection{Community analysis for tissue gene expression
+data}\label{community-analysis-for-tissue-gene-expression-data}}
 
@@
 consisting of 22,215-dimensional gene expression data from 189 tissue
-samples. A \texttt{dist} object was created using this data set and is
-included in the \CRANpkg{pald} package in an object called
+samples. A (Euclidean) \texttt{dist} object was created using this data
+set and is included in the \CRANpkg{pald} package in an object called
 \texttt{tissue\_dist}.
@@
 
-The \texttt{community\_clusters()} function can be used to identify the
-community cluster corresponding to each tissue sample. Since the output
-is a data frame, we can summarize the result using commonly employed
-data analysis techniques. For demonstration purposes, we will use the
-\CRANpkg{dplyr} package to summarize the contribution of clusters.
+The \texttt{community\_clusters} function can be used to identify the
+community corresponding to each tissue sample. Since the output is a
+data frame, we can summarize the result using commonly employed R
+functions. For demonstration purposes, we will use the \CRANpkg{dplyr}
+package to summarize the community contributions.
 
@@
 community_clusters(tissue_cohesion) |>
-  dplyr::count(cluster, point)
+  dplyr::count(community, point)
 \end{Sinput}
 \begin{Soutput}
-#>    cluster       point  n
+#>    community       point  n
 #> 1          1 endometrium 15
@@
 
-From this, we can glean that Cluster 1 consists of two types of tissue,
-the kidney and endometrium. Cluster 2 is comprised of only the
-hippocampus.
-
-We can also display the relationships between tissue samples using the
-\texttt{plot\_community\_graphs()} function (Figure \ref{fig:fig4}). For
-clarity of the display, we show how to remove the labels using
-\texttt{show\_labels\ =\ FALSE}. We will instead color by the labels by
-passing these to the \texttt{vertex.color} parameter to the
+From this, we can glean, for instance, that Community 1 consists of two
+types of tissue, the kidney and endometrium, Community 2 is comprised of
+only the hippocampus, etc. More informatively, we can display
+relationships between tissue samples, both locally and globally through
+the \texttt{plot\_community\_graphs} function (Figure \ref{fig:fig4}).
+For clarity of the display, we show how to remove the labels using
+\texttt{show\_labels\ =\ FALSE}. We will instead color according to the
+labels by passing these to the \texttt{vertex.color} argument for the
 \texttt{plot.igraph} function (via the \texttt{...} argument).
@@
 the \texttt{edge\_width\_factor} and \texttt{emph\_strong} arguments to
-adjust the width of the lines between and within PaLD clusters.
+adjust the width of the lines between and within PaLD communities.
 
@@
 \includegraphics[width=1\linewidth]{fig5} \caption[Community cluster network for the tissue data]{Community cluster network for the tissue data.
-The line colors indicate the PaLD clusters, the point colors indicate the tissue classification.}\label{fig:fig4}
+The line colors indicate the PaLD communities; the point colors indicate the tissue classification.}\label{fig:fig4}
 \end{figure}
+\end{Schunk}
+
+Note that while Community 1 does contain kidney and endometrium tissues,
+the respective sub-communities are well-separated. Of the 357 strong
+inter-community ties within Community 1, only three are between the two
+sub-communities. In fact, these are the only strong ties between
+different tissue types in the community cluster network (of 1064 total);
+a summary of community strong ties can be produced via the following
+code. Note that \texttt{tissue\_graph\_strong} is an \CRANpkg{igraph}
+object corresponding to the graph displayed in Figure \ref{fig:fig4},
+restricted only to strong ties.
+
+\begin{Schunk}
+\begin{Sinput}
+tissue_graphs <- community_graphs(tissue_cohesion)
+tissue_graph_strong <- tissue_graphs[["G_strong"]]
+E <- igraph::get.edgelist(tissue_graph_strong)
+table(apply(E, 1, paste, collapse = ","))
+\end{Sinput}
+\begin{Soutput}
+#> 
+#>   cerebellum,cerebellum             colon,colon endometrium,endometrium 
+#>                     186                     241                      81 
+#> hippocampus,hippocampus      kidney,endometrium           kidney,kidney 
+#>                     191                       3                     273 
+#>             liver,liver       placenta,placenta 
+#>                      85                       4
+\end{Soutput}
 \end{Schunk}
@@
 
-This example performs a PaLD analysis on a data set from \citet{dyen92}
-that examines the relationship between 87 Indo-European languages from
-the perspective of cognates, coded using 2,655-dimensional binary
-vectors. A \texttt{dist} object was created from this data set and is
-included in the \CRANpkg{pald} package in an object called
-\texttt{cognate\_dist}.
+This example explors a data set from \citet{dyen92} that summarizes
+relationships between 87 Indo-European languages from the perspective of
+cognates, coded using 2,655-dimensional binary vectors. A \texttt{dist}
+object was created from this data set and is included in the
+\CRANpkg{pald} package in an object called \texttt{cognate\_dist}.
 
-This example will demonstrate how you can apply functions in the
+Here we will demonstrate how one can further apply functions in the
 \CRANpkg{igraph} package to objects output from the \CRANpkg{pald}
-package. We can first use the \texttt{cohesion\_matrix()} function to
-calculate the cohesion matrix and the \texttt{community\_graphs()}
+package. We can first use the \texttt{cohesion\_matrix} function to
+calculate the cohesion matrix and the \texttt{community\_graphs}
 function to create a list with the weighted community graph, the
@@
 
-We can then use the \texttt{neighbors()} function from the
+We can then use the \texttt{neighbors} function from the
 \CRANpkg{igraph} package to extract the strong neighbors in this graph.
-For example, if we wanted to extract all neighbors for the language
-``French'', we could run the following.
+For example, we can extract all neighbors for the language ``French'',
+via the followign code.
 
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from a461d26:
+#> + 8/87 vertices, named, from dc83183:
 #> [1] Italian         Ladin           Provencal       Walloon        
@@
 
-We can again use the \texttt{plot\_community\_graphs()} function to
+We can again use the \texttt{plot\_community\_graphs} function to
 visualize the community clusters (Figure \ref{fig:figlang}). One may
@@
 \item
-  \texttt{exdata1} is a data set consisting of 8 points to recreate
+  \texttt{exdata1} is a data set consisting of 8 points used to create
   Figure 1 in \citet{berenhaut2022social}
 \item
-  \texttt{exdata2} is a data set consisting of 16 points to recreate
+  \texttt{exdata2} is a data set consisting of 16 points used to create
   Figure 2 in \citet{berenhaut2022social}
 \item
-  \texttt{exdata3} is a data set consisting of 240 points to recreate
+  \texttt{exdata3} is a data set consisting of 240 points used to create
   Figure 4D in \citet{berenhaut2022social}
@@
 
-We will demonstrate how we can compare PaLD to two clustering methods:
-\emph{k}-means and hierarchical clustering. The code below calculates
-the cohesion matrix (\texttt{exdata\_cohesion}) as well as the clusters
-via PaLD (\texttt{exdata\_pald}), \emph{k}-means
-(\texttt{exdata\_kmeans}) and hierarchical clustering using complete
-linkage (\texttt{exdata\_hclust}).
+We will contrast community analysis via PaLD to standard culuster
+analysis by considering two common clustering methods. The code below
+calculates the cohesion matrix (\texttt{exdata\_cohesion}) as well as
+the community clusters obtained via PaLD (\texttt{exdata\_pald}), along
+with clusters obtained via \emph{k}-means (\texttt{exdata\_kmeans}) and
+hierarchical clustering using complete linkage
+(\texttt{exdata\_hclust}).
 
@@
 
-We can compare this to the clustering generated by \emph{k}-means and
-hierarchical clustering (Figure \ref{fig:fig5}).
+The information is displayed in Figure \ref{fig:fig5}.
 
@@
 \begin{figure}
-\includegraphics{dagostino-mcgowan_files/figure-latex/fig5-1} \caption[PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al]{PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al.
-(2022)) compared to k-means and hierarchical clustering with k = 8.}\label{fig:fig5}
+\includegraphics{dagostino-mcgowan_files/figure-latex/fig5-1} \caption[PaLD, k-means, and hierarchical 8-clustering of randomly generated example data (from \citet{berenhaut2022social}]{PaLD, k-means, and hierarchical 8-clustering of randomly generated example data (from \citet{berenhaut2022social}; Figure 4D).}\label{fig:fig5}
 \end{figure}
@@
 discussion, and theoretical results. Note that the PaLD algorithm is
-able to detect the eight natural groups within the data without the use
+able to detect the eight natural groups within the data (along with
+inter- and intra-community structure not displayed here) without the use
 of any additional inputs (e.g., number of clusters) nor optimization
 criteria. Despite the user input of the ``correct'' number of clusters
-(i.e., \(k = 8\)) both \emph{k}-means and hierarchical clustering did
-not give the desired result.
+(i.e., \(k = 8\)) both \emph{k}-means and hierarchical clustering do not
+provide the desired result.
 
@@
 
-\textbf{TODO intro to data here} \citep{muthukrishna2020beyond}
-
-First, we will create the cohesion matrix using the \texttt{cultures}
-data, provided in \CRANpkg{pald}, and proceed to plot the community
-structure.
+In this example we perform a PaLD analysis for cultural distances
+obtained in \citet{muthukrishna2020beyond} from two recent waves of the
+World Values Survey (2005 to 2009 and 2010 to 2014; see
+\citet{inglehart2014world}). Distances are computed using the cultural
+fixation index (CFST), which is a measure built on the framework of
+fixation indices from population biology
+(\citet{bell2009culture,cavalli1994history}). Recall that the foundation
+of PaLD in within-triplet comparisons allows for the employment of
+application-dependent and non-Euclidean measures of dissimilarity. The
+\texttt{dist} object is included in the \CRANpkg{pald} package
+(\texttt{cultures}). We will first create the cohesion matrix using the
+\texttt{dist} object \texttt{cultures}, and proceed to plot the
+community graph.
 
@@
 
-This example demonstrates key features of the PaLD approach that
-distinguish it from commonly used clustering techniques. In addition to
-viewing the community structure as seen in Figure \ref{fig:figculture},
-we can also visualize a two dimensional plot of cohesion versus distance
-for each of the included distance pairs using the
-\texttt{dist\_cohesion\_plot()} function, as seen below (Figure
+In addition to viewing the local and global community structure as seen
+in Figure \ref{fig:figculture}, the \CRANpkg{pald} package allows for a
+two-dimensional display of cohesion against distance for the data, via
+the \texttt{dist\_cohesion\_plot} function, as seen below (Figure
 \ref{fig:figco}).
@@
 \begin{Sinput}
-dist_cohesion_plot(cultures)
+dist_cohesion_plot(cultures, mutual = TRUE)
 \end{Sinput}
 \begin{figure}
-\includegraphics{dagostino-mcgowan_files/figure-latex/figco-1} \caption[Distance versus cohesion for each TODO The identified communties are shown via the 5 colors -- these map to the same clusters idenfitied in Figure 7]{Distance versus cohesion for each TODO The identified communties are shown via the 5 colors -- these map to the same clusters idenfitied in Figure 7.}\label{fig:figco}
+\includegraphics{dagostino-mcgowan_files/figure-latex/figco-1} \caption{A plot of cohesion versus distance for the data.
+The identified communities are colored as in Figure \ref{fig:figculture}.}\label{fig:figco}
 \end{figure}
@@
 
-Notice here the magnitude of the distances within each of the identified
-communities varies substantially between regions, yet the cohesion
-remains strong. In other words, focusing particularly on identified
-community consisting of regions in the United States (light purple)
-compared to the community consisting of regions in India (turquoise) the
-strong cohesions are similar (Figure \ref{fig:figculture}), yet the
-distances themselves are markedly different (Figure \ref{fig:figco}). In
-fact, the largest distance between regions in the Unites States is
-smaller than that between any two regions in India (Figure
-\ref{fig:figco}). Classic distance-based clustering techniques tend to
-be unable to maintain this distinction often breaking India into several
-smaller communities. Additionally, currently available techniques
-require specification of parameters, as seen in the previous section.
-For further discussion of the cultural distance data in relation to
-community analysis see \citet{berenhaut2022social}.
+Notice here that the magnitude of the distances within each of the
+identified communities varies substantially between regions; in fact,
+the most disparate two regions in the United States (at distance
+\(\approx\) 0.027) are far closer than the two most similar in India (at
+distance \(\approx\) 0.043). Despite this, India remains a cohesive
+whole, and locally disparate regions in the United States such as East
+South Central and California are not strongly cohesive. For discussion
+of subtleties in local density (see \citet{berenhaut2022social}.
+Additionally, currently available techniques require specification of
+parameters, as seen in the previous section. For further discussion of
+the cultural distance data in relation to community analysis see
+\citet{berenhaut2022social}.
 
@@
 This paper introduces the \CRANpkg{pald} package, demonstrating its
-utility for providing parameter-free clustering which can easily be
-implemented for a variety of data sets. We provide example code as well
-as compare the method to commonly used clustering techniques,
-\emph{k}-means and hierarchical clustering.
+utility for providing parameter-free community analysis which can easily
+be implemented for a variety of data sets, supplementing results from
+other methods for clustering, embedding, nearest neighbors, etc. Example
+code is provided advantages over other techniques are highlighted.
 

### 2023-09-28 2023-r-journal/dagostino-mcgowan.R

@@ -1,0 +1,272 @@
+# Generated by `rjournal_article()` using `knitr::purl()`: do not edit by hand
+# Please edit dagostino-mcgowan.Rmd to modify this file
+
+## ---- echo = FALSE------------------------------------------------------------
+knitr::opts_chunk$set(fig.height = 4, message = FALSE, warning = FALSE)
+
+
+## -----------------------------------------------------------------------------
+library(pald)
+df <- data.frame(
+  x1 = c(6, 8, 8, 16, 4, 14),
+  x2 = c(5, 4, 10, 8, 4, 10)
+)
+rownames(df) <- c("A", "B", "C", "D", "E", "F")
+
+
+## -----------------------------------------------------------------------------
+d <- dist(df)
+
+
+## -----------------------------------------------------------------------------
+d <- dist(df)
+cohesion_matrix(d)
+
+
+## -----------------------------------------------------------------------------
+df |>
+  dist() |>
+  cohesion_matrix()
+
+
+## -----------------------------------------------------------------------------
+df |>
+  dist() |>
+  cohesion_matrix() -> cohesion
+
+
+## -----------------------------------------------------------------------------
+local_depths(cohesion)
+
+
+## -----------------------------------------------------------------------------
+strong_threshold(cohesion)
+
+
+## -----------------------------------------------------------------------------
+cohesion_strong(cohesion)
+
+
+## -----------------------------------------------------------------------------
+graphs <- community_graphs(cohesion)
+graphs[["G_strong"]]
+
+
+## -----------------------------------------------------------------------------
+any_isolated(cohesion)
+
+
+## -----------------------------------------------------------------------------
+community_clusters(cohesion)
+
+
+## ----fig1, fig.cap = "Visualization of the points from the data frame `df`", message = FALSE, warning = FALSE----
+library(ggplot2)
+ggplot(df, aes(x1, x2)) +
+  geom_text(label = rownames(df)) + 
+  coord_fixed(ratio = 1) + 
+  xlim(c(4, 16)) + 
+  ylim(c(4, 16))
+
+
+## ----fig2, fig.cap = "PaLD graph displaying the relationship between the points in data frame `df`"----
+plot_community_graphs(cohesion, 
+                      vertex.label.cex = 2,
+                      vertex.label.dist = 0.9)
+
+
[…]

### 2023-09-28 2023-r-journal/dagostino-mcgowan.tex
update fig

@@
 \begin{Soutput}
-#> IGRAPH 198a54e UNW- 6 3 -- 
+#> IGRAPH 90d60fd UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from 198a54e (vertex names):
+#> + edges from 90d60fd (vertex names):
 #> [1] A--B A--E D--F
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from dc83183:
+#> + 8/87 vertices, named, from aab3808:
 #> [1] Italian         Ladin           Provencal       Walloon        
@@
 
-exdata_pald <- community_clusters(exdata_cohesion)$cluster
+exdata_pald <- community_clusters(exdata_cohesion)$community
 

### 2023-09-28 2023-r-journal/dagostino-mcgowan.R

@@ -193,7 +193,7 @@ exdata_cohesion <- exdata3 |>
   dist() |>
   cohesion_matrix()
 
-exdata_pald <- community_clusters(exdata_cohesion)$cluster
+exdata_pald <- community_clusters(exdata_cohesion)$community
 
 exdata_kmeans <- kmeans(exdata3, 8)$cluster
 

### 2025-02-03 2023-r-journal/dagostino-mcgowan.Rmd
update fig
update

@@
 
-The `community_clusters` function can be used to identify the community corresponding to each tissue sample.
-Since the output is a data frame, we can summarize the result using commonly employed R functions.
-For demonstration purposes, we will use the \CRANpkg{dplyr} package to summarize the community contributions.
-
-```{r}
-community_clusters(tissue_cohesion) |>
-  dplyr::count(community, point)
-```
-
-From this, we can glean, for instance, that Community 1 consists of two types of tissue, the kidney and endometrium, Community 2 is comprised of only the hippocampus, etc.
-More informatively, we can display relationships between tissue samples, both locally and globally through the `plot_community_graphs` function  (Figure \ref{fig:fig4}).
+We can display relationships between tissue samples, both locally and globally through the `plot_community_graphs` function  (Figure \ref{fig:fig4}).
 For clarity of the display, we show how to remove the labels using `show_labels = FALSE`.
@@
 
-Note that while Community 1 does contain kidney and endometrium tissues, the respective sub-communities are well-separated.
-Of the 357 strong inter-community ties within Community 1, only three are between the two sub-communities.
-In fact, these are the only strong ties between different tissue types in the community cluster network (of 1064 total); a summary of community strong ties can be produced via the following code.
+A summary of community strong ties can be produced via the following code.
 Note that `tissue_graph_strong` is an \CRANpkg{igraph} object corresponding to the graph displayed in Figure \ref{fig:fig4}, restricted only to strong ties.
+
+<!-- The `community_clusters` function can be used to identify the community corresponding to each tissue sample.
+Since the output is a data frame, we can summarize the result using commonly employed R functions.
+For demonstration purposes, we will use the \CRANpkg{dplyr} package to summarize the community contributions.
+-->
+
+<!-- ```{r} -->
+<!-- community_clusters(tissue_cohesion) |> -->
+<!--   dplyr::count(community, point) -->
+<!-- ``` -->
+
+<!-- From this, we can glean, for instance, that Community 1 consists of two types of tissue, the kidney and endometrium, Community 2 is comprised of only the hippocampus, etc.
+More informatively,  -->
 
@@
 E <- igraph::get.edgelist(tissue_graph_strong)
-table(apply(E, 1, paste, collapse = ","))
+tissue_strong_ties <- data.frame(
+  strong_ties = apply(E, 1, paste, collapse = ",")
+)
+tissue_strong_ties |>
+  dplyr::count(strong_ties) |>
+  dplyr::arrange(-n)
 ```
+
+Note that there are only three strong ties between different tissue types (kidney and endometrium) in the community cluster network of 1,064 strong ties total.
 
@@
 
-This example explors a data set from \citet{dyen92} that summarizes relationships between 87 Indo-European languages from the perspective of cognates, coded using 2,655-dimensional binary vectors.
+This example explores a data set from \citet{dyen92} that summarizes relationships between 87 Indo-European languages from the perspective of cognates, coded using 2,655-dimensional binary vectors.
 A `dist` object was created from this data set and is included in the \CRANpkg{pald} package in an object called `cognate_dist`.
@@
 We can then use the `neighbors` function from the \CRANpkg{igraph} package to extract the strong neighbors in this graph.
-For example, we can extract all neighbors for the language "French", via the followign code.
+For example, we can extract all neighbors for the language "French", via the following code.
 
@@
 
-Similarly, we can print the associated neighborhood weights by subsetting the cohesion matrix.
+Similarly, we can sort and print the associated neighborhood weights by subsetting the cohesion matrix.
 
 ```{r}
-cognate_cohesion["French", french_neighbors]
+cognate_cohesion["French", french_neighbors] |>
+  sort(decreasing = TRUE)
 ```
@@
 
-We will contrast community analysis via PaLD to standard culuster analysis by considering two common clustering methods.
-The code below calculates the cohesion matrix (`exdata_cohesion`) as well as the community clusters obtained via PaLD (`exdata_pald`), along with clusters obtained via *k*-means (`exdata_kmeans`) and hierarchical clustering using complete linkage (`exdata_hclust`).
+We will contrast community analysis via PaLD to standard cluster analysis by considering two common clustering methods.
+The code below calculates the cohesion matrix (`exdata_cohesion`) as well as the community clusters obtained via PaLD (`exdata_pald`), along with $k=8$ clusters obtained via *k*-means (`exdata_kmeans`) and hierarchical clustering using complete linkage (`exdata_hclust`).
+Recall that there is no need to determine values for extraneous inputs in the case of PaLD (e.g.
+the number of clusters, as is necessary to specify for *k*-means and hierarchical clustering).
 
@@
 
-This paper introduces the \CRANpkg{pald} package, demonstrating its utility for providing parameter-free community analysis which can easily be implemented for a variety of data sets, supplementing results from other methods for clustering, embedding, nearest neighbors, etc.
-Example code is provided advantages over other techniques are highlighted.
+This paper introduces the \CRANpkg{pald} package, demonstrating its utility for providing novel parameter-free community analysis which can easily be implemented for a variety of data sets, supplementing results from other methods for clustering, embedding, nearest neighbors, etc.
+Example code is provided along with discussion in the context of other commonly used R-based approaches.
 

### 2025-02-03 2023-r-journal/dagostino-mcgowan.tex
update

@@
 \begin{Soutput}
-#> IGRAPH 90d60fd UNW- 6 3 -- 
+#> IGRAPH 5de0378 UNW- 6 3 --
 #> + attr: name (v/c), weight (e/n)
-#> + edges from 90d60fd (vertex names):
+#> + edges from 5de0378 (vertex names):
 #> [1] A--B A--E D--F
@@
 \end{Sinput}
-\begin{figure}
-\includegraphics{dagostino-mcgowan_files/figure-latex/fig1-1} \caption[Visualization of the points from the data frame `df`]{Visualization of the points from the data frame `df`}\label{fig:fig1}
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=5in]{dagostino-mcgowan_files/figure-latex/fig1-1} \caption[Visualization of the points from the data frame `df`]{Visualization of the points from the data frame `df`}\label{fig:fig1}
+\end{center}
 \end{figure}
@@
 \end{Sinput}
-\begin{figure}
-\includegraphics{dagostino-mcgowan_files/figure-latex/fig2-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`]{PaLD graph displaying the relationship between the points in data frame `df`}\label{fig:fig2}
+\begin{figure}[H]
+\includegraphics[width=5in,trim=0in 1in 0in 0.7in,clip]{dagostino-mcgowan_files/figure-latex/fig2-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`]{PaLD graph displaying the relationship between the points in data frame `df`}\label{fig:fig2}
 \end{figure}
@@
 \end{Sinput}
-\begin{figure}
-\includegraphics{dagostino-mcgowan_files/figure-latex/fig3-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1]{PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1}\label{fig:fig3}
+\begin{figure}[H]
+\includegraphics[width=5in,trim=0in 0.6in 0in 0.8in,clip]{dagostino-mcgowan_files/figure-latex/fig3-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1]{PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1}\label{fig:fig3}
 \end{figure}
@@
 
-The \texttt{community\_clusters} function can be used to identify the
-community corresponding to each tissue sample. Since the output is a
-data frame, we can summarize the result using commonly employed R
-functions. For demonstration purposes, we will use the \CRANpkg{dplyr}
-package to summarize the community contributions.
-
-\begin{Schunk}
-\begin{Sinput}
-community_clusters(tissue_cohesion) |>
-  dplyr::count(community, point)
-\end{Sinput}
-\begin{Soutput}
-#>    community       point  n
-#> 1          1 endometrium 15
-#> 2          1      kidney 39
-#> 3          2 hippocampus 31
-#> 4          3  cerebellum 26
-#> 5          4  cerebellum  1
-#> 6          5       colon 33
-#> 7          6       colon  1
-#> 8          7       liver  7
-#> 9          8  cerebellum  1
-#> 10         9       liver 17
-#> 11        10  cerebellum  2
-#> 12        11       liver  2
-#> 13        12  cerebellum  1
-#> 14        13  cerebellum  4
-#> 15        14  cerebellum  2
-#> 16        15  cerebellum  1
-#> 17        16    placenta  2
-#> 18        17    placenta  1
-#> 19        18    placenta  3
-\end{Soutput}
-\end{Schunk}
-
-From this, we can glean, for instance, that Community 1 consists of two
-types of tissue, the kidney and endometrium, Community 2 is comprised of
-only the hippocampus, etc. More informatively, we can display
-relationships between tissue samples, both locally and globally through
-the \texttt{plot\_community\_graphs} function (Figure \ref{fig:fig4}).
-For clarity of the display, we show how to remove the labels using
-\texttt{show\_labels\ =\ FALSE}. We will instead color according to the
-labels by passing these to the \texttt{vertex.color} argument for the
-\texttt{plot.igraph} function (via the \texttt{...} argument).
-Similarly, we can add a legend using the \texttt{legend()} function, as
-you would for an \CRANpkg{igraph} visualization. Additionally, we use
-the \texttt{edge\_width\_factor} and \texttt{emph\_strong} arguments to
-adjust the width of the lines between and within PaLD communities.
+We can display relationships between tissue samples, both locally and
+globally through the \texttt{plot\_community\_graphs} function (Figure
+\ref{fig:fig4}). For clarity of the display, we show how to remove the
+labels using \texttt{show\_labels\ =\ FALSE}. We will instead color
+according to the labels by passing these to the \texttt{vertex.color}
+argument for the \texttt{plot.igraph} function (via the \texttt{...}
+argument). Similarly, we can add a legend using the \texttt{legend()}
+function, as you would for an \CRANpkg{igraph} visualization.
+Additionally, we use the \texttt{edge\_width\_factor} and
+\texttt{emph\_strong} arguments to adjust the width of the lines between
+and within PaLD communities.
 
@@
 \begin{Schunk}
-\begin{figure}
-\includegraphics[width=1\linewidth]{fig5} \caption[Community cluster network for the tissue data]{Community cluster network for the tissue data.
+\begin{figure}[H]
+\includegraphics[width=5.5in,trim=0in 5in 0in 3in,clip]{fig5} \caption[Community cluster network for the tissue data]{Community cluster network for the tissue data.
 The line colors indicate the PaLD communities; the point colors indicate the tissue classification.}\label{fig:fig4}
@@
 
-Note that while Community 1 does contain kidney and endometrium tissues,
-the respective sub-communities are well-separated. Of the 357 strong
-inter-community ties within Community 1, only three are between the two
-sub-communities. In fact, these are the only strong ties between
-different tissue types in the community cluster network (of 1064 total);
-a summary of community strong ties can be produced via the following
+A summary of community strong ties can be produced via the following
 code. Note that \texttt{tissue\_graph\_strong} is an \CRANpkg{igraph}
@@
 E <- igraph::get.edgelist(tissue_graph_strong)
-table(apply(E, 1, paste, collapse = ","))
+tissue_strong_ties <- data.frame(
+  strong_ties = apply(E, 1, paste, collapse = ",")
+)
+tissue_strong_ties |>
+  dplyr::count(strong_ties) |>
+  dplyr::arrange(-n)
 \end{Sinput}
 \begin{Soutput}
-#> 
-#>   cerebellum,cerebellum             colon,colon endometrium,endometrium 
-#>                     186                     241                      81 
-#> hippocampus,hippocampus      kidney,endometrium           kidney,kidney 
-#>                     191                       3                     273 
-#>             liver,liver       placenta,placenta 
-#>                      85                       4
+#>               strong_ties   n
+#> 1           kidney,kidney 273
+#> 2             colon,colon 241
+#> 3 hippocampus,hippocampus 191
+#> 4   cerebellum,cerebellum 186
+#> 5             liver,liver  85
+#> 6 endometrium,endometrium  81
+#> 7       placenta,placenta   4
+#> 8      kidney,endometrium   3
 \end{Soutput}
 \end{Schunk}
+
+Note that there are only three strong ties between different tissue
+types (kidney and endometrium) in the community cluster network of 1,064
+strong ties total.
 
@@
 
-This example explors a data set from \citet{dyen92} that summarizes
+This example explores a data set from \citet{dyen92} that summarizes
 relationships between 87 Indo-European languages from the perspective of
@@
 For example, we can extract all neighbors for the language ``French'',
-via the followign code.
+via the following code.
 
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from aab3808:
+#> + 8/87 vertices, named, from b09505b:
 #> [1] Italian         Ladin           Provencal       Walloon
@@
 
-Similarly, we can print the associated neighborhood weights by
+Similarly, we can sort and print the associated neighborhood weights by
 subsetting the cohesion matrix.
@@
 \begin{Sinput}
-cognate_cohesion["French", french_neighbors]
+cognate_cohesion["French", french_neighbors] |>
+  sort(decreasing = TRUE)
 \end{Sinput}
 \begin{Soutput}
-#>         Italian           Ladin       Provencal         Walloon French_Creole_C 
-#>      0.01997696      0.02094596      0.02871174      0.03258771      0.02406057 
-#> French_Creole_D         Spanish         Catalan 
-#>      0.02406057      0.01679733      0.01859688
+#>         Walloon       Provencal French_Creole_C French_Creole_D           Ladin
+#>      0.03258771      0.02871174      0.02406057      0.02406057      0.02094596
+#>         Italian         Catalan         Spanish
+#>      0.01997696      0.01859688      0.01679733
 \end{Soutput}
@@
 \begin{Schunk}
-\begin{figure}
-\includegraphics[width=1\linewidth]{fig6} \caption[Community structure for 87 Indo-European languages, which employs cognate information that was coded via 2,665-dimensional binary vectors]{Community structure for 87 Indo-European languages, which employs cognate information that was coded via 2,665-dimensional binary vectors.
+\begin{figure}[H]
+\includegraphics[width=5.5in,trim=0in 5in 0in 1in,clip]{fig6} \caption[Community structure for 87 Indo-European languages, which employs cognate information that was coded via 2,665-dimensional binary vectors]{Community structure for 87 Indo-European languages, which employs cognate information that was coded via 2,665-dimensional binary vectors.
 Commonly identifiable language clusters arise along with informative inter- and intra-cluster structure.
@@
 
-We will contrast community analysis via PaLD to standard culuster
+We will contrast community analysis via PaLD to standard cluster
 analysis by considering two common clustering methods. The code below
@@
 the community clusters obtained via PaLD (\texttt{exdata\_pald}), along
-with clusters obtained via \emph{k}-means (\texttt{exdata\_kmeans}) and
-hierarchical clustering using complete linkage
-(\texttt{exdata\_hclust}).
+with \(k=8\) clusters obtained via \emph{k}-means
+(\texttt{exdata\_kmeans}) and hierarchical clustering using complete
+linkage (\texttt{exdata\_hclust}). Recall that there is no need to
+determine values for extraneous inputs in the case of PaLD (e.g.~the
+number of clusters, as is necessary to specify for \emph{k}-means and
+hierarchical clustering).
 
@@
 \end{Sinput}
-\begin{figure}
-\includegraphics{dagostino-mcgowan_files/figure-latex/fig5-1} \caption[PaLD, k-means, and hierarchical 8-clustering of randomly generated example data (from \citet{berenhaut2022social}]{PaLD, k-means, and hierarchical 8-clustering of randomly generated example data (from \citet{berenhaut2022social}; Figure 4D).}\label{fig:fig5}
+\begin{figure}[H]
+\includegraphics[width=5.5in,trim=0in 1in 0in 0.8in,clip]{dagostino-mcgowan_files/figure-latex/fig5-1} \caption[PaLD, k-means, and hierarchical 8-clustering of randomly generated example data (from \citet{berenhaut2022social}]{PaLD, k-means, and hierarchical 8-clustering of randomly generated example data (from \citet{berenhaut2022social}; Figure 4D).}\label{fig:fig5}
 \end{figure}
@@
 \begin{Schunk}
-\begin{figure}
-\includegraphics[width=1\linewidth]{fig7} \caption[Community structure for cultural distance data]{Community structure for cultural distance data.}\label{fig:figculture}
+\begin{figure}[H]
+\includegraphics[width=5.5in,trim=0in 5in 0in 3in,clip]{fig7} \caption[Community structure for cultural distance data]{Community structure for cultural distance data.}\label{fig:figculture}
 \end{figure}
@@
 \end{Sinput}
-\begin{figure}
-\includegraphics{dagostino-mcgowan_files/figure-latex/figco-1} \caption{A plot of cohesion versus distance for the data.
+\begin{figure}[H]
+\includegraphics[width=5.5in,trim=0in 0.5in 0in 0.5in,clip]{dagostino-mcgowan_files/figure-latex/figco-1} \caption{A plot of cohesion versus distance for the data.
 The identified communities are colored as in Figure \ref{fig:figculture}.}\label{fig:figco}
@@
 This paper introduces the \CRANpkg{pald} package, demonstrating its
-utility for providing parameter-free community analysis which can easily
-be implemented for a variety of data sets, supplementing results from
-other methods for clustering, embedding, nearest neighbors, etc. Example
-code is provided advantages over other techniques are highlighted.
+utility for providing novel parameter-free community analysis which can
+easily be implemented for a variety of data sets, supplementing results
+from other methods for clustering, embedding, nearest neighbors, etc.
+Example code is provided along with discussion in the context of other
+commonly used R-based approaches.
 

### 2025-02-03 2023-r-journal/dagostino-mcgowan.R

@@ -93,11 +93,6 @@ plot_community_graphs(cohesion,
 tissue_cohesion <- cohesion_matrix(tissue_dist)
 
 
-## -----------------------------------------------------------------------------
-community_clusters(tissue_cohesion) |>
-  dplyr::count(community, point)
-
-
 ## ---- eval = FALSE------------------------------------------------------------
 #> labels <- rownames(tissue_cohesion)
 #> plot_community_graphs(tissue_cohesion,
@@ -140,7 +135,12 @@ knitr::include_graphics("fig5.png")
 tissue_graphs <- community_graphs(tissue_cohesion)
 tissue_graph_strong <- tissue_graphs[["G_strong"]]
 E <- igraph::get.edgelist(tissue_graph_strong)
-table(apply(E, 1, paste, collapse = ","))
+tissue_strong_ties <- data.frame(
+  strong_ties = apply(E, 1, paste, collapse = ",")
+)
+tissue_strong_ties |>
+  dplyr::count(strong_ties) |>
+  dplyr::arrange(-n)
 
 
 ## -----------------------------------------------------------------------------
@@ -156,7 +156,8 @@ french_neighbors
 
 
 ## -----------------------------------------------------------------------------
-cognate_cohesion["French", french_neighbors]
+cognate_cohesion["French", french_neighbors] |>
+  sort(decreasing = TRUE)
 
 
 ## ---- eval = FALSE------------------------------------------------------------

### 2025-02-03 2025-r-journal/dagostino-mcgowan.Rmd
add recent revision

--- 2025-02-03 2023-r-journal/dagostino-mcgowan.Rmd
@@
 
-In addition, the package provides a number of pertinent example data sets which may be used to demonstrate community analysis, including a synthetic data set of two-dimensional points created by \citet{gionis1clustering} to demonstrate aggregation, clustering data generated from the scikit-learn Python package \citep{pedregosa2011scikit}, data describing cognate relationships between words across 87 Indo-European languages \citep{dyen92}, data compiled by \cite{tissue} of tissue gene expressions, data emplying information from the World Values Survey \citep{inglehart2014world} on cultural values regarding family, religion, education, and institutions for several regions \citep{muthukrishna2020beyond}, and three example data sets generated for the \citet{berenhaut2022social} paper.
+In addition, the package provides a number of pertinent example data sets which may be used to demonstrate community analysis, including a synthetic data set of two-dimensional points created by \citet{gionis1clustering} to demonstrate aggregation, clustering data generated from the scikit-learn Python package \citep{pedregosa2011scikit}, data describing cognate relationships between words across 87 Indo-European languages \citep{dyen92}, data compiled by \cite{tissue} of tissue gene expressions, data employing information from the World Values Survey \citep{inglehart2014world} on cultural values regarding family, religion, education, and institutions for several regions \citep{muthukrishna2020beyond}, and three example data sets generated for the \citet{berenhaut2022social} paper.
 
@@
 ```{r}
-d <- dist(df)
 cohesion_matrix(d)
@@
 
+```{r, echo=FALSE}
+options(width = 20)
+```
+
 ```{r}
@@
 tissue_strong_ties |>
-  dplyr::count(strong_ties) |>
-  dplyr::arrange(-n)
+  table() |>
+  sort(decreasing = TRUE)
 ```

### 2025-02-03 2025-r-journal/dagostino-mcgowan.tex
add recent revision

--- 2025-02-03 2023-r-journal/dagostino-mcgowan.tex
@@
 between words across 87 Indo-European languages \citep{dyen92}, data
-compiled by \cite{tissue} of tissue gene expressions, data emplying
+compiled by \cite{tissue} of tissue gene expressions, data employing
 information from the World Values Survey \citep{inglehart2014world} on
@@
 \begin{Sinput}
-d <- dist(df)
 cohesion_matrix(d)
@@
 \begin{Soutput}
-#> IGRAPH 5de0378 UNW- 6 3 --
+#> IGRAPH cc642b5 UNW- 6 3 -- 
 #> + attr: name (v/c), weight (e/n)
-#> + edges from 5de0378 (vertex names):
+#> + edges from cc642b5 (vertex names):
 #> [1] A--B A--E D--F
@@
 \end{Sinput}
-\begin{figure}[H]
-\begin{center}
-\includegraphics[width=5in]{dagostino-mcgowan_files/figure-latex/fig1-1} \caption[Visualization of the points from the data frame `df`]{Visualization of the points from the data frame `df`}\label{fig:fig1}
-\end{center}
+\begin{figure}
+\includegraphics{dagostino-mcgowan_files/figure-latex/fig1-1} \caption[Visualization of the points from the data frame `df`]{Visualization of the points from the data frame `df`}\label{fig:fig1}
 \end{figure}
@@
 \end{Sinput}
-\begin{figure}[H]
-\includegraphics[width=5in,trim=0in 1in 0in 0.7in,clip]{dagostino-mcgowan_files/figure-latex/fig2-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`]{PaLD graph displaying the relationship between the points in data frame `df`}\label{fig:fig2}
+\begin{figure}
+\includegraphics{dagostino-mcgowan_files/figure-latex/fig2-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`]{PaLD graph displaying the relationship between the points in data frame `df`}\label{fig:fig2}
 \end{figure}
@@
 \end{Sinput}
-\begin{figure}[H]
-\includegraphics[width=5in,trim=0in 0.6in 0in 0.8in,clip]{dagostino-mcgowan_files/figure-latex/fig3-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1]{PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1}\label{fig:fig3}
+\begin{figure}
+\includegraphics{dagostino-mcgowan_files/figure-latex/fig3-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1]{PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1}\label{fig:fig3}
 \end{figure}
@@
 \begin{Schunk}
-\begin{figure}[H]
-\includegraphics[width=5.5in,trim=0in 5in 0in 3in,clip]{fig5} \caption[Community cluster network for the tissue data]{Community cluster network for the tissue data.
+\begin{figure}
+\includegraphics[width=1\linewidth]{fig5} \caption[Community cluster network for the tissue data]{Community cluster network for the tissue data.
 The line colors indicate the PaLD communities; the point colors indicate the tissue classification.}\label{fig:fig4}
@@
 tissue_strong_ties |>
-  dplyr::count(strong_ties) |>
-  dplyr::arrange(-n)
+  table() |>
+  sort(decreasing = TRUE)
 \end{Sinput}
 \begin{Soutput}
-#>               strong_ties   n
-#> 1           kidney,kidney 273
-#> 2             colon,colon 241
-#> 3 hippocampus,hippocampus 191
-#> 4   cerebellum,cerebellum 186
-#> 5             liver,liver  85
-#> 6 endometrium,endometrium  81
-#> 7       placenta,placenta   4
-#> 8      kidney,endometrium   3
+#> strong_ties
+#>           kidney,kidney 
+#>                     273 
+#>             colon,colon 
+#>                     241 
+#> hippocampus,hippocampus 
+#>                     191 
+#>   cerebellum,cerebellum 
+#>                     186 
+#>             liver,liver 
+#>                      85 
+#> endometrium,endometrium 
+#>                      81 
+#>       placenta,placenta 
+#>                       4 
+#>      kidney,endometrium 
+#>                       3
 \end{Soutput}
@@
 \begin{Soutput}
-#> + 8/87 vertices, named, from b09505b:
-#> [1] Italian         Ladin           Provencal       Walloon
-#> [5] French_Creole_C French_Creole_D Spanish         Catalan
+#> + 8/87 vertices, named, from 4454cb9:
+#> [1] Italian        
+#> [2] Ladin          
+#> [3] Provencal      
+#> [4] Walloon        
+#> [5] French_Creole_C
+#> [6] French_Creole_D
+#> [7] Spanish        
+#> [8] Catalan
 \end{Soutput}
@@
 \begin{Soutput}
-#>         Walloon       Provencal French_Creole_C French_Creole_D           Ladin
-#>      0.03258771      0.02871174      0.02406057      0.02406057      0.02094596
-#>         Italian         Catalan         Spanish
-#>      0.01997696      0.01859688      0.01679733
+#>         Walloon 
+#>      0.03258771 
+#>       Provencal 
+#>      0.02871174 
+#> French_Creole_C 
+#>      0.02406057 
+#> French_Creole_D 
+#>      0.02406057 
+#>           Ladin 
+#>      0.02094596 
+#>         Italian 
+#>      0.01997696 
+#>         Catalan 
+#>      0.01859688 
+#>         Spanish 
+#>      0.01679733
 \end{Soutput}
@@
 \begin{Schunk}
-\begin{figure}[H]
-\includegraphics[width=5.5in,trim=0in 5in 0in 1in,clip]{fig6} \caption[Community structure for 87 Indo-European languages, which employs cognate information that was coded via 2,665-dimensional binary vectors]{Community structure for 87 Indo-European languages, which employs cognate information that was coded via 2,665-dimensional binary vectors.
+\begin{figure}
+\includegraphics[width=1\linewidth]{fig6} \caption[Community structure for 87 Indo-European languages, which employs cognate information that was coded via 2,665-dimensional binary vectors]{Community structure for 87 Indo-European languages, which employs cognate information that was coded via 2,665-dimensional binary vectors.
 Commonly identifiable language clusters arise along with informative inter- and intra-cluster structure.
@@
 \end{Sinput}
-\begin{figure}[H]
-\includegraphics[width=5.5in,trim=0in 1in 0in 0.8in,clip]{dagostino-mcgowan_files/figure-latex/fig5-1} \caption[PaLD, k-means, and hierarchical 8-clustering of randomly generated example data (from \citet{berenhaut2022social}]{PaLD, k-means, and hierarchical 8-clustering of randomly generated example data (from \citet{berenhaut2022social}; Figure 4D).}\label{fig:fig5}
+\begin{figure}
+\includegraphics{dagostino-mcgowan_files/figure-latex/fig5-1} \caption[PaLD, k-means, and hierarchical 8-clustering of randomly generated example data (from \citet{berenhaut2022social}]{PaLD, k-means, and hierarchical 8-clustering of randomly generated example data (from \citet{berenhaut2022social}; Figure 4D).}\label{fig:fig5}
 \end{figure}
@@
 \begin{Schunk}
-\begin{figure}[H]
-\includegraphics[width=5.5in,trim=0in 5in 0in 3in,clip]{fig7} \caption[Community structure for cultural distance data]{Community structure for cultural distance data.}\label{fig:figculture}
+\begin{figure}
+\includegraphics[width=1\linewidth]{fig7} \caption[Community structure for cultural distance data]{Community structure for cultural distance data.}\label{fig:figculture}
 \end{figure}
@@
 \end{Sinput}
-\begin{figure}[H]
-\includegraphics[width=5.5in,trim=0in 0.5in 0in 0.5in,clip]{dagostino-mcgowan_files/figure-latex/figco-1} \caption{A plot of cohesion versus distance for the data.
+\begin{figure}
+\includegraphics{dagostino-mcgowan_files/figure-latex/figco-1} \caption{A plot of cohesion versus distance for the data.
 The identified communities are colored as in Figure \ref{fig:figculture}.}\label{fig:figco}

### 2025-02-03 2025-r-journal/dagostino-mcgowan.R

@@ -1,0 +1,276 @@
+# Generated by `rjournal_article()` using `knitr::purl()`: do not edit by hand
+# Please edit dagostino-mcgowan.Rmd to modify this file
+
+## ----echo = FALSE----
+knitr::opts_chunk$set(fig.height = 4, message = FALSE, warning = FALSE)
+
+
+## -----------------
+library(pald)
+df <- data.frame(
+  x1 = c(6, 8, 8, 16, 4, 14),
+  x2 = c(5, 4, 10, 8, 4, 10)
+)
+rownames(df) <- c("A", "B", "C", "D", "E", "F")
+
+
+## -----------------
+d <- dist(df)
+
+
+## -----------------
+cohesion_matrix(d)
+
+
+## -----------------
+df |>
+  dist() |>
+  cohesion_matrix()
+
+
+## -----------------
+df |>
+  dist() |>
+  cohesion_matrix() -> cohesion
+
+
+## -----------------
+local_depths(cohesion)
+
+
+## -----------------
+strong_threshold(cohesion)
+
+
+## -----------------
+cohesion_strong(cohesion)
+
+
+## -----------------
+graphs <- community_graphs(cohesion)
+graphs[["G_strong"]]
+
+
+## -----------------
+any_isolated(cohesion)
+
+
+## -----------------
+community_clusters(cohesion)
+
+
+## ----fig1, fig.cap = "Visualization of the points from the data frame `df`", message = FALSE, warning = FALSE----
+library(ggplot2)
+ggplot(df, aes(x1, x2)) +
+  geom_text(label = rownames(df)) + 
+  coord_fixed(ratio = 1) + 
+  xlim(c(4, 16)) + 
+  ylim(c(4, 16))
+
+
+## ----fig2, fig.cap = "PaLD graph displaying the relationship between the points in data frame `df`"----
+plot_community_graphs(cohesion, 
+                      vertex.label.cex = 2,
+                      vertex.label.dist = 0.9)
+
+
+## ----fig3, fig.cap = "PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1"----
+par(pty = "s")
+
+plot_community_graphs(cohesion, 
+                      layout = as.matrix(df),
+                      vertex.size = 100,
+                      vertex.label.color = "white",
+                      axes = TRUE,
+                      rescale = FALSE,
+                      asp = 1,
[…]
